---
title: "Sentence Memorability Experiment"
author: "Team:- Metamatics"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    latex_engine: xelatex
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    code_folding: show
    fig_caption: true
    highlight: tango
    df_print: paged
subtitle: "Testing Hypotheses 1 and 2: Role of Lexical Condition and Grammatical Voice in Recognition Memory"
---
```{r setup, include=FALSE}


knitr::opts_chunk$set(
  echo       = TRUE,
  message    = FALSE,
  warning    = FALSE,
  fig.align  = "center",
  fig.width  = 10,
  fig.height = 6,
  out.width  = "95%",
  dpi        = 150
)

# Our chosen colour scheme for plots throughout the report
PALETTE_4 <- c("#6C5CE7", "#00B894", "#E17055", "#0984E3")
PALETTE_2 <- c("#6C5CE7", "#00B894")
```

---

# Introduction

## Background

This report was prepared by **Team Metamatics** as part of our research
project on sentence memorability. We ran a **continuous recognition experiment**
to explore how two factors shape whether people remember a sentence:

1. **Lexical condition** -- do the nouns in a sentence have high or low memorability?
2. **Grammatical voice** -- is the sentence Active or Passive?

Participants read a stream of simple Subject-Verb-Object (SVO) sentences.
Each sentence appeared twice: first in a study phase, then in a test phase.
Their job was to spot repeated sentences and to say whether the wording
had stayed the same.

## Experimental Procedure

- **Study Phase**: A sentence appears on screen for the first time.
- **Test Phase**: The sentence appears again. The participant must:
  - Press **Spacebar** (IR = Immediate Recognition) if they recognise it as a repeat.
  - Press **Yes/No** (WR = Wording Recognition) to judge if the wording has changed.

We built sentences by pairing nouns that differ in memorability:

| Code | Subject Noun | Object Noun | Example |
|------|-------------|-------------|---------|
| **HH** | High | High | *The doctor examined the patient.* |
| **HL** | High | Low | *The teacher read the passage.* |
| **LH** | Low | High | *The device recorded the music.* |
| **LL** | Low | Low | *The item covered the surface.* |

## Our Hypotheses

> **Hypothesis 1**: Corrected memorability will differ across the four noun
> categories (HH, HL, LH, LL). Sentences made from more memorable nouns
> should be easier to recognise.

> **Hypothesis 2**: Corrected memorability will differ between Active and
> Passive voice. We want to find out if voice alone changes how well
> people remember a sentence.

## Our Key Measure: Corrected Memorability

$$CM = \text{Hit Rate} - \text{False Alarm Rate}$$

This simple subtraction removes guessing bias: we take the proportion of
correctly spotted repeats and subtract the proportion of new sentences
that were wrongly called "old".

---

# Setup and Data Loading

## Libraries We Need

```{r packages}
# Auto-install anything that is missing
pkgs <- c(
  "dplyr", "tidyr", "ggplot2", "readr", "stringr", "purrr",
  "knitr", "kableExtra", "RColorBrewer", "scales"
)
for (p in pkgs) {
  if (!requireNamespace(p, quietly = TRUE)) {
    install.packages(p, repos = "https://cran.r-project.org")
  }
}

library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(stringr)
library(purrr)
library(knitr)
library(kableExtra)
library(scales)

# A minimal, modern ggplot theme we will reuse everywhere
theme_set(theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15),
    plot.subtitle = element_text(color = "grey40"),
    legend.position = "bottom"
  ))
```

## Reading the Log Files

Our dataset lives in **114 separate log files** -- one for each participant.
Every file records all the events that happened during that person's session.

```{r load-data}
log_dir <- file.path("Sentence Memorability", "NewLogsAnonymized")
output_dir <- "R_output"
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

log_files <- list.files(log_dir, pattern = "\\.log$", full.names = TRUE)
cat(sprintf("Found %d participant log files.\n", length(log_files)))

# Load every file and stack them into a single table
raw_data <- log_files %>%
  map_dfr(~ tryCatch(
    read_csv(.x, show_col_types = FALSE, col_types = cols(.default = "c")),
    error = function(e) tibble()
  ))

cat(sprintf("Total rows loaded: %s\n", format(nrow(raw_data), big.mark = ",")))
```

---

# Data Cleaning

## Step 1: Remove Practice Trials

Practice trials let people get used to the task before the real experiment.
We drop them so they do not skew our results.

```{r remove-practice}
# Count and remove practice events
n_practice <- sum(str_starts(raw_data$Event, "Practice") & !is.na(raw_data$Event))
raw_data <- raw_data %>% filter(!str_starts(Event, "Practice") | is.na(Event))

cat(sprintf("Removed %d practice events.\n", n_practice))
cat(sprintf("Remaining rows: %s\n", format(nrow(raw_data), big.mark = ",")))

# Save cleaned data
write_csv(raw_data, file.path(output_dir, "step1_cleaned_data.csv"))
cat("Saved: step1_cleaned_data.csv\n")
```

## Step 2: Parse Stimulus Codes

Every stimulus has a code like `"HH_172_A"`. The last letter is the voice
(A = Active, P = Passive), and everything before it is the sentence ID.
We split these apart so we can group by condition and voice later.

```{r parse-stimulus}
# Parse each stimulus into Core_ID and Voice
parse_stimulus <- function(stim) {
  if (is.na(stim)) {
    return(c(NA_character_, NA_character_))
  }
  parts <- str_split(stim, "_")[[1]]
  n <- length(parts)
  if (n >= 2) {
    return(c(paste(parts[1:(n - 1)], collapse = "_"), parts[n]))
  }
  return(c(stim, NA_character_))
}

# Apply to all rows — use Stimulus column
stim_col <- raw_data[["Stimulus"]]
if (is.null(stim_col)) {
  # Fallback: try to find a column containing stimulus info
  stim_col <- rep(NA_character_, nrow(raw_data))
  warning("Stimulus column not found — using NA")
}

parsed <- t(sapply(stim_col, parse_stimulus, USE.NAMES = FALSE))
raw_data$Core_ID <- parsed[, 1]
raw_data$Voice <- parsed[, 2]

cat("Stimulus codes parsed successfully.\n")
cat(sprintf("Unique voices found: %s\n", paste(unique(na.omit(raw_data$Voice)), collapse = ", ")))
```

## Step 3: Convert Data Types

All columns arrive as text because we read every file with `col_types = "c"`.
Here we turn numbers into actual numbers and flag columns into TRUE/FALSE.

```{r convert-types}
# Turn text columns into the right types
# NA in a boolean column means the event did not happen, so we treat it as FALSE
raw_data <- raw_data %>%
  mutate(
    participant_ID   = as.integer(participant_ID),
    Timestamp        = as.numeric(Timestamp),
    `Accuracy IR`    = suppressWarnings(as.numeric(`Accuracy IR`)),
    `Accuracy WR`    = suppressWarnings(as.numeric(`Accuracy WR`)),
    Reaction_time_IR = suppressWarnings(as.numeric(Reaction_time_IR)),
    Reaction_time_WR = suppressWarnings(as.numeric(Reaction_time_WR)),
    isTarget         = coalesce(tolower(trimws(isTarget)) == "true", FALSE),
    isValidation     = coalesce(tolower(trimws(isValidation)) == "true", FALSE),
    isRepeat         = coalesce(tolower(trimws(isRepeat)) == "true", FALSE)
  ) %>%
  arrange(participant_ID, Timestamp)

cat(sprintf(
  "Data types converted. %d unique participants.\n",
  n_distinct(raw_data$participant_ID)
))
```

## Step 4: Reconstruct Trials

Each trial starts when a sentence appears on screen and includes every
response event until the next sentence. We stitch these rows together
into one row per trial.

```{r build-trials}
# Assign trial IDs: each "Sentence shown" starts a new trial
raw_data <- raw_data %>%
  group_by(participant_ID) %>%
  mutate(trial_id = cumsum(Event == "Sentence shown")) %>%
  ungroup() %>%
  filter(trial_id > 0)

# Extract metadata from the "Sentence shown" row
sent_rows <- raw_data %>%
  filter(Event == "Sentence shown") %>%
  group_by(participant_ID, trial_id) %>%
  slice(1) %>%
  ungroup() %>%
  select(
    participant_ID, trial_id, Core_ID, Voice,
    isTarget, isValidation, isRepeat
  )

# Aggregate response events within each trial
trial_agg <- raw_data %>%
  group_by(participant_ID, trial_id) %>%
  summarise(
    ir_pressed = any(str_detect(Event, "IR pressed"), na.rm = TRUE),
    wr_pressed = any(str_detect(Event, "WR pressed"), na.rm = TRUE),
    has_validation = any(str_detect(Event, "Validation"), na.rm = TRUE),
    ir_accuracy = max(`Accuracy IR`, na.rm = TRUE),
    wr_accuracy = max(`Accuracy WR`, na.rm = TRUE),
    ir_rt = suppressWarnings(min(Reaction_time_IR, na.rm = TRUE)),
    wr_rt = suppressWarnings(min(Reaction_time_WR, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  mutate(across(
    c(ir_accuracy, wr_accuracy, ir_rt, wr_rt),
    ~ ifelse(is.infinite(.), NA_real_, .)
  ))

# Combine metadata + responses
trials <- sent_rows %>%
  inner_join(trial_agg, by = c("participant_ID", "trial_id")) %>%
  mutate(
    is_validation = isValidation | has_validation,
    is_target     = isTarget,
    is_repeat     = isRepeat
  )

cat(sprintf(
  "Reconstructed %s trials from %d participants.\n",
  format(nrow(trials), big.mark = ","), n_distinct(trials$participant_ID)
))

# Save trial-level data
write_csv(trials, file.path(output_dir, "step2_trials.csv"))
cat("Saved: step2_trials.csv\n")
```

---

# Participant Exclusion

## Validation-Based Quality Control

Our experiment has built-in **validation filler sentences**. These are
easy catch trials that tell us whether a participant was actually paying
attention. We use a simple rule to decide who stays and who gets dropped.

**Exclusion rule** -- a participant passes only if:

$$\text{correct\_ir} > \frac{\text{wrong\_ir}}{2} + \text{missed\_ir}$$

```{r exclusion}
val_trials <- trials %>% filter(is_validation)

validation_summary <- val_trials %>%
  group_by(participant_ID) %>%
  summarise(
    correct_ir = sum(is_repeat == TRUE & ir_pressed == TRUE, na.rm = TRUE),
    wrong_ir = sum(is_repeat == FALSE & ir_pressed == TRUE, na.rm = TRUE),
    missed_ir = sum(is_repeat == TRUE & ir_pressed == FALSE, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(passes = correct_ir > (wrong_ir / 2) + missed_ir)

valid_ids <- validation_summary %>%
  filter(passes) %>%
  pull(participant_ID)
excluded_ids <- validation_summary %>%
  filter(!passes) %>%
  pull(participant_ID)

# Display validation results table
validation_summary %>%
  mutate(Status = ifelse(passes, "KEPT", "EXCLUDED")) %>%
  select(participant_ID, correct_ir, wrong_ir, missed_ir, Status) %>%
  kable(
    caption = "Participant Validation Results",
    col.names = c("Participant", "Correct IR", "Wrong IR", "Missed IR", "Status"),
    booktabs = TRUE
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    latex_options = c("striped", "scale_down", "hold_position"),
    full_width = FALSE, position = "center"
  )
```

```{r exclusion-summary}
cat(sprintf("Total participants:    %d\n", nrow(validation_summary)))
cat(sprintf("Participants excluded: %d  %s\n", length(excluded_ids), paste(excluded_ids, collapse = ", ")))
cat(sprintf("Participants kept:     %d\n", length(valid_ids)))

# Save validation summary
write_csv(validation_summary, file.path(output_dir, "step3_validation_summary.csv"))
cat("Saved: step3_validation_summary.csv\n")

# Filter to valid participants only
trials_valid <- trials %>% filter(participant_ID %in% valid_ids)
```

---

# Computing Corrected Memorability

## Deriving Original Voice

When a sentence first appears (study phase), it is in either Active or
Passive voice. We call this the **Original Voice**. We grab it from the
first non-repeat presentation of each sentence for every participant.

```{r original-voice}
# Get Original Voice from first (non-repeat) presentation of each Core_ID
first_pres <- trials_valid %>%
  filter(!is_repeat) %>%
  distinct(participant_ID, Core_ID, .keep_all = TRUE) %>%
  select(participant_ID, Core_ID, Voice) %>%
  rename(Original_Voice = Voice)

# Attach Original Voice to valid trials
trials_valid <- trials_valid %>%
  left_join(first_pres, by = c("participant_ID", "Core_ID"))

cat(sprintf(
  "Original Voice attached. Unique voices: %s\n",
  paste(unique(na.omit(trials_valid$Original_Voice)), collapse = ", ")
))
```

## Mapping Lexical Conditions

We tag each sentence with its lexical condition based on the prefix
of its Core_ID (HH, HVL, LVH, LVL, or HF for fillers).

```{r map-conditions}
get_condition <- function(core_id) {
  if (is.na(core_id)) {
    return("Other")
  }
  prefix <- str_to_upper(str_extract(core_id, "^[A-Za-z]+"))
  mapping <- c("HH" = "HH", "HVL" = "HL", "LVH" = "LH", "LVL" = "LL", "HF" = "Filler")
  result <- mapping[prefix]
  ifelse(is.na(result), "Other", result)
}

trials_valid <- trials_valid %>%
  mutate(label_group = sapply(Core_ID, get_condition))

# Show distribution of conditions
trials_valid %>%
  count(label_group, name = "Trial_Count") %>%
  kable(
    caption = "Number of Trials per Condition",
    col.names = c("Condition", "Number of Trials"),
    booktabs = TRUE
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    latex_options = c("striped", "hold_position"),
    full_width = FALSE, position = "center"
  )
```

We now calculate our main dependent variable. For each participant and
each Original Voice, we work out:

- **Hit Rate** = proportion of repeated sentences correctly recognised
- **FA Rate** = proportion of new sentences wrongly called "old"
  (computed from raw IR-pressed events with Accuracy IR == 0)
- **Corrected Memorability** = Hit Rate - FA Rate

```{r compute-cm-voice}
# Keep only trials with valid Original Voice
voice_trials <- trials_valid %>%
  filter(Original_Voice %in% c("A", "P"))

# --- Hits: from repeat trials grouped by participant x Original_Voice ---
hit_data <- voice_trials %>%
  group_by(participant_ID, Original_Voice) %>%
  summarise(
    n_old = sum(is_repeat == TRUE, na.rm = TRUE),
    hits = sum(is_repeat == TRUE & ir_pressed == TRUE, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(hit_rate = ifelse(n_old > 0, hits / n_old, NA_real_))

# --- FAs: from raw IR pressed events where Accuracy IR == 0 ---
# (kept participants only, using raw_data not trial-level)
raw_kept <- raw_data %>%
  filter(participant_ID %in% valid_ids)

# All "IR pressed" events
ir_all <- raw_kept %>%
  filter(Event == "IR pressed") %>%
  left_join(first_pres, by = c("participant_ID", "Core_ID"))

# All first-presentation (non-repeat) Sentence shown events (FA denominator)
first_shown <- raw_kept %>%
  filter(Event == "Sentence shown", isRepeat == FALSE) %>%
  left_join(first_pres, by = c("participant_ID", "Core_ID"))

# Compute FA rate per participant x Original_Voice
fa_list <- list()
for (pid in valid_ids) {
  for (voice in c("A", "P")) {
    fa_count <- ir_all %>%
      filter(
        participant_ID == pid,
        Original_Voice == voice,
        `Accuracy IR` == 0
      ) %>%
      nrow()
    nontarget <- first_shown %>%
      filter(
        participant_ID == pid,
        Original_Voice == voice
      ) %>%
      nrow()
    fa_rate <- ifelse(nontarget > 0, fa_count / nontarget, NA_real_)
    fa_list[[length(fa_list) + 1]] <- tibble(
      participant_ID = pid,
      Original_Voice = voice,
      false_alarms = fa_count,
      n_new = nontarget,
      fa_rate = fa_rate
    )
  }
}
fa_df <- bind_rows(fa_list)

# Merge hits and FAs
mem_df <- hit_data %>%
  full_join(fa_df, by = c("participant_ID", "Original_Voice")) %>%
  mutate(corrected_memorability = hit_rate - fa_rate)

# Quick overview of hit/FA/CM for each voice
cat("\n=== Corrected Memorability per participant x voice ===\n")
for (v in c("A", "P")) {
  sub <- mem_df %>% filter(Original_Voice == v)
  cat(sprintf("\nVoice %s:\n", v))
  cat(sprintf("  Mean Hit Rate:    %.4f\n", mean(sub$hit_rate, na.rm = TRUE)))
  cat(sprintf("  Mean FA Rate:     %.4f\n", mean(sub$fa_rate, na.rm = TRUE)))
  cat(sprintf("  Mean Corrected M: %.4f\n", mean(sub$corrected_memorability, na.rm = TRUE)))
}

# Save mem_df
write_csv(mem_df, file.path(output_dir, "step4_mem_df.csv"))
cat("Saved: step4_mem_df.csv\n")
```

## Descriptive Statistics by Original Voice

```{r voice-describe}
# Full descriptive statistics table for each voice
voice_describe <- mem_df %>%
  pivot_longer(
    cols = c(hit_rate, fa_rate, corrected_memorability),
    names_to = "Metric", values_to = "Value"
  ) %>%
  group_by(Original_Voice, Metric) %>%
  summarise(
    Count = sum(!is.na(Value)),
    Mean = round(mean(Value, na.rm = TRUE), 6),
    SD = round(sd(Value, na.rm = TRUE), 6),
    Min = round(min(Value, na.rm = TRUE), 6),
    Q25 = round(quantile(Value, 0.25, na.rm = TRUE), 6),
    Median = round(median(Value, na.rm = TRUE), 6),
    Q75 = round(quantile(Value, 0.75, na.rm = TRUE), 6),
    Max = round(max(Value, na.rm = TRUE), 6),
    .groups = "drop"
  ) %>%
  mutate(
    Metric = factor(Metric,
      levels = c("hit_rate", "fa_rate", "corrected_memorability"),
      labels = c("Hit Rate", "FA Rate", "Corrected Memorability")
    )
  )

voice_describe %>%
  kable(
    caption = "Descriptive Statistics -- Hit Rate, FA Rate, and CM by Original Voice",
    col.names = c("Voice", "Metric", "N", "Mean", "SD", "Min", "Q1", "Median", "Q3", "Max"),
    booktabs = TRUE,
    longtable = TRUE
  ) %>%
  kable_styling(
    latex_options = c("striped", "repeat_header", "scale_down"),
    full_width = FALSE,
    position = "center",
    font_size = 10
  )
```

## Linking Study and Test Trials (WR Accuracy)

```{r build-analysis-df}
# Study Events: First presentation (non-repeat)
study_df <- trials_valid %>%
  filter(!is_repeat) %>%
  distinct(participant_ID, Core_ID, .keep_all = TRUE) %>%
  select(participant_ID, Core_ID, Voice) %>%
  rename(Original_Voice = Voice)

# Test Events: Repeat trials where WR was pressed
test_df <- trials_valid %>%
  filter(is_repeat == TRUE, wr_pressed == TRUE) %>%
  distinct(participant_ID, Core_ID, .keep_all = TRUE) %>%
  select(participant_ID, Core_ID, Voice, wr_accuracy, wr_rt) %>%
  rename(Test_Voice = Voice, `Accuracy WR` = wr_accuracy, Reaction_time_WR = wr_rt)

# Merge Study and Test
analysis_df <- test_df %>%
  inner_join(study_df, by = c("participant_ID", "Core_ID")) %>%
  mutate(Condition = ifelse(Original_Voice == Test_Voice, "Identical", "Transformed")) %>%
  filter(Original_Voice %in% c("A", "P"))

# Merge corrected memorability from mem_df
analysis_df <- analysis_df %>%
  left_join(
    mem_df %>% select(participant_ID, Original_Voice, corrected_memorability, hit_rate, fa_rate),
    by = c("participant_ID", "Original_Voice")
  )

cat(sprintf("\nTotal matched test trials: %d\n", nrow(analysis_df)))
cat(sprintf("Unique participants: %d\n", n_distinct(analysis_df$participant_ID)))
cat(sprintf("\nOriginal Voice counts:\n"))
print(table(analysis_df$Original_Voice))
cat(sprintf("\nCondition counts:\n"))
print(table(analysis_df$Condition))

# Save analysis_df
write_csv(analysis_df, file.path(output_dir, "step4_analysis_df.csv"))
cat("Saved: step4_analysis_df.csv\n")
```

## Computing CM per Condition x Voice (for H1 and H2)

```{r compute-sentence-mem}
# Filter to target conditions only (exclude fillers and validation)
target <- trials_valid %>%
  filter(
    !is_validation,
    label_group %in% c("HH", "HL", "LH", "LL"),
    Voice %in% c("A", "P")
  )

sentence_mem <- target %>%
  group_by(participant_ID, label_group, Voice) %>%
  summarise(
    n_old = sum(is_repeat == TRUE, na.rm = TRUE),
    n_new = sum(is_repeat == FALSE, na.rm = TRUE),
    hits = sum(is_repeat == TRUE & ir_pressed == TRUE, na.rm = TRUE),
    false_alarms = sum(is_repeat == FALSE & ir_pressed == TRUE, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    hit_rate               = ifelse(n_old > 0, hits / n_old, NA_real_),
    false_alarm_rate       = ifelse(n_new > 0, false_alarms / n_new, NA_real_),
    corrected_memorability = hit_rate - false_alarm_rate
  )

# Save detailed memorability data
write_csv(sentence_mem, file.path(output_dir, "step4_sentence_memorability.csv"))
cat(sprintf("sentence_mem: %d rows\n", nrow(sentence_mem)))
```

## Preparing Data for Each Hypothesis

```{r prepare-hypotheses}
COND_ORDER <- c("HH", "HL", "LH", "LL")

# H1: average across voice so we get one score per condition per participant
mem_h1 <- sentence_mem %>%
  group_by(participant_ID, label_group) %>%
  summarise(
    corrected_memorability = mean(corrected_memorability, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(label_group = factor(label_group, levels = COND_ORDER))

# H2: average across condition so we get one score per voice per participant
mem_h2 <- sentence_mem %>%
  group_by(participant_ID, Voice) %>%
  summarise(
    corrected_memorability = mean(corrected_memorability, na.rm = TRUE),
    .groups = "drop"
  )

write_csv(mem_h1, file.path(output_dir, "step5_mem_h1.csv"))
write_csv(mem_h2, file.path(output_dir, "step5_mem_h2.csv"))
cat(sprintf("H1 data: %d rows | H2 data: %d rows\n", nrow(mem_h1), nrow(mem_h2)))
```

---

# Hypothesis 1: Effect of Lexical Condition

## Descriptive Statistics

Here is a quick look at the average corrected memorability in each of the
four noun-memorability groups.

```{r h1-descriptive}
h1_desc <- mem_h1 %>%
  group_by(label_group) %>%
  summarise(
    N = n(),
    Mean = round(mean(corrected_memorability, na.rm = TRUE), 3),
    Median = round(median(corrected_memorability, na.rm = TRUE), 3),
    SD = round(sd(corrected_memorability, na.rm = TRUE), 3),
    Min = round(min(corrected_memorability, na.rm = TRUE), 3),
    Max = round(max(corrected_memorability, na.rm = TRUE), 3),
    .groups = "drop"
  )

h1_desc %>%
  kable(
    caption = "Descriptive Statistics -- Corrected Memorability by Lexical Condition",
    col.names = c("Condition", "N", "Mean", "Median", "SD", "Min", "Max"),
    booktabs = TRUE
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "bordered"),
    latex_options = c("striped", "hold_position"),
    full_width = FALSE, position = "center"
  )
```

## Normality Testing

We need to know if the data are roughly normal before picking a test.
The Shapiro-Wilk test checks this for each group. If the p-value is
below 0.05, we cannot assume normality.

```{r h1-shapiro}
# Safe wrapper for shapiro.test (handles small samples)
safe_shapiro <- function(x) {
  x <- x[!is.na(x)]
  if (length(x) < 3 || length(x) > 5000) {
    return(data.frame(W = NA_real_, p_value = NA_real_))
  }
  st <- shapiro.test(x)
  data.frame(W = round(as.numeric(st$statistic), 4), p_value = as.numeric(st$p.value))
}

sw_h1 <- mem_h1 %>%
  group_by(label_group) %>%
  summarise(
    N = n(), sw = list(safe_shapiro(corrected_memorability)),
    .groups = "drop"
  ) %>%
  unnest(sw) %>%
  mutate(
    Normal = case_when(is.na(p_value) ~ "N/A", p_value > 0.05 ~ "Yes", TRUE ~ "No"),
    p_display = case_when(
      is.na(p_value) ~ "N/A", p_value < 0.001 ~ "< 0.001",
      TRUE ~ as.character(round(p_value, 4))
    )
  )

sw_h1 %>%
  select(label_group, N, W, p_display, Normal) %>%
  kable(
    caption = "Shapiro-Wilk Normality Test per Lexical Condition",
    col.names = c("Condition", "N", "W", "p-value", "Normal?"),
    booktabs = TRUE
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    latex_options = c("striped", "hold_position"),
    full_width = FALSE, position = "center"
  )
```

## Q-Q Plots for Normality Assessment

Another way to check normality visually. If the dots sit close to the
straight line, the data are roughly normal.

```{r h1-qqplot, fig.cap="Q-Q Plots of Corrected Memorability by Lexical Condition."}
p_qq_h1 <- ggplot(mem_h1, aes(sample = corrected_memorability)) +
  stat_qq(colour = "#6C5CE7", alpha = 0.6, size = 2) +
  stat_qq_line(colour = "#E17055", linewidth = 0.8) +
  facet_wrap(~label_group, scales = "free") +
  labs(
    title = "Q-Q Plots -- Corrected Memorability by Lexical Condition",
    x = "Theoretical Quantiles", y = "Sample Quantiles"
  )

p_qq_h1
ggsave(file.path(output_dir, "H1_QQ_plots.png"), p_qq_h1, width = 10, height = 8, dpi = 150)
cat("Saved: H1_QQ_plots.png\n")
```

## Visualization: Distribution of Corrected Memorability

The plot below combines a violin shape (showing the distribution) with a
box plot and individual data points. Wider parts of the violin mean more
participants scored around that level.

```{r h1-violin, fig.cap="Corrected Memorability across Lexical Conditions. Each dot represents one participant's mean score."}
p_h1 <- ggplot(mem_h1, aes(
  x = label_group, y = corrected_memorability,
  fill = label_group
)) +
  geom_violin(alpha = 0.5, colour = NA, width = 0.9) +
  geom_boxplot(width = 0.15, outlier.alpha = 0.3, fill = "white") +
  geom_jitter(width = 0.08, alpha = 0.4, size = 1.8, color = "grey30") +
  scale_fill_manual(values = PALETTE_4) +
  labs(
    title = "Corrected Memorability by Lexical Condition",
    subtitle = "Hypothesis 1: Does noun memorability affect sentence recognition?",
    x = "Lexical Condition", y = "Corrected Memorability (Hit Rate − FA Rate)"
  ) +
  theme(legend.position = "none")

p_h1
ggsave(file.path(output_dir, "H1_violin_box.png"), p_h1, width = 10, height = 6, dpi = 150)
cat("Saved: H1_violin_box.png\n")
```

## Statistical Test: Kruskal-Wallis

Since some groups show departures from normality, we use the **Kruskal-Wallis test** —
the non-parametric equivalent of one-way ANOVA. This tests whether at least
two of the four conditions differ significantly.

```{r h1-kruskal}
kw_test <- kruskal.test(corrected_memorability ~ label_group, data = mem_h1)

# Effect size: Epsilon-squared
n_total <- nrow(mem_h1)
k <- length(COND_ORDER)
epsilon_sq <- round((as.numeric(kw_test$statistic) - k + 1) / (n_total - k), 4)

tibble(
  Test       = "Kruskal-Wallis",
  Statistic  = round(as.numeric(kw_test$statistic), 3),
  df         = as.numeric(kw_test$parameter),
  `p-value`  = round(as.numeric(kw_test$p.value), 4),
  `Epsilon2` = epsilon_sq,
  Decision   = ifelse(as.numeric(kw_test$p.value) < 0.05, "Reject H0", "Retain H0")
) %>%
  kable(caption = "Kruskal-Wallis Test -- H1", booktabs = TRUE) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "bordered"),
    latex_options = c("striped", "hold_position"),
    full_width = FALSE, position = "center"
  )
```

## Post-Hoc Pairwise Comparisons

```{r h1-posthoc}
if (as.numeric(kw_test$p.value) < 0.05) {
  pairs <- combn(COND_ORDER, 2, simplify = FALSE)
  posthoc <- map_dfr(pairs, function(pr) {
    x1 <- mem_h1 %>%
      filter(label_group == pr[1]) %>%
      pull(corrected_memorability)
    x2 <- mem_h1 %>%
      filter(label_group == pr[2]) %>%
      pull(corrected_memorability)
    wt <- wilcox.test(x1, x2, exact = FALSE)
    tibble(
      Pair = paste(pr[1], "vs", pr[2]),
      p_raw = round(as.numeric(wt$p.value), 4),
      p_Bonferroni = round(min(as.numeric(wt$p.value) * length(pairs), 1), 4),
      Significant = ifelse(p_Bonferroni < 0.05, "Yes *", "No")
    )
  })

  posthoc %>%
    kable(
      caption = "Post-Hoc Pairwise Wilcoxon Tests (Bonferroni corrected)",
      col.names = c("Comparison", "p (raw)", "p (Bonferroni)", "Significant?"),
      booktabs = TRUE
    ) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover"),
      latex_options = c("striped", "hold_position"),
      full_width = FALSE, position = "center"
    )
} else {
  cat("Kruskal-Wallis was not significant -- no follow-up comparisons needed.\n")
}
```

---

# Hypothesis 2: Effect of Grammatical Voice

## Descriptive Statistics

```{r h2-descriptive}
h2_desc <- mem_h2 %>%
  group_by(Voice) %>%
  summarise(
    N = n(),
    Mean = round(mean(corrected_memorability, na.rm = TRUE), 3),
    Median = round(median(corrected_memorability, na.rm = TRUE), 3),
    SD = round(sd(corrected_memorability, na.rm = TRUE), 3),
    .groups = "drop"
  ) %>%
  mutate(Voice = c(A = "Active", P = "Passive")[Voice])

h2_desc %>%
  kable(
    caption = "Descriptive Statistics -- CM by Grammatical Voice",
    col.names = c("Voice", "N", "Mean", "Median", "SD"),
    booktabs = TRUE
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "bordered"),
    latex_options = c("striped", "hold_position"),
    full_width = FALSE, position = "center"
  )
```

## Paired Differences

Since every participant has a score for both Active and Passive, we can
work out each person's difference (Active minus Passive) and test whether
that difference is significantly different from zero.

```{r h2-paired}
h2_wide <- mem_h2 %>%
  pivot_wider(names_from = Voice, values_from = corrected_memorability) %>%
  drop_na(A, P) %>%
  mutate(diff = A - P)

cat(sprintf("Paired participants: %d\n", nrow(h2_wide)))
cat(sprintf("Mean difference (A - P): %.4f\n", mean(h2_wide$diff)))
```

## Normality of Differences

```{r h2-normality}
sw_diff <- shapiro.test(h2_wide$diff)

tibble(
  Test = "Shapiro-Wilk (on differences)",
  W = round(as.numeric(sw_diff$statistic), 4),
  p_value = round(as.numeric(sw_diff$p.value), 4),
  Normal = ifelse(as.numeric(sw_diff$p.value) > 0.05, "Yes", "No"),
  `Recommended Test` = ifelse(as.numeric(sw_diff$p.value) > 0.05, "Paired t-test", "Wilcoxon signed-rank")
) %>%
  kable(caption = "Normality Test on Paired Differences (A - P)", booktabs = TRUE) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    latex_options = c("striped", "scale_down", "hold_position"),
    full_width = FALSE, position = "center"
  )
```

## Q-Q Plot of Paired Differences

```{r h2-qqplot, fig.cap="Q-Q Plot of paired differences (Active - Passive). Departures from the line suggest non-normality."}
p_qq_h2 <- ggplot(h2_wide, aes(sample = diff)) +
  stat_qq(colour = "#6C5CE7", alpha = 0.6, size = 2) +
  stat_qq_line(colour = "#E17055", linewidth = 0.8) +
  labs(
    title = "Q-Q Plot -- Paired Differences (Active - Passive)",
    x = "Theoretical Quantiles", y = "Sample Quantiles"
  )

p_qq_h2
ggsave(file.path(output_dir, "H2_QQ_plot.png"), p_qq_h2, width = 8, height = 6, dpi = 150)
cat("Saved: H2_QQ_plot.png\n")
```

## Visualization: Voice Comparison

```{r h2-violin, fig.cap="Corrected Memorability by Grammatical Voice."}
mem_h2_plot <- mem_h2 %>%
  mutate(Voice_label = c(A = "Active", P = "Passive")[Voice])

p_h2 <- ggplot(mem_h2_plot, aes(
  x = Voice_label, y = corrected_memorability,
  fill = Voice_label
)) +
  geom_violin(alpha = 0.5, colour = NA, width = 0.8) +
  geom_boxplot(width = 0.15, outlier.alpha = 0.3, fill = "white") +
  geom_jitter(width = 0.08, alpha = 0.4, size = 1.8, color = "grey30") +
  scale_fill_manual(values = PALETTE_2) +
  labs(
    title = "Corrected Memorability by Grammatical Voice",
    subtitle = "Hypothesis 2: Does Active vs Passive voice affect recognition?",
    x = "Voice", y = "Corrected Memorability"
  ) +
  theme(legend.position = "none")

p_h2
ggsave(file.path(output_dir, "H2_violin_box.png"), p_h2, width = 8, height = 6, dpi = 150)
cat("Saved: H2_violin_box.png\n")
```

```{r h2-histogram, fig.cap="Distribution of paired differences (Active − Passive). Red line = zero, orange line = observed mean."}
p_hist <- ggplot(h2_wide, aes(x = diff)) +
  geom_histogram(bins = 20, fill = "#6C5CE7", colour = "white", alpha = 0.8) +
  geom_vline(xintercept = 0, colour = "red", linetype = "dashed", linewidth = 1) +
  geom_vline(xintercept = mean(h2_wide$diff), colour = "#E17055", linewidth = 1.2) +
  labs(
    title = "Distribution of Paired Differences (Active − Passive)",
    subtitle = "Orange line = observed mean difference",
    x = "Active − Passive CM", y = "Count"
  )

p_hist
ggsave(file.path(output_dir, "H2_histogram_diff.png"), p_hist, width = 8, height = 5, dpi = 150)
cat("Saved: H2_histogram_diff.png\n")
```

## Statistical Test: Wilcoxon Signed-Rank

```{r h2-test}
w_test <- wilcox.test(h2_wide$A, h2_wide$P, paired = TRUE, exact = FALSE)

tibble(
  Test      = "Wilcoxon Signed-Rank (paired)",
  W         = round(as.numeric(w_test$statistic), 3),
  `p-value` = round(as.numeric(w_test$p.value), 4),
  Decision  = ifelse(as.numeric(w_test$p.value) < 0.05, "Reject H0", "Retain H0")
) %>%
  kable(caption = "Wilcoxon Signed-Rank Test -- H2", booktabs = TRUE) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "bordered"),
    latex_options = c("striped", "hold_position"),
    full_width = FALSE, position = "center"
  )
```

---

# Reaction Time Analysis

We also looked at how fast people responded when they correctly
identified a repeated sentence. The y-axis starts at zero so the
scale gives an honest picture.

```{r rt-data}
rt_data <- target %>%
  filter(is_repeat, ir_pressed, !is.na(ir_rt), ir_rt > 0) %>%
  mutate(
    label_group = factor(label_group, levels = COND_ORDER),
    Voice_label = ifelse(Voice == "A", "Active", "Passive")
  )
```

```{r rt-condition, fig.cap="IR Reaction Time by Lexical Condition (y-axis from 0)."}
p_rt1 <- ggplot(rt_data, aes(x = label_group, y = ir_rt, fill = label_group)) +
  geom_violin(alpha = 0.45, colour = NA) +
  geom_boxplot(width = 0.15, outlier.alpha = 0.3, fill = "white") +
  scale_fill_manual(values = PALETTE_4) +
  scale_y_continuous(limits = c(0, NA), labels = comma) +
  labs(
    title = "IR Reaction Time by Lexical Condition",
    x = "Condition", y = "Reaction Time (ms)"
  ) +
  theme(legend.position = "none")

p_rt1
ggsave(file.path(output_dir, "RT_by_condition.png"), p_rt1, width = 10, height = 6, dpi = 150)
```

```{r rt-voice, fig.cap="IR Reaction Time by Voice (y-axis from 0)."}
p_rt2 <- ggplot(rt_data, aes(x = Voice_label, y = ir_rt, fill = Voice_label)) +
  geom_violin(alpha = 0.45, colour = NA) +
  geom_boxplot(width = 0.15, outlier.alpha = 0.3, fill = "white") +
  scale_fill_manual(values = PALETTE_2) +
  scale_y_continuous(limits = c(0, NA), labels = comma) +
  labs(
    title = "IR Reaction Time by Grammatical Voice",
    x = "Voice", y = "Reaction Time (ms)"
  ) +
  theme(legend.position = "none")

p_rt2
ggsave(file.path(output_dir, "RT_by_voice.png"), p_rt2, width = 8, height = 6, dpi = 150)
```

---

# Conclusion

## Summary of Findings

```{r summary-table}
tibble(
  Hypothesis = c("H1: Lexical Condition", "H2: Grammatical Voice"),
  Test = c("Kruskal-Wallis", "Wilcoxon Signed-Rank"),
  Statistic = c(round(as.numeric(kw_test$statistic), 3), round(as.numeric(w_test$statistic), 3)),
  `p-value` = c(round(as.numeric(kw_test$p.value), 4), round(as.numeric(w_test$p.value), 4)),
  Result = c(
    ifelse(as.numeric(kw_test$p.value) < 0.05, "Significant *", "Not significant"),
    ifelse(as.numeric(w_test$p.value) < 0.05, "Significant *", "Not significant")
  )
) %>%
  kable(caption = "Summary of All Hypothesis Tests", booktabs = TRUE) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "bordered"),
    latex_options = c("striped", "hold_position"),
    full_width = FALSE, position = "center"
  )
```

## Interpretation

```{r conclusion-text, results='asis'}
# ----- Hypothesis 1 conclusion -----
kw_p <- as.numeric(kw_test$p.value)

if (kw_p < 0.05) {
  cat("**Hypothesis 1 -- Lexical Condition:**\n\n")
  cat(sprintf("The Kruskal-Wallis test was **significant** (p = %.4f, p < 0.05).\n", kw_p))
  cat("We therefore **reject the null hypothesis (H0)** and conclude that\n")
  cat("corrected memorability **does differ** across at least two of the four\n")
  cat("lexical conditions (HH, HL, LH, LL). In other words, the memorability\n")
  cat("of the nouns in a sentence appears to affect how well people recognise\n")
  cat("that sentence.\n\n")
  # report significant post-hoc pairs if the posthoc object exists
  if (exists("posthoc")) {
    sig_pairs <- posthoc %>% filter(Significant == "Yes *") %>% pull(Pair)
    if (length(sig_pairs) > 0) {
      cat(sprintf("The post-hoc pairwise Wilcoxon tests (with Bonferroni correction) identified\n"))
      cat(sprintf("the following significant pairs: **%s**.\n\n", paste(sig_pairs, collapse = ", ")))
    } else {
      cat("However, after Bonferroni correction none of the individual pairwise\n")
      cat("comparisons reached significance.\n\n")
    }
  }
} else {
  cat("**Hypothesis 1 -- Lexical Condition:**\n\n")
  cat(sprintf("The Kruskal-Wallis test was **not significant** (p = %.4f, p >= 0.05).\n", kw_p))
  cat("We therefore **retain the null hypothesis (H0)** and conclude that there is\n")
  cat("no statistically significant difference in corrected memorability across the\n")
  cat("four lexical conditions.\n\n")
}

# ----- Hypothesis 2 conclusion -----
w_p <- as.numeric(w_test$p.value)

if (w_p < 0.05) {
  cat("**Hypothesis 2 -- Grammatical Voice:**\n\n")
  cat(sprintf("The Wilcoxon signed-rank test was **significant** (p = %.4f, p < 0.05).\n", w_p))
  cat("We therefore **reject the null hypothesis (H0)** and conclude that there **is**\n")
  cat("a statistically significant paired difference in corrected memorability between\n")
  cat("Active and Passive sentences.\n\n")
} else {
  cat("**Hypothesis 2 -- Grammatical Voice:**\n\n")
  cat(sprintf("The Wilcoxon signed-rank test was **not significant** (p = %.4f, p >= 0.05).\n", w_p))
  cat("We therefore **retain the null hypothesis (H0)**. There is no statistically\n")
  cat("significant paired difference in corrected memorability between Active and\n")
  cat("Passive sentences.\n\n")
}

cat("All findings should be read together with the effect sizes, Q-Q plots,\n")
cat("and the distribution plots shown earlier in this report.\n")
```

---


