---
title: "Hypothesis 5: Does Practice Performance Predict Memory?"
subtitle: "Using Practice Phase Data as a Covariate in ANCOVA"
author: "Team: Metamatics"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    latex_engine: xelatex
    number_sections: true
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    code_folding: show
    fig_caption: true
    highlight: tango
    df_print: paged
fontsize: 11pt
geometry: margin=1in
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{amsmath}
  - \usepackage{amssymb}
  - \usepackage{fontspec}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo      = TRUE,
  warning   = FALSE,
  message   = FALSE,
  fig.align = "center",
  fig.width  = 7,
  fig.height = 4,
  out.width  = "100%",
  dpi        = 150
)

# Install and load all needed packages
pkgs <- c(
  "dplyr", "tidyr", "tibble", "ggplot2", "readr", "stringr", "purrr",
  "knitr", "kableExtra", "broom", "car", "lsr", "patchwork", "scales"
)
for (p in pkgs) {
  if (!requireNamespace(p, quietly = TRUE)) {
    install.packages(p, repos = "https://cran.r-project.org")
  }
  library(p, character.only = TRUE)
}

# Colour palette and theme settings
PAL_H5    <- c("#A23B72", "#2E86AB", "#F18F01", "#C73E1D")
PAL_VOICE <- c("Active" = "#2E86AB", "Passive" = "#A23B72")
HDR_COLOR <- "#1B3A57"

theme_set(
  theme_minimal(base_size = 13) +
    theme(
      plot.title    = element_text(face = "bold", size = 15, colour = "#2C2C2C"),
      plot.subtitle = element_text(colour = "grey45", size = 11),
      legend.position = "bottom",
      panel.grid.minor = element_blank()
    )
)

# Folder paths
log_dir    <- file.path("Sentence Memorability", "NewLogsAnonymized")
output_dir <- "R_output"
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)
```

---

# What is Hypothesis 5?

## The Simple Idea

Before the real experiment started, every participant did a short **practice round**
(about 8 warm-up trials). This was just to help them learn how to press the right
keys. In our earlier analysis (H1 to H4), we simply threw away this practice data.

But here is the key question: **What if the way someone performs during practice
tells us something useful about how they will perform in the real experiment?**

- A person who was **fast and accurate** in practice probably finds the task easy.
- A person who was **slow and made lots of mistakes** in practice probably finds it harder.

If this is true, then practice performance is what we call a **covariate** --- a
background factor that quietly influences the results. By measuring it and
including it in our model, we can remove its effect and get a cleaner picture
of what our real variables (like sentence type and grammatical voice) actually do.

## Where H5 Fits

| Hypothesis | What we tested | Main tool |
|---|---|---|
| H1 | Does noun type (HH/HL/LH/LL) affect memory? | Kruskal-Wallis |
| H2 | Does sentence voice (Active/Passive) affect memory? | Wilcoxon / t-test |
| H3 | Does voice affect how accurately people judge wording? | Paired test |
| H4 | Does voice affect how quickly people respond? | Paired test |
| **H5** | Does practice performance predict final memory scores? | **ANCOVA** |

H5 does not test a new condition. It tests whether **individual differences**
captured in the practice phase carry over and affect the main results.

---

# The Formal Hypothesis

## In Plain Words

> **Null Hypothesis (H0):** How well someone does in practice does NOT predict
> how well they do in the real experiment. Controlling for practice does not
> change the main findings.

> **Alternative Hypothesis (H5):** How well someone does in practice DOES
> predict their final memory score. People who picked up the task quickly in
> practice tend to remember sentences better overall.

## The Math Behind It

We fit a model that looks like this:

$$\text{Memory Score} = \beta_0 + \beta_1 \times \text{Voice} + \beta_2 \times \text{Practice Speed} + \beta_3 \times \text{Practice Errors} + \varepsilon$$

- **Memory Score** = Corrected Memorability (how well each person remembered sentences)
- **Voice** = Was the sentence Active or Passive? (our main variable from H2)
- **Practice Speed** = Average reaction time during practice (in milliseconds)
- **Practice Errors** = Proportion of wrong answers during practice
- **$\varepsilon$** = Leftover unexplained variation (noise)

If $\beta_2$ or $\beta_3$ are statistically significant, then **H5 is supported**.

## Decision Rule

| Setting | Value |
|---|---|
| Significance level (alpha) | 0.05 |
| Effect size measure | Partial eta-squared |
| Statistical test | ANCOVA (Analysis of Covariance) |

---

# Step A: Collecting Practice Data

## Reading the Log Files

We go back to the raw log files but this time we **keep only the practice rows**
(the ones labelled "Practice"), which we had previously removed.

```{r extract-practice}
log_files <- list.files(log_dir, pattern = "\\.log$", full.names = TRUE)
cat(sprintf("Found %d participant log files.\n", length(log_files)))

# Read all files, keep only Practice-phase events
practice_raw <- log_files %>%
  map_dfr(~ {
    tryCatch(
      {
        d <- read_csv(.x, show_col_types = FALSE, col_types = cols(.default = "c"))
        d %>% filter(str_starts(Event, "Practice"))
      },
      error = function(e) tibble()
    )
  }) %>%
  mutate(
    participant_ID   = as.integer(participant_ID),
    `Accuracy IR`    = suppressWarnings(as.numeric(`Accuracy IR`)),
    Reaction_time_IR = suppressWarnings(as.numeric(Reaction_time_IR))
  )

cat(sprintf(
  "Extracted %s practice events from all participants.\n",
  format(nrow(practice_raw), big.mark = ",")
))
cat(sprintf(
  "Number of participants with practice data: %d\n",
  n_distinct(practice_raw$participant_ID)
))
```

## Computing Two Scores Per Participant

For each participant, we calculate two simple numbers from their practice data:

| Score Name | What it measures | What it tells us |
|---|---|---|
| `covariate_RT` | Average time to press the key in practice (ms) | How fast their basic reactions were |
| `covariate_Error` | Fraction of wrong answers in practice | How quickly they grasped the task rules |

```{r compute-covariates}
# Keep only the key-press events from practice
practice_prof <- practice_raw %>%
  filter(str_detect(Event, "IR pressed")) %>%
  group_by(participant_ID) %>%
  summarise(
    practice_trials = n(),
    covariate_RT    = mean(Reaction_time_IR, na.rm = TRUE),
    covariate_Error = 1 - mean(`Accuracy IR`, na.rm = TRUE),
    .groups         = "drop"
  ) %>%
  filter(!is.na(covariate_RT), !is.na(covariate_Error))

write_csv(practice_prof, file.path(output_dir, "H5_practice_covariates.csv"))

cat(sprintf("Practice scores ready for %d participants.\n", nrow(practice_prof)))

# Show a preview of the data
practice_prof %>%
  head(10) %>%
  kable(
    caption   = "Table 1. Preview of Practice Scores (first 10 participants)",
    col.names = c("Participant", "Practice Trials", "Mean RT (ms)", "Error Rate"),
    digits    = 3,
    booktabs  = TRUE
  ) %>%
  kable_styling(
    latex_options     = c("striped", "HOLD_position"),
    bootstrap_options = c("striped", "hover"),
    full_width        = FALSE,
    position          = "center"
  ) %>%
  row_spec(0, bold = TRUE, background = HDR_COLOR, color = "white")
```

## Summary Statistics for Practice Scores

```{r cov-descriptives}
practice_prof %>%
  summarise(
    N          = n(),
    Mean_RT    = round(mean(covariate_RT, na.rm = TRUE), 2),
    SD_RT      = round(sd(covariate_RT, na.rm = TRUE), 2),
    Min_RT     = round(min(covariate_RT, na.rm = TRUE), 2),
    Max_RT     = round(max(covariate_RT, na.rm = TRUE), 2),
    Mean_Error = round(mean(covariate_Error, na.rm = TRUE), 4),
    SD_Error   = round(sd(covariate_Error, na.rm = TRUE), 4),
    Min_Error  = round(min(covariate_Error, na.rm = TRUE), 4),
    Max_Error  = round(max(covariate_Error, na.rm = TRUE), 4)
  ) %>%
  pivot_longer(everything(), names_to = "Statistic", values_to = "Value") %>%
  kable(
    caption  = "Table 2. Summary of Practice Scores Across All Participants",
    digits   = 4,
    booktabs = TRUE
  ) %>%
  kable_styling(
    latex_options     = c("striped", "HOLD_position"),
    bootstrap_options = c("striped", "hover"),
    full_width        = FALSE,
    position          = "center"
  )
```

---

# Step B: Merging Practice Scores with Main Results

We now attach each participant's practice scores to their main experiment memory
scores. We load the memory data produced by the H3/H4 analysis.

```{r merge-main-data}
h3h4_file <- file.path(output_dir, "H3H4_step3_analysis.csv")

if (!file.exists(h3h4_file)) {
  stop(paste(
    "File not found:", h3h4_file,
    "\nPlease knit H3_and_H4_Metametrics.Rmd first."
  ))
}

main_df <- read_csv(h3h4_file, show_col_types = FALSE)

# Get one corrected memorability score per participant per voice condition
cm_by_voice <- main_df %>%
  distinct(participant_ID, Voice_Label, corrected_memorability) %>%
  filter(!is.na(corrected_memorability), Voice_Label %in% c("Active", "Passive"))

# Attach practice scores
ancova_df <- cm_by_voice %>%
  inner_join(practice_prof, by = "participant_ID")

cat(sprintf("Final dataset: %d rows (%d participants x 2 voice conditions)\n",
            nrow(ancova_df), n_distinct(ancova_df$participant_ID)))

write_csv(ancova_df, file.path(output_dir, "H5_ancova_dataset.csv"))
```

---

# Step C: Checking ANCOVA Assumptions

Before we run ANCOVA, we need to check 4 rules. If these rules are broken,
ANCOVA cannot be trusted.

## Rule 1 - Linearity

**What it means:** The relationship between practice scores and memory scores
should follow a straight line (not a curve).

**How we check:** Plot the data and compare a straight line fit with a curved
(LOESS) smoother. If they match closely, we are good.

```{r assumption-linearity, fig.cap="Figure 1. Linearity check. Top: Practice speed vs. memory score. Bottom: Practice error rate vs. memory score. Solid line = straight-line fit. Dashed line = curved fit. They should look similar.", fig.width=7, fig.height=7, out.width="100%"}
p_lin1 <- ggplot(ancova_df, aes(x = covariate_RT, y = corrected_memorability)) +
  geom_point(alpha = 0.5, colour = PAL_H5[2], size = 1.8) +
  geom_smooth(method = "lm",    colour = PAL_H5[1], fill = "grey80",
              se = TRUE,  linewidth = 1.0, linetype = "solid") +
  geom_smooth(method = "loess", colour = PAL_H5[3], fill = NA,
              se = FALSE, linewidth = 0.9, linetype = "dashed") +
  labs(
    title    = "Fig 1a: Practice Speed vs. Memory Score",
    subtitle = "Solid = straight-line fit | Dashed = curved fit",
    x        = "Average Practice Reaction Time (ms)",
    y        = "Corrected Memorability (CM)"
  )

p_lin2 <- ggplot(ancova_df, aes(x = covariate_Error, y = corrected_memorability)) +
  geom_point(alpha = 0.5, colour = PAL_H5[4], size = 1.8) +
  geom_smooth(method = "lm",    colour = PAL_H5[1], fill = "grey80",
              se = TRUE,  linewidth = 1.0, linetype = "solid") +
  geom_smooth(method = "loess", colour = PAL_H5[3], fill = NA,
              se = FALSE, linewidth = 0.9, linetype = "dashed") +
  labs(
    title    = "Fig 1b: Practice Error Rate vs. Memory Score",
    subtitle = "Solid = straight-line fit | Dashed = curved fit",
    x        = "Practice Error Rate (0 = no errors, 1 = all wrong)",
    y        = "Corrected Memorability (CM)"
  )

# Stack vertically so neither plot is clipped
p_lin1 / p_lin2
ggsave(file.path(output_dir, "H5_assumption_linearity.png"),
       p_lin1 / p_lin2, width = 7, height = 7, dpi = 150)
```

> **Reading this plot:** If the solid and dashed lines are close together,
> the linearity assumption is OK. A big difference means the relationship is curved
> and we might need to transform the data.

## Rule 2 - Residuals Should Be Normally Distributed

**What it means:** After the model explains what it can, the leftover errors
(called residuals) should follow a bell-curve shape.

**How we check:** Shapiro-Wilk test. A p-value above 0.05 means we are OK.

```{r assumption-normality}
# Fit a test model and inspect its residuals
prelim_fit <- lm(
  corrected_memorability ~ Voice_Label + covariate_RT + covariate_Error,
  data = ancova_df
)

resid_vals <- residuals(prelim_fit)
sw_test    <- shapiro.test(resid_vals)

tibble(
  Test      = "Shapiro-Wilk test on model residuals",
  W         = round(as.numeric(sw_test$statistic), 4),
  p_value   = round(as.numeric(sw_test$p.value), 4),
  Result    = ifelse(as.numeric(sw_test$p.value) > 0.05,
                     "OK (p > 0.05)", "Not OK (p <= 0.05)")
) %>%
  kable(
    caption   = "Table 3. Are the model residuals normally distributed? (Shapiro-Wilk Test)",
    col.names = c("Test", "W statistic", "p-value", "Conclusion"),
    booktabs  = TRUE
  ) %>%
  kable_styling(
    latex_options     = c("striped", "HOLD_position"),
    bootstrap_options = c("striped", "hover"),
    full_width        = FALSE,
    position          = "center"
  ) %>%
  row_spec(0, bold = TRUE, background = HDR_COLOR, color = "white")
```

```{r assumption-normality-plot, fig.cap="Figure 2. Are the model errors normally distributed? Top: Q-Q plot (points should follow the line). Bottom: Histogram (should be a bell shape).", fig.width=7, fig.height=7, out.width="100%"}
qq_df <- tibble(residuals = resid_vals)

p_qq <- ggplot(qq_df, aes(sample = residuals)) +
  stat_qq(colour = PAL_H5[2], alpha = 0.6, size = 1.8) +
  stat_qq_line(colour = PAL_H5[1], linewidth = 1) +
  labs(
    title    = "Fig 2a: Q-Q Plot of Model Residuals",
    subtitle = "Points should lie close to the diagonal line",
    x        = "Expected (theoretical) values",
    y        = "Actual residual values"
  )

p_hist <- ggplot(qq_df, aes(x = residuals)) +
  geom_histogram(bins = 20, fill = PAL_H5[2], alpha = 0.7, colour = "white") +
  geom_vline(xintercept = 0, colour = PAL_H5[1], linewidth = 1, linetype = "dashed") +
  labs(
    title    = "Fig 2b: Histogram of Model Residuals",
    subtitle = "Should be a bell shape centred at zero",
    x        = "Residual value",
    y        = "Count"
  )

# Stack vertically so neither plot is clipped
p_qq / p_hist
ggsave(file.path(output_dir, "H5_assumption_normality.png"),
       p_qq / p_hist, width = 7, height = 7, dpi = 150)
```

## Rule 3 - Equal Variance Across Groups (Levene's Test)

**What it means:** The spread of memory scores should be roughly the same
for the Active voice group and the Passive voice group.

**How we check:** Levene's test. A p-value above 0.05 means the spreads are
similar enough.

```{r assumption-levene}
levene_result <- leveneTest(
  corrected_memorability ~ Voice_Label,
  data   = ancova_df,
  center = mean
)

tibble(
  Test      = "Levene's Test (Active vs Passive groups)",
  F_stat    = round(levene_result$`F value`[1], 3),
  df1       = levene_result$Df[1],
  df2       = levene_result$Df[2],
  p_value   = round(levene_result$`Pr(>F)`[1], 4),
  Result    = ifelse(levene_result$`Pr(>F)`[1] > 0.05,
                     "OK (p > 0.05)", "Not OK (p <= 0.05)")
) %>%
  kable(
    caption   = "Table 4. Are the group variances roughly equal? (Levene's Test)",
    col.names = c("Test", "F", "df1", "df2", "p-value", "Conclusion"),
    booktabs  = TRUE
  ) %>%
  kable_styling(
    latex_options     = c("striped", "HOLD_position"),
    bootstrap_options = c("striped", "hover"),
    full_width        = FALSE,
    position          = "center"
  ) %>%
  row_spec(0, bold = TRUE, background = HDR_COLOR, color = "white")
```

## Rule 4 - Parallel Regression Slopes (Most Important Rule)

**What it means:** The practice scores should affect Active and Passive sentences
**in the same way**. If practice speed helps Active-sentence memory but hurts
Passive-sentence memory (or vice versa), ANCOVA becomes unreliable.

**How we check:** We add a "practice x voice" interaction term to the model. If
it is **not significant** (p > 0.05), the slopes are parallel and ANCOVA is valid.

```{r assumption-hors}
# Add interaction terms to test if slopes differ across Voice conditions
hors_fit <- lm(
  corrected_memorability ~
    Voice_Label * covariate_RT +
    Voice_Label * covariate_Error,
  data = ancova_df
)

hors_anova <- Anova(hors_fit, type = "III")

hors_anova %>%
  as.data.frame() %>%
  tibble::rownames_to_column("Term") %>%
  rename(
    SS        = `Sum Sq`,
    df        = Df,
    F_value   = `F value`,
    p_value   = `Pr(>F)`
  ) %>%
  mutate(
    F_value   = round(F_value, 3),
    p_value   = round(p_value, 4),
    Important = ifelse(!is.na(p_value) & p_value < 0.05,
                       "Significant (problem!)", "Not significant (good)")
  ) %>%
  kable(
    caption  = "Table 5. Parallel Slopes Test (interaction terms should NOT be significant)",
    booktabs = TRUE,
    digits   = 4
  ) %>%
  kable_styling(
    latex_options     = c("striped", "HOLD_position", "scale_down"),
    bootstrap_options = c("striped", "hover"),
    full_width        = FALSE,
    position          = "center"
  ) %>%
  row_spec(0, bold = TRUE, background = HDR_COLOR, color = "white")
```

```{r hors-visualise, fig.cap="Figure 3. Parallel slopes check. Top: Practice speed vs. memory by voice. Bottom: Practice errors vs. memory by voice. Parallel lines across blue (Active) and pink (Passive) mean the assumption is met.", fig.width=7, fig.height=7, out.width="100%"}
p_hors1 <- ggplot(ancova_df,
  aes(x = covariate_RT, y = corrected_memorability, colour = Voice_Label)) +
  geom_point(alpha = 0.5, size = 1.8) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 1.0) +
  scale_colour_manual(values = PAL_VOICE, name = "Voice") +
  labs(
    title    = "Fig 3a: Practice Speed vs. Memory (by Voice)",
    subtitle = "Parallel lines = slopes assumption met",
    x        = "Mean Practice Reaction Time (ms)",
    y        = "Corrected Memorability"
  )

p_hors2 <- ggplot(ancova_df,
  aes(x = covariate_Error, y = corrected_memorability, colour = Voice_Label)) +
  geom_point(alpha = 0.5, size = 1.8) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 1.0) +
  scale_colour_manual(values = PAL_VOICE, name = "Voice") +
  labs(
    title    = "Fig 3b: Practice Errors vs. Memory (by Voice)",
    subtitle = "Parallel lines = slopes assumption met",
    x        = "Practice Error Rate",
    y        = "Corrected Memorability"
  )

# Stack vertically so neither plot is clipped
p_hors1 / p_hors2
ggsave(file.path(output_dir, "H5_assumption_hors.png"),
       p_hors1 / p_hors2, width = 7, height = 7, dpi = 150)
```

> **Reading this plot:** We want the blue and pink lines to run **roughly parallel**.
> If one line tilts up while the other tilts down, the assumption is broken.

---

# Step D: Running the ANCOVA

## The Main Model

Now we run the actual ANCOVA --- voice condition plus both practice covariates
together predicting memory scores.

```{r ancova-model}
# Main ANCOVA model
ancova_fit <- lm(
  corrected_memorability ~ Voice_Label + covariate_RT + covariate_Error,
  data = ancova_df
)

# Get the full ANOVA table (Type III controls for overlap between predictors)
ancova_anova <- Anova(ancova_fit, type = "III")
```

## Effect Sizes

We calculate **partial eta-squared** ($\eta^2_p$) for each predictor. This tells
us what percentage of the remaining variation each predictor explains.

- Below 0.01 = negligible effect
- 0.01 to 0.06 = small effect
- 0.06 to 0.14 = medium effect
- Above 0.14 = large effect

```{r ancova-eta-squared}
ss_error <- ancova_anova["Residuals", "Sum Sq"]

ancova_table <- ancova_anova %>%
  as.data.frame() %>%
  tibble::rownames_to_column("Term") %>%
  filter(Term != "(Intercept)") %>%
  rename(
    SS      = `Sum Sq`,
    df      = Df,
    F_value = `F value`,
    p_value = `Pr(>F)`
  ) %>%
  mutate(
    Partial_eta2 = ifelse(
      Term != "Residuals",
      round(SS / (SS + ss_error), 4),
      NA_real_
    ),
    F_value     = round(F_value, 3),
    p_value     = round(p_value, 4),
    Significant = case_when(
      is.na(p_value)   ~ "N/A",
      p_value < 0.001  ~ "Yes (p < 0.001)",
      p_value < 0.01   ~ "Yes (p < 0.01)",
      p_value < 0.05   ~ "Yes (p < 0.05)",
      TRUE             ~ "No"
    ),
    Effect_Size = case_when(
      is.na(Partial_eta2)    ~ "N/A",
      Partial_eta2 >= 0.14   ~ "Large",
      Partial_eta2 >= 0.06   ~ "Medium",
      Partial_eta2 >= 0.01   ~ "Small",
      TRUE                   ~ "Negligible"
    )
  )

ancova_table %>%
  kable(
    caption  = "Table 6. ANCOVA Results: What predicts memory scores?",
    booktabs = TRUE,
    digits   = 4,
    col.names = c("Predictor", "SS", "df", "F", "p-value",
                  "Partial eta-sq", "Significant?", "Effect Size")
  ) %>%
  kable_styling(
    latex_options     = c("striped", "HOLD_position", "scale_down"),
    bootstrap_options = c("striped", "hover"),
    full_width        = FALSE,
    position          = "center"
  ) %>%
  row_spec(0, bold = TRUE, background = HDR_COLOR, color = "white")
```

## Regression Coefficients

This table shows the direction and size of each predictor's effect.
A **negative coefficient** for Practice RT means: slower practice = lower memory score.
A **positive coefficient** would mean the opposite.

```{r ancova-coefficients}
tidy(ancova_fit) %>%
  mutate(
    estimate  = round(estimate, 5),
    std.error = round(std.error, 5),
    statistic = round(statistic, 3),
    p.value   = round(p.value, 4),
    Direction = case_when(
      estimate > 0 ~ "Positive",
      estimate < 0 ~ "Negative",
      TRUE         ~ "Zero"
    ),
    Significant = ifelse(p.value < 0.05, "Yes", "No")
  ) %>%
  rename(
    Predictor = term,
    Estimate  = estimate,
    Std_Error = std.error,
    t_value   = statistic,
    p_value   = p.value
  ) %>%
  kable(
    caption  = "Table 7. How much does each predictor change the memory score?",
    booktabs = TRUE
  ) %>%
  kable_styling(
    latex_options     = c("striped", "HOLD_position", "scale_down"),
    bootstrap_options = c("striped", "hover"),
    full_width        = FALSE,
    position          = "center"
  ) %>%
  row_spec(0, bold = TRUE, background = HDR_COLOR, color = "white")
```

## Overall Model Quality

```{r ancova-fit-summary}
f_stat    <- summary(ancova_fit)$fstatistic
overall_p <- pf(f_stat[1], f_stat[2], f_stat[3], lower.tail = FALSE)

tibble(
  Metric  = c("R-squared (how much variance explained)",
              "Adjusted R-squared",
              "F-statistic",
              "Overall p-value"),
  Value   = c(
    round(summary(ancova_fit)$r.squared, 4),
    round(summary(ancova_fit)$adj.r.squared, 4),
    round(f_stat[1], 3),
    round(overall_p, 4)
  )
) %>%
  kable(
    caption  = "Table 8. How well does the overall ANCOVA model fit the data?",
    booktabs = TRUE
  ) %>%
  kable_styling(
    latex_options     = c("striped", "HOLD_position"),
    bootstrap_options = c("striped", "hover"),
    full_width        = FALSE,
    position          = "center"
  ) %>%
  row_spec(0, bold = TRUE, background = HDR_COLOR, color = "white")

cat(sprintf("\nR-squared      : %.4f\n", summary(ancova_fit)$r.squared))
cat(sprintf("Adjusted R-sq  : %.4f\n", summary(ancova_fit)$adj.r.squared))
cat(sprintf("Overall p-value: %.4f\n", overall_p))

if (overall_p < 0.05) {
  cat("\nDECISION: REJECT H0 --- The overall ANCOVA model is significant.\n")
} else {
  cat("\nDECISION: KEEP H0 --- The overall ANCOVA model is NOT significant.\n")
}
```

---

# Visualising the Results

## Adjusted vs Raw Memory Scores by Voice

This plot compares memory scores **before** and **after** we control for practice
performance. If the adjusted bars look different from the raw bars, practice
proficiency was affecting our results.

```{r adjusted-means, fig.cap="Figure 4. Memory scores by voice condition. Purple bars = adjusted scores after removing the practice effect. Blue bars = raw unadjusted scores. If they differ, practice was influencing the results.", fig.width=7, fig.height=4, out.width="100%"}
mean_rt    <- mean(ancova_df$covariate_RT,    na.rm = TRUE)
mean_error <- mean(ancova_df$covariate_Error, na.rm = TRUE)

adj_means <- tibble(
  Voice_Label     = c("Active", "Passive"),
  covariate_RT    = mean_rt,
  covariate_Error = mean_error
) %>%
  mutate(adj_CM = predict(ancova_fit, newdata = .))

raw_means <- ancova_df %>%
  group_by(Voice_Label) %>%
  summarise(raw_CM = mean(corrected_memorability, na.rm = TRUE), .groups = "drop")

comp_df <- adj_means %>%
  left_join(raw_means, by = "Voice_Label") %>%
  pivot_longer(
    cols      = c(adj_CM, raw_CM),
    names_to  = "Type",
    values_to = "CM"
  ) %>%
  mutate(
    Type = case_when(
      Type == "adj_CM" ~ "Adjusted (ANCOVA)",
      Type == "raw_CM" ~ "Unadjusted (raw mean)",
      TRUE             ~ Type
    )
  )

ggplot(comp_df, aes(x = Voice_Label, y = CM, fill = Type)) +
  geom_col(position = position_dodge(0.65), width = 0.55, alpha = 0.85) +
  scale_fill_manual(values = c("Adjusted (ANCOVA)"     = PAL_H5[1],
                               "Unadjusted (raw mean)" = PAL_H5[2])) +
  labs(
    title    = "Adjusted vs. Raw Memory Scores by Voice Condition",
    subtitle = "Adjusted = after removing practice effect | Raw = before adjustment",
    x        = "Voice Condition",
    y        = "Corrected Memorability (CM)",
    fill     = "Score Type"
  ) +
  theme(legend.position = "right")

ggsave(file.path(output_dir, "H5_adjusted_vs_raw_means.png"),
       last_plot(), width = 7, height = 4, dpi = 150)
```

## Practice Speed and Memory Scores

```{r scatter-rt, fig.cap="Figure 5. Each dot is one participant. Does being faster in practice lead to higher memory scores? Blue = Active sentences. Pink = Passive sentences.", fig.width=7, fig.height=4, out.width="100%"}
ggplot(ancova_df, aes(x = covariate_RT, y = corrected_memorability,
                      colour = Voice_Label)) +
  geom_point(alpha = 0.6, size = 2.8) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 1.1) +
  scale_colour_manual(values = PAL_VOICE, name = "Voice") +
  labs(
    title    = "Practice Speed vs. Final Memory Score",
    subtitle = "Does being faster in practice help you remember better?",
    x        = "Average Practice Reaction Time (ms) --- lower = faster",
    y        = "Corrected Memorability (CM)"
  )

ggsave(file.path(output_dir, "H5_scatter_RT.png"),
       last_plot(), width = 7, height = 4, dpi = 150)
```

## Practice Errors and Memory Scores

```{r scatter-error, fig.cap="Figure 6. Each dot is one participant. Does making more errors in practice lead to lower memory scores? Blue = Active sentences. Pink = Passive sentences.", fig.width=7, fig.height=4, out.width="100%"}
ggplot(ancova_df, aes(x = covariate_Error, y = corrected_memorability,
                      colour = Voice_Label)) +
  geom_point(alpha = 0.6, size = 2.8) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 1.1) +
  scale_colour_manual(values = PAL_VOICE, name = "Voice") +
  labs(
    title    = "Practice Errors vs. Final Memory Score",
    subtitle = "Does making more mistakes in practice hurt your memory later?",
    x        = "Practice Error Rate (0 = no errors, 1 = all wrong)",
    y        = "Corrected Memorability (CM)"
  )

ggsave(file.path(output_dir, "H5_scatter_Error.png"),
       last_plot(), width = 7, height = 4, dpi = 150)
```

---

# Collinearity Check

Are the two practice scores highly correlated with each other? If they are,
they are measuring the same thing and we only need to keep one.

```{r collinearity-check}
r_cov <- cor(ancova_df$covariate_RT, ancova_df$covariate_Error,
             use = "complete.obs")

cat(sprintf("Correlation between Practice Speed and Practice Errors: r = %.4f\n", r_cov))

if (abs(r_cov) > 0.8) {
  cat("WARNING: The two scores are very similar (|r| > 0.8). Consider using only one.\n")
} else {
  cat("The two scores are different enough to use both together. No problem.\n")
}
```

---

# Final Decision

```{r h5-decision, echo=FALSE, results='asis'}
p_rt  <- ancova_table %>% filter(Term == "covariate_RT")    %>% pull(p_value)
p_err <- ancova_table %>% filter(Term == "covariate_Error") %>% pull(p_value)
covariates_significant <- (p_rt < 0.05) | (p_err < 0.05)

cat("## Was H5 Supported?\n\n")

if (covariates_significant) {
  cat(paste0(
    "**Decision: REJECT the Null Hypothesis (H0)**\n\n",
    "At least one practice score significantly predicted how well participants ",
    "remembered sentences. This means **practice performance does carry over** ",
    "into the real experiment.\n\n",
    "In simple terms: participants who did better in the warm-up phase also tended ",
    "to remember more sentences in the actual experiment.\n\n",
    "**What this means for future research:** When running experiments like this, ",
    "researchers should always include practice phase scores as a covariate. ",
    "This makes the statistical models more accurate by removing background noise ",
    "caused by individual differences in how quickly people pick up new tasks."
  ))
} else {
  cat(paste0(
    "**Decision: KEEP the Null Hypothesis (H0)**\n\n",
    "Neither practice score significantly predicted final memory performance. ",
    "This means that how well someone did in warm-up did NOT consistently predict ",
    "how they did in the real experiment.\n\n",
    "In simple terms: some people were slow in practice but did fine in the real ",
    "task, and vice versa. The warm-up did not measure something that sticks.\n\n",
    "**What this means:** In this dataset, the main experiment results are driven ",
    "by the actual experimental conditions (noun type, grammatical voice), not by ",
    "how fast people learned the button presses. The H1-H4 findings stand on their own."
  ))
}
```

---

# Data Flow Summary and Conclusion

This section shows the complete journey of the data for Hypothesis 5, from raw
log files all the way to the final conclusion.

```
STEP 1: Raw Data (114 participant log files)
        |
        | Filter: keep ONLY practice-phase rows
        v
STEP 2: Practice Data Extracted
        - Events labelled "Practice" kept
        - All main-experiment rows ignored here
        |
        | Calculate per participant:
        |   (a) Average reaction time in practice  --> covariate_RT
        |   (b) Error rate in practice             --> covariate_Error
        v
STEP 3: Practice Scores Ready (one row per participant)
        |
        | Load main experiment memory scores from H3/H4 output
        | (H3H4_step3_analysis.csv)
        |   --> Corrected Memorability per participant x Voice condition
        |
        | Merge by participant_ID
        v
STEP 4: Combined Dataset
        - Columns: participant_ID, Voice_Label,
                   corrected_memorability, covariate_RT, covariate_Error
        |
        | Check 4 ANCOVA assumptions:
        |   (1) Linearity          -- visual check with scatter + LOESS
        |   (2) Normal residuals   -- Shapiro-Wilk test
        |   (3) Equal variances    -- Levene's test
        |   (4) Parallel slopes    -- interaction model test
        v
STEP 5: Assumptions Checked
        |
        | Fit ANCOVA model:
        |   Memory = Voice + Practice_RT + Practice_Error
        |
        | Extract: F-statistics, p-values, partial eta-squared
        v
STEP 6: ANCOVA Results
        |
        | Compare: adjusted means vs raw means
        | Visualise: scatter plots for each covariate
        v
STEP 7: FINAL CONCLUSION
        If Practice RT or Practice Errors are significant (p < 0.05):
            --> H0 rejected: practice performance predicts memory
            --> Future models should include these as covariates
        Else:
            --> H0 retained: practice performance does NOT predict memory
            --> Main experiment effects are independent of practice performance
```

## Summary Table

```{r summary-table, echo=FALSE}
tibble(
  Step   = c(
    "Data Source",
    "Practice Covariates",
    "Main Outcome (DV)",
    "Main Group Variable (IV)",
    "Statistical Test",
    "Assumption 1: Linearity",
    "Assumption 2: Normal Residuals",
    "Assumption 3: Equal Variance",
    "Assumption 4: Parallel Slopes",
    "Practice RT significant?",
    "Practice Errors significant?",
    "Overall Model significant?",
    "Final Decision"
  ),
  Result = c(
    "114 log files from Sentence Memorability experiment",
    "covariate_RT (practice speed), covariate_Error (practice errors)",
    "Corrected Memorability = Hit Rate minus False Alarm Rate",
    "Grammatical Voice (Active or Passive)",
    "ANCOVA with Type III sums of squares and partial eta-squared",
    "Checked visually with scatter plot and LOESS smoother",
    ifelse(sw_test$p.value > 0.05,
           paste0("Met (Shapiro-Wilk p = ", round(sw_test$p.value, 3), ")"),
           paste0("Violated (Shapiro-Wilk p = ", round(sw_test$p.value, 3), ")")),
    ifelse(levene_result$`Pr(>F)`[1] > 0.05,
           paste0("Met (Levene p = ", round(levene_result$`Pr(>F)`[1], 3), ")"),
           paste0("Violated (Levene p = ", round(levene_result$`Pr(>F)`[1], 3), ")")),
    ifelse(all(ancova_table %>%
      filter(grepl(":", Term)) %>%
      pull(p_value) > 0.05, na.rm = TRUE),
      "Met -- slopes are parallel across Voice conditions",
      "Violated -- slopes differ across Voice conditions"),
    ifelse(p_rt < 0.05,
           paste0("Yes (p = ", round(p_rt, 3), ")"),
           paste0("No (p = ", round(p_rt, 3), ")")),
    ifelse(p_err < 0.05,
           paste0("Yes (p = ", round(p_err, 3), ")"),
           paste0("No (p = ", round(p_err, 3), ")")),
    ifelse(overall_p < 0.05,
           paste0("Yes (p = ", round(overall_p, 3), ")"),
           paste0("No (p = ", round(overall_p, 3), ")")),
    ifelse(covariates_significant,
      "REJECT H0: Practice performance predicts memory scores",
      "KEEP H0: Practice performance does NOT predict memory scores")
  )
) %>%
  kable(
    caption   = "Table 9. Complete H5 Summary -- from data to decision",
    col.names = c("What We Checked", "Result"),
    booktabs  = TRUE
  ) %>%
  kable_styling(
    latex_options     = c("striped", "HOLD_position"),
    bootstrap_options = c("striped", "hover"),
    full_width        = FALSE,
    position          = "center"
  ) %>%
  column_spec(1, bold = TRUE, width = "6cm") %>%
  column_spec(2, width = "9cm") %>%
  row_spec(0, bold = TRUE, background = HDR_COLOR, color = "white") %>%
  row_spec(13, bold = TRUE, background = "#EBF5FB")
```
