---
title: "Hypothesis 6: Does Fatigue Slow People Down Over Time?"
subtitle: "Checking if Reaction Time and Accuracy Change Across the Experiment"
author: "Team: Metamatics"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    latex_engine: xelatex
    number_sections: true
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    code_folding: show
    fig_caption: true
    highlight: tango
    df_print: paged
fontsize: 11pt
geometry: margin=1in
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{amsmath}
  - \usepackage{amssymb}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo       = TRUE,
  warning    = FALSE,
  message    = FALSE,
  fig.align  = "center",
  fig.width  = 7,
  fig.height = 4,
  out.width  = "100%",
  dpi        = 150
)

pkgs <- c(
  "dplyr", "tidyr", "tibble", "ggplot2", "readr", "stringr",
  "purrr", "knitr", "kableExtra", "scales", "patchwork"
)
for (p in pkgs) {
  if (!requireNamespace(p, quietly = TRUE))
    install.packages(p, repos = "https://cran.r-project.org")
  library(p, character.only = TRUE)
}

PAL_BLOCKS <- c("1: Early" = "#2E86AB", "2: Middle" = "#F18F01", "3: Late" = "#C73E1D")
HDR_COLOR  <- "#1B3A57"

theme_set(
  theme_minimal(base_size = 13) +
    theme(
      plot.title       = element_text(face = "bold", size = 14, colour = "#2C2C2C"),
      plot.subtitle    = element_text(colour = "grey45", size = 11),
      legend.position  = "bottom",
      panel.grid.minor = element_blank()
    )
)

log_dir    <- "Sentence Memorability/NewLogsAnonymized"
output_dir <- "R_output"
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

# ── Pre-initialise test-result variables so the summary table never crashes ──
# These are overwritten by the actual test chunks below.
chi_rt       <- NA_real_;  df_rt  <- NA_real_;  p_rt  <- NA_real_
chi_acc      <- NA_real_;  df_acc <- NA_real_;  p_acc <- NA_real_
test_name_rt  <- "Not yet run"
test_name_acc <- "Not yet run"
```

---

# What is Hypothesis 6?

## The Simple Idea

Think about any long task you have done — like sitting through a 3-hour exam.
By the end, your brain is tired. You might take longer to answer questions or
make more mistakes than at the start. This is called **cognitive fatigue**.

In our experiment, each participant sat through roughly **30 minutes** of
reading and recognising sentences — lots of trials in total.

> **Do people get slower or less accurate as the experiment goes on?**

If yes, then fatigue is a real problem in this experiment. It means the results
from H1 to H5 could be slightly biased — the differences we see might partly
be caused by tiredness, not just the sentence type or voice.

## How We Test It

We split each participant's experiment into **3 equal time blocks**:

| Block | When | What it represents |
|---|---|---|
| Block 1: Early | First third of trials | Fresh, not yet tired |
| Block 2: Middle | Middle third of trials | Settling in |
| Block 3: Late | Last third of trials | Potentially fatigued |

We then check **two things**:

1. **Reaction Time (RT)** — Does it take longer to press the key later in the experiment?
2. **Accuracy** — Do people make more mistakes later in the experiment?

---

# Formal Hypothesis

## In Plain Words

> **Null Hypothesis (H0):** Reaction time and accuracy do NOT change significantly
> across the three time blocks. People perform the same throughout the experiment.

> **Alternative Hypothesis (H6):** Reaction time increases and/or accuracy decreases
> significantly from Early to Late blocks, showing that cognitive fatigue builds up.

## The Tests We Use

| Outcome | Test (if non-normal) | Why |
|---|---|---|
| Reaction Time (RT) | Friedman Test | RT is right-skewed; non-parametric test is safer |
| Accuracy | Friedman Test | Accuracy is bounded 0–1; often non-normal |

We check normality first with **Shapiro-Wilk** and **Q-Q plots** before deciding.

---

# Step 0: Participant Validation

## Why Do We Need This?

Not every participant should be in the analysis. Some people may have clicked
randomly, zoned out, or not understood the task. We check this using special
**validation sentences** — easy, obviously repeated sentences that anyone paying
attention should recognise.

## The Validation Formula

$$\text{Correct Hits} > \frac{\text{Wrong IR Presses}}{2} + \text{Missed IRs}$$

Where:

- **Correct Hits** = validation repeats where the person pressed spacebar ✅
- **Wrong IR Presses** = non-repeats where they pressed spacebar ❌
- **Missed IRs** = validation repeats where they did NOT press spacebar ❌

In simple words: **your correct responses must clearly outnumber your mistakes.**

```{r load-all-for-validation}
log_files <- list.files(log_dir, pattern = "\\.log$", full.names = TRUE)
cat(sprintf("Found %d log files.\n", length(log_files)))

raw_all <- log_files %>%
  map_dfr(~ {
    tryCatch(
      read_csv(.x, show_col_types = FALSE, col_types = cols(.default = "c")),
      error = function(e) tibble()
    )
  }) %>%
  mutate(
    participant_ID   = as.integer(participant_ID),
    Timestamp        = suppressWarnings(as.numeric(Timestamp)),
    `Accuracy IR`    = suppressWarnings(as.numeric(`Accuracy IR`)),
    Reaction_time_IR = suppressWarnings(as.numeric(Reaction_time_IR)),
    isRepeat         = coalesce(tolower(trimws(isRepeat))     == "true", FALSE),
    isValidation     = coalesce(tolower(trimws(isValidation)) == "true", FALSE)
  ) %>%
  arrange(participant_ID, Timestamp)

cat(sprintf("Total rows loaded: %s across %d participants.\n",
            format(nrow(raw_all), big.mark = ","),
            n_distinct(raw_all$participant_ID)))
```

```{r participant-validation}
# Sort by time, remove practice rows, assign trial IDs
raw_trials <- raw_all %>%
  filter(!str_starts(Event, "Practice") | is.na(Event)) %>%
  arrange(participant_ID, Timestamp) %>%
  group_by(participant_ID) %>%
  mutate(trial_id = cumsum(Event == "Sentence shown")) %>%
  ungroup() %>%
  filter(trial_id > 0)

# One row per trial — from "Sentence shown" event
sentence_meta <- raw_trials %>%
  filter(Event == "Sentence shown") %>%
  group_by(participant_ID, trial_id) %>%
  slice(1) %>%
  ungroup() %>%
  select(participant_ID, trial_id, isRepeat, isValidation)

# Check BOTH the isValidation column AND the Event string
# (matches H1-H4 validation pipeline exactly)
has_validation_flag <- raw_trials %>%
  group_by(participant_ID, trial_id) %>%
  summarise(
    has_validation = any(str_detect(Event, "Validation"), na.rm = TRUE),
    ir_pressed     = any(str_detect(Event, "IR pressed"), na.rm = TRUE),
    .groups        = "drop"
  )

all_trials <- sentence_meta %>%
  inner_join(has_validation_flag, by = c("participant_ID", "trial_id")) %>%
  mutate(is_validation = isValidation | has_validation)

val_trials <- all_trials %>% filter(is_validation == TRUE)
cat(sprintf("Total validation trials found: %d\n", nrow(val_trials)))

# Apply the formula: Correct Hits > (False Alarms / 2) + Missed
validation_summary <- val_trials %>%
  group_by(participant_ID) %>%
  summarise(
    Correct_Hits = sum(isRepeat == TRUE  & ir_pressed == TRUE,  na.rm = TRUE),
    False_Alarms = sum(isRepeat == FALSE & ir_pressed == TRUE,  na.rm = TRUE),
    Missed       = sum(isRepeat == TRUE  & ir_pressed == FALSE, na.rm = TRUE),
    .groups      = "drop"
  ) %>%
  mutate(
    Passes = Correct_Hits > (False_Alarms / 2) + Missed,
    Status = ifelse(Passes, "Kept", "Excluded")
  )

valid_ids    <- validation_summary %>% filter( Passes) %>% pull(participant_ID)
excluded_ids <- validation_summary %>% filter(!Passes) %>% pull(participant_ID)

cat(sprintf("Total participants:    %d\n", nrow(validation_summary)))
cat(sprintf("Participants KEPT:     %d\n", length(valid_ids)))
cat(sprintf("Participants EXCLUDED: %d\n", length(excluded_ids)))
if (length(excluded_ids) > 0)
  cat(sprintf("Excluded IDs: %s\n", paste(excluded_ids, collapse = ", ")))
```

## Validation Results Table

Green = kept, Red = excluded.

```{r validation-table}
validation_summary %>%
  select(participant_ID, Correct_Hits, False_Alarms, Missed, Status) %>%
  kable(
    caption   = "Table 0. Participant Validation Results",
    col.names = c("Participant", "Correct Hits", "Wrong Presses", "Missed", "Status"),
    booktabs  = TRUE
  ) %>%
  kable_styling(
    latex_options     = c("striped", "HOLD_position", "scale_down"),
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width        = FALSE, position = "center"
  ) %>%
  column_spec(5,
    bold       = TRUE,
    color      = ifelse(validation_summary$Passes, "#1B5E20", "#B71C1C"),
    background = ifelse(validation_summary$Passes, "#E8F5E9", "#FDECEA")
  ) %>%
  row_spec(0, bold = TRUE, background = HDR_COLOR, color = "white")

write_csv(validation_summary, file.path(output_dir, "H6_validation.csv"))
```

---

# Step A: Loading Fatigue Data

```{r data-prep}
fatigue_raw <- log_files %>%
  map_dfr(~ {
    tryCatch({
      d <- read_csv(.x, show_col_types = FALSE, col_types = cols(.default = "c"))
      d %>%
        filter(!str_detect(Event, "Practice"), str_detect(Event, "IR pressed"))
    }, error = function(e) tibble())
  }) %>%
  mutate(
    participant_ID   = as.integer(participant_ID),
    Timestamp        = suppressWarnings(as.numeric(Timestamp)),
    Reaction_time_IR = suppressWarnings(as.numeric(Reaction_time_IR)),
    `Accuracy IR`    = suppressWarnings(as.numeric(`Accuracy IR`))
  ) %>%
  filter(!is.na(Reaction_time_IR), participant_ID %in% valid_ids)

cat(sprintf("IR press events (valid participants): %s\n",
            format(nrow(fatigue_raw), big.mark = ",")))
cat(sprintf("Valid participants in dataset:        %d\n",
            n_distinct(fatigue_raw$participant_ID)))
```

## Splitting into 3 Time Blocks

```{r block-segmentation}
# Sort by Timestamp first — so row 1 = FIRST event in real time
fatigue_blocks <- fatigue_raw %>%
  arrange(participant_ID, Timestamp) %>%
  group_by(participant_ID) %>%
  mutate(
    trial_sequence = row_number(),
    total_trials   = n(),
    block_id       = ntile(trial_sequence, 3)
  ) %>%
  mutate(
    Block_Name = case_when(
      block_id == 1 ~ "1: Early",
      block_id == 2 ~ "2: Middle",
      block_id == 3 ~ "3: Late"
    )
  ) %>%
  ungroup()

cat(sprintf("Split %s events into 3 blocks across %d participants.\n",
            format(nrow(fatigue_blocks), big.mark = ","),
            n_distinct(fatigue_blocks$participant_ID)))

min_trials <- fatigue_blocks %>%
  group_by(participant_ID) %>%
  summarise(n = n(), .groups = "drop") %>%
  pull(n) %>% min()
cat(sprintf("Minimum IR trials per participant: %d\n", min_trials))
```

## Preview Table 1 — Raw Rows with Block Assignment

```{r preview-raw-rows}
sample_ids <- fatigue_blocks %>%
  distinct(participant_ID) %>% slice(1:3) %>% pull(participant_ID)

fatigue_blocks %>%
  filter(participant_ID %in% sample_ids) %>%
  mutate(Timestamp_s = round(Timestamp / 1000, 1)) %>%
  select(
    Participant    = participant_ID,
    `Trial #`      = trial_sequence,
    `Total Trials` = total_trials,
    `Block ID`     = block_id,
    `Block Name`   = Block_Name,
    `RT (ms)`      = Reaction_time_IR,
    `Timestamp (s)` = Timestamp_s
  ) %>%
  kable(
    caption  = "Preview Table 1. Sample rows — trial number, block, RT and Timestamp.",
    booktabs = TRUE
  ) %>%
  kable_styling(
    latex_options     = c("striped", "HOLD_position", "scale_down"),
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE, position = "center", font_size = 10
  ) %>%
  row_spec(0, bold = TRUE, background = HDR_COLOR, color = "white")
```

## Preview Table 2 — Block Boundaries Per Participant

```{r preview-block-boundaries}
block_boundaries <- fatigue_blocks %>%
  group_by(participant_ID, Block_Name) %>%
  summarise(
    First_Trial       = min(trial_sequence),
    Last_Trial        = max(trial_sequence),
    N_Trials          = n(),
    First_Timestamp_s = round(min(Timestamp, na.rm = TRUE) / 1000, 0),
    Last_Timestamp_s  = round(max(Timestamp, na.rm = TRUE) / 1000, 0),
    .groups           = "drop"
  ) %>%
  arrange(participant_ID, Block_Name)

block_boundaries %>%
  filter(participant_ID %in% sample_ids) %>%
  kable(
    caption   = "Preview Table 2. Block boundaries for 3 sample participants.",
    col.names = c("Participant", "Block", "First Trial#", "Last Trial#",
                  "N Trials", "Block Start (s)", "Block End (s)"),
    booktabs  = TRUE
  ) %>%
  kable_styling(
    latex_options     = c("striped", "HOLD_position"),
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE, position = "center"
  ) %>%
  row_spec(0, bold = TRUE, background = HDR_COLOR, color = "white") %>%
  pack_rows("Participant 1", 1, 3) %>%
  pack_rows("Participant 2", 4, 6) %>%
  pack_rows("Participant 3", 7, 9)
```

## Preview Table 3 — Full Block Size Summary

```{r preview-block-sizes}
block_sizes <- fatigue_blocks %>%
  group_by(participant_ID, Block_Name) %>%
  summarise(N_Trials = n(), .groups = "drop") %>%
  pivot_wider(names_from = Block_Name, values_from = N_Trials) %>%
  mutate(
    Total    = `1: Early` + `2: Middle` + `3: Late`,
    Max_N    = pmax(`1: Early`, `2: Middle`, `3: Late`),
    Min_N    = pmin(`1: Early`, `2: Middle`, `3: Late`),
    Balanced = ifelse(Max_N - Min_N <= 1, "Yes", "No")
  ) %>%
  select(-Max_N, -Min_N)

block_sizes %>%
  kable(
    caption   = "Preview Table 3. Trials per block per participant.",
    col.names = c("Participant", "Early", "Middle", "Late", "Total", "Balanced?"),
    booktabs  = TRUE
  ) %>%
  kable_styling(
    latex_options     = c("striped", "HOLD_position", "scale_down"),
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE, position = "center", font_size = 10
  ) %>%
  row_spec(0, bold = TRUE, background = HDR_COLOR, color = "white") %>%
  column_spec(6,
    bold       = TRUE,
    color      = ifelse(block_sizes$Balanced == "Yes", "#1B5E20", "#B71C1C"),
    background = ifelse(block_sizes$Balanced == "Yes", "#E8F5E9", "#FDECEA")
  )

cat(sprintf("\nAll participants balanced: %s\n",
            ifelse(all(block_sizes$Balanced == "Yes"), "YES", "NO — check data!")))
```

## Computing One Score Per Person Per Block

We use **median RT** (not mean) because it is less affected by the occasional
very slow response. We use **mean accuracy** within each block.

```{r compute-block-scores}
person_blocks <- fatigue_blocks %>%
  group_by(participant_ID, Block_Name) %>%
  summarise(
    Median_RT = median(Reaction_time_IR, na.rm = TRUE),
    Mean_Acc  = mean(`Accuracy IR`, na.rm = TRUE),
    n_trials  = n(),
    .groups   = "drop"
  ) %>%
  mutate(Block_Name = factor(Block_Name,
                             levels = c("1: Early", "2: Middle", "3: Late")))

write_csv(person_blocks, file.path(output_dir, "H6_block_scores.csv"))
cat(sprintf("Block scores: %d rows (%d participants x 3 blocks)\n",
            nrow(person_blocks), n_distinct(person_blocks$participant_ID)))
```

---

# Step B: Descriptive Statistics

## Table 1 — Reaction Time by Block

```{r rt-descriptives}
rt_desc <- person_blocks %>%
  group_by(Block_Name) %>%
  summarise(
    N         = n(),
    Mean_RT   = round(mean(Median_RT,   na.rm = TRUE), 1),
    Median_RT = round(median(Median_RT, na.rm = TRUE), 1),
    SD_RT     = round(sd(Median_RT,     na.rm = TRUE), 1),
    Min_RT    = round(min(Median_RT,    na.rm = TRUE), 1),
    Max_RT    = round(max(Median_RT,    na.rm = TRUE), 1),
    .groups   = "drop"
  )

rt_desc %>%
  kable(
    caption   = "Table 1. Reaction Time (ms) by Time Block",
    col.names = c("Block", "N", "Mean (ms)", "Median (ms)", "SD", "Min", "Max"),
    booktabs  = TRUE
  ) %>%
  kable_styling(
    latex_options     = c("striped", "HOLD_position"),
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE, position = "center"
  ) %>%
  row_spec(0, bold = TRUE, background = HDR_COLOR, color = "white")
```

## Table 2 — Accuracy by Block

```{r acc-descriptives}
acc_desc <- person_blocks %>%
  group_by(Block_Name) %>%
  summarise(
    N        = n(),
    Mean_Acc = round(mean(Mean_Acc,   na.rm = TRUE), 4),
    SD_Acc   = round(sd(Mean_Acc,     na.rm = TRUE), 4),
    Min_Acc  = round(min(Mean_Acc,    na.rm = TRUE), 4),
    Max_Acc  = round(max(Mean_Acc,    na.rm = TRUE), 4),
    .groups  = "drop"
  )

acc_desc %>%
  kable(
    caption   = "Table 2. Accuracy (proportion correct) by Time Block",
    col.names = c("Block", "N", "Mean Accuracy", "SD", "Min", "Max"),
    booktabs  = TRUE
  ) %>%
  kable_styling(
    latex_options     = c("striped", "HOLD_position"),
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE, position = "center"
  ) %>%
  row_spec(0, bold = TRUE, background = HDR_COLOR, color = "white")
```

---

# Step C: Visualising the Data

## Figure 1 — Reaction Time Distributions

```{r fig1-rt-violin, fig.cap="Figure 1. Reaction time violin + boxplot per block. Each dot = one participant. Y-axis starts at 0. If people get slower, boxes shift upward from Early to Late.", fig.width=7, fig.height=5, out.width="100%"}
p1 <- ggplot(person_blocks,
  aes(x = Block_Name, y = Median_RT, fill = Block_Name)) +
  geom_violin(alpha = 0.55, colour = "grey60", width = 0.8) +
  geom_boxplot(width = 0.18, alpha = 0.85, outlier.shape = NA, colour = "grey20") +
  geom_jitter(width = 0.07, alpha = 0.35, size = 1.5, colour = "grey30") +
  scale_fill_manual(values = PAL_BLOCKS, guide = "none") +
  scale_y_continuous(limits = c(0, 3200), oob = scales::squish,
                     labels = comma_format()) +
  labs(
    title    = "Figure 1: Reaction Time by Time Block",
    subtitle = "Y-axis starts at 0. If people get slower, boxes shift up.",
    x        = "Time Block",
    y        = "Median Reaction Time per Participant (ms)"
  )

p1
ggsave(file.path(output_dir, "H6_RT_violin.png"), p1, width = 7, height = 5, dpi = 150)
```

## Figure 2 — Accuracy Distributions

```{r fig2-acc-violin, fig.cap="Figure 2. Accuracy violin + boxplot per block. Each dot = one participant. If people make more mistakes later, boxes shift downward from Early to Late.", fig.width=7, fig.height=5, out.width="100%"}
p2 <- ggplot(person_blocks,
  aes(x = Block_Name, y = Mean_Acc, fill = Block_Name)) +
  geom_violin(alpha = 0.55, colour = "grey60", width = 0.8) +
  geom_boxplot(width = 0.18, alpha = 0.85, outlier.shape = NA, colour = "grey20") +
  geom_jitter(width = 0.07, alpha = 0.35, size = 1.5, colour = "grey30") +
  scale_fill_manual(values = PAL_BLOCKS, guide = "none") +
  scale_y_continuous(limits = c(0, 1), labels = percent_format()) +
  labs(
    title    = "Figure 2: Accuracy by Time Block",
    subtitle = "Does accuracy drop from Early to Late?",
    x        = "Time Block",
    y        = "Mean Accuracy per Participant"
  )

p2
ggsave(file.path(output_dir, "H6_Acc_violin.png"), p2, width = 7, height = 5, dpi = 150)
```

## Figure 3 — Density Plots (Shape of the Data)

Density plots show WHERE most participants' scores cluster. A curve shifted
to the right (for RT) means more participants were slow in that block.

```{r fig3-density, fig.cap="Figure 3. Density curves for RT (top) and Accuracy (bottom) per block. The peak shows where most participants scored. A rightward shift in RT = slower responses.", fig.width=7, fig.height=7, out.width="100%"}
p3a <- ggplot(person_blocks,
  aes(x = Median_RT, fill = Block_Name, colour = Block_Name)) +
  geom_density(alpha = 0.3, linewidth = 1) +
  scale_fill_manual(values   = PAL_BLOCKS, name = "Block") +
  scale_colour_manual(values = PAL_BLOCKS, name = "Block") +
  scale_x_continuous(limits = c(0, NA), labels = comma_format()) +
  labs(
    title    = "Fig 3a: RT Density by Block",
    subtitle = "Right shift = slower. Overlap = similar performance.",
    x        = "Median RT (ms)", y        = "Density"
  )

p3b <- ggplot(person_blocks,
  aes(x = Mean_Acc, fill = Block_Name, colour = Block_Name)) +
  geom_density(alpha = 0.3, linewidth = 1) +
  scale_fill_manual(values   = PAL_BLOCKS, name = "Block") +
  scale_colour_manual(values = PAL_BLOCKS, name = "Block") +
  scale_x_continuous(limits = c(0, 1), labels = percent_format()) +
  labs(
    title    = "Fig 3b: Accuracy Density by Block",
    subtitle = "Left shift = lower accuracy. Overlap = similar performance.",
    x        = "Mean Accuracy", y = "Density"
  )

p3a / p3b
ggsave(file.path(output_dir, "H6_density.png"),
       p3a / p3b, width = 7, height = 7, dpi = 150)
```

## Figure 4 — Average Trend Line

This plot shows the **group average** at each block with error bars.
A clear upward slope for RT or downward slope for accuracy = fatigue.

```{r fig4-trend, fig.cap="Figure 4. Average RT (top) and Accuracy (bottom) at each block. Error bars = +/- 1 standard error. A clear slope = performance changed across the experiment.", fig.width=7, fig.height=7, out.width="100%"}
trend_rt <- person_blocks %>%
  group_by(Block_Name) %>%
  summarise(Mean = mean(Median_RT, na.rm = TRUE),
            SE   = sd(Median_RT,   na.rm = TRUE) / sqrt(n()),
            .groups = "drop")

trend_acc <- person_blocks %>%
  group_by(Block_Name) %>%
  summarise(Mean = mean(Mean_Acc, na.rm = TRUE),
            SE   = sd(Mean_Acc,   na.rm = TRUE) / sqrt(n()),
            .groups = "drop")

p4a <- ggplot(trend_rt, aes(x = Block_Name, y = Mean, group = 1)) +
  geom_ribbon(aes(ymin = Mean - SE, ymax = Mean + SE), alpha = 0.15,
              fill = PAL_BLOCKS["1: Early"]) +
  geom_line(colour = PAL_BLOCKS["1: Early"], linewidth = 1.2) +
  geom_point(aes(colour = Block_Name), size = 4) +
  geom_errorbar(aes(ymin = Mean - SE, ymax = Mean + SE),
                width = 0.15, colour = "grey40", linewidth = 0.8) +
  scale_colour_manual(values = PAL_BLOCKS, guide = "none") +
  scale_y_continuous(limits = c(0, NA), labels = comma_format()) +
  labs(title    = "Fig 4a: Average RT Trend Across Blocks",
       subtitle = "Y starts at 0. Each point = group mean. Shaded band = standard error.",
       x = "Time Block", y = "Mean Median RT (ms)")

p4b <- ggplot(trend_acc, aes(x = Block_Name, y = Mean, group = 1)) +
  geom_ribbon(aes(ymin = Mean - SE, ymax = Mean + SE), alpha = 0.15,
              fill = PAL_BLOCKS["1: Early"]) +
  geom_line(colour = PAL_BLOCKS["1: Early"], linewidth = 1.2) +
  geom_point(aes(colour = Block_Name), size = 4) +
  geom_errorbar(aes(ymin = Mean - SE, ymax = Mean + SE),
                width = 0.15, colour = "grey40", linewidth = 0.8) +
  scale_colour_manual(values = PAL_BLOCKS, guide = "none") +
  scale_y_continuous(limits = c(0, 1), labels = percent_format()) +
  labs(title    = "Fig 4b: Average Accuracy Trend Across Blocks",
       subtitle = "Each point = group mean. Shaded band = standard error.",
       x = "Time Block", y = "Mean Accuracy")

p4a / p4b
ggsave(file.path(output_dir, "H6_trend_lines.png"),
       p4a / p4b, width = 7, height = 7, dpi = 150)
```

## Figure 5 — Individual Participant Trajectories (Spaghetti)

Each thin line = one participant. If **most** lines go up for RT or down for
accuracy, fatigue is widespread — not just a few outliers.

```{r fig5-spaghetti, fig.cap="Figure 5. Spaghetti plots. Each thin line = one participant. Thick red line = group average. Top = RT, Bottom = Accuracy.", fig.width=7, fig.height=7, out.width="100%"}
p5a <- ggplot(person_blocks,
  aes(x = Block_Name, y = Median_RT, group = participant_ID)) +
  geom_line(alpha = 0.25, colour = "#2E86AB", linewidth = 0.5) +
  stat_summary(aes(group = 1), fun = mean, geom = "line",
               colour = "#C73E1D", linewidth = 1.5) +
  stat_summary(aes(group = 1), fun = mean, geom = "point",
               colour = "#C73E1D", size = 3) +
  scale_y_continuous(limits = c(0, NA), labels = comma_format()) +
  labs(title    = "Fig 5a: Individual RT Trajectories",
       subtitle = "Y starts at 0. Thin = each person. Thick red = group average.",
       x = "Time Block", y = "Median RT (ms)")

p5b <- ggplot(person_blocks,
  aes(x = Block_Name, y = Mean_Acc, group = participant_ID)) +
  geom_line(alpha = 0.25, colour = "#F18F01", linewidth = 0.5) +
  stat_summary(aes(group = 1), fun = mean, geom = "line",
               colour = "#C73E1D", linewidth = 1.5) +
  stat_summary(aes(group = 1), fun = mean, geom = "point",
               colour = "#C73E1D", size = 3) +
  scale_y_continuous(limits = c(0, 1), labels = percent_format()) +
  labs(title    = "Fig 5b: Individual Accuracy Trajectories",
       subtitle = "Thin = each person. Thick red = group average.",
       x = "Time Block", y = "Mean Accuracy")

p5a / p5b
ggsave(file.path(output_dir, "H6_spaghetti.png"),
       p5a / p5b, width = 7, height = 7, dpi = 150)
```

## Figure 6 — Summary Boxplots

```{r fig6-summary-box, fig.cap="Figure 6. Clean summary boxplots. Red dot = group mean. Box = spread of individual medians. This is the clearest single visual for whether fatigue is present.", fig.width=7, fig.height=7, out.width="100%"}
p6a <- ggplot(person_blocks,
  aes(x = Block_Name, y = Median_RT, fill = Block_Name)) +
  geom_boxplot(width = 0.45, alpha = 0.75, outlier.alpha = 0.3) +
  stat_summary(fun = mean, geom = "point",
               shape = 21, size = 3, fill = "#C73E1D", colour = "white") +
  scale_fill_manual(values = PAL_BLOCKS, guide = "none") +
  scale_y_continuous(limits = c(0, NA), labels = comma_format()) +
  labs(title = "Fig 6a: RT by Block (Y from 0 | red dot = mean)",
       x = "Time Block", y = "Median RT (ms)")

p6b <- ggplot(person_blocks,
  aes(x = Block_Name, y = Mean_Acc, fill = Block_Name)) +
  geom_boxplot(width = 0.45, alpha = 0.75, outlier.alpha = 0.3) +
  stat_summary(fun = mean, geom = "point",
               shape = 21, size = 3, fill = "#C73E1D", colour = "white") +
  scale_fill_manual(values = PAL_BLOCKS, guide = "none") +
  scale_y_continuous(limits = c(0, 1), labels = percent_format()) +
  labs(title = "Fig 6b: Accuracy by Block (red dot = mean)",
       x = "Time Block", y = "Mean Accuracy")

p6a / p6b
ggsave(file.path(output_dir, "H6_summary_boxplots.png"),
       p6a / p6b, width = 7, height = 7, dpi = 150)
```

---

# Step D: Normality Check

Before picking the statistical test, we check if each block's data is normally
distributed. We use:

- **Shapiro-Wilk test** — gives a number (p < 0.05 = NOT normal)
- **Q-Q plot** — visual: dots on the line = normal, curving away = NOT normal

## Table 3 — Shapiro-Wilk Results

```{r normality-check}
blocks <- c("1: Early", "2: Middle", "3: Late")

sw_rt <- lapply(blocks, function(b) {
  vals   <- person_blocks %>% filter(Block_Name == b) %>% pull(Median_RT)
  result <- shapiro.test(vals)
  tibble(Outcome = "Reaction Time", Block = b,
         W       = round(as.numeric(result$statistic), 4),
         p_value = round(as.numeric(result$p.value),   4),
         Normal  = ifelse(result$p.value > 0.05,
                          "Yes (p > 0.05)", "No (p <= 0.05)"))
}) %>% bind_rows()

sw_acc <- lapply(blocks, function(b) {
  vals   <- person_blocks %>% filter(Block_Name == b) %>% pull(Mean_Acc)
  result <- shapiro.test(vals)
  tibble(Outcome = "Accuracy", Block = b,
         W       = round(as.numeric(result$statistic), 4),
         p_value = round(as.numeric(result$p.value),   4),
         Normal  = ifelse(result$p.value > 0.05,
                          "Yes (p > 0.05)", "No (p <= 0.05)"))
}) %>% bind_rows()

sw_all <- bind_rows(sw_rt, sw_acc)

sw_all %>%
  kable(
    caption   = "Table 3. Shapiro-Wilk Normality Test (per block, per outcome)",
    col.names = c("Outcome", "Block", "W statistic", "p-value", "Normal?"),
    booktabs  = TRUE
  ) %>%
  kable_styling(
    latex_options     = c("striped", "HOLD_position"),
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE, position = "center"
  ) %>%
  column_spec(5,
    bold       = TRUE,
    color      = ifelse(sw_all$Normal == "Yes (p > 0.05)", "#1B5E20", "#B71C1C"),
    background = ifelse(sw_all$Normal == "Yes (p > 0.05)", "#E8F5E9", "#FDECEA")
  ) %>%
  row_spec(0, bold = TRUE, background = HDR_COLOR, color = "white")

rt_normal_all  <- all(sw_rt$p_value  > 0.05)
acc_normal_all <- all(sw_acc$p_value > 0.05)
```

## Figure 7 — Q-Q Plots

```{r fig7-qq, fig.cap="Figure 7. Q-Q plots for all six distributions (RT and Accuracy per block). Points on the diagonal = normal. Curves away = NOT normal.", fig.width=7, fig.height=9, out.width="100%"}
qq_rt <- lapply(blocks, function(b) {
  d   <- person_blocks %>% filter(Block_Name == b)
  lbl <- sw_rt$Normal[sw_rt$Block == b]
  ggplot(d, aes(sample = Median_RT)) +
    stat_qq(colour = PAL_BLOCKS[b], alpha = 0.7, size = 2) +
    stat_qq_line(colour = "#C73E1D", linewidth = 1) +
    labs(title    = paste("RT Q-Q:", b),
         subtitle = lbl,
         x = "Expected", y = "Actual RT (ms)")
})

qq_acc <- lapply(blocks, function(b) {
  d   <- person_blocks %>% filter(Block_Name == b)
  lbl <- sw_acc$Normal[sw_acc$Block == b]
  ggplot(d, aes(sample = Mean_Acc)) +
    stat_qq(colour = PAL_BLOCKS[b], alpha = 0.7, size = 2) +
    stat_qq_line(colour = "#C73E1D", linewidth = 1) +
    labs(title    = paste("Acc Q-Q:", b),
         subtitle = lbl,
         x = "Expected", y = "Actual Accuracy")
})

(qq_rt[[1]]  | qq_rt[[2]]  | qq_rt[[3]]) /
(qq_acc[[1]] | qq_acc[[2]] | qq_acc[[3]])
ggsave(file.path(output_dir, "H6_qq_all.png"),
       (qq_rt[[1]]  | qq_rt[[2]]  | qq_rt[[3]]) /
       (qq_acc[[1]] | qq_acc[[2]] | qq_acc[[3]]),
       width = 7, height = 7, dpi = 150)
```

## Test Selection Decision

```{r test-decision, echo=FALSE, results='asis'}
cat("| Outcome | All blocks normal? | Test chosen |\n")
cat("|---|---|---|\n")
cat(sprintf("| Reaction Time | %s | %s |\n",
    ifelse(rt_normal_all,  "Yes — all pass", "No — at least one fails"),
    ifelse(rt_normal_all,  "Repeated Measures ANOVA", "Friedman Test")))
cat(sprintf("| Accuracy | %s | %s |\n",
    ifelse(acc_normal_all, "Yes — all pass", "No — at least one fails"),
    ifelse(acc_normal_all, "Repeated Measures ANOVA", "Friedman Test")))
```

---

# Step E: Statistical Tests

## Test 1 — Reaction Time

```{r test-rt}
if (rt_normal_all) {
  cat("All RT blocks normal → Repeated Measures ANOVA\n\n")
  rm_rt        <- summary(aov(Median_RT ~ Block_Name +
                              Error(participant_ID / Block_Name), data = person_blocks))
  print(rm_rt)
  p_rt         <- rm_rt[["Error: participant_ID:Block_Name"]][[1]][["Pr(>F)"]][1]
  test_name_rt <- "Repeated Measures ANOVA"
  chi_rt       <- NA_real_;  df_rt <- NA_real_
  cat(sprintf("\nRM-ANOVA p-value for RT: %.4f\n", p_rt))
} else {
  cat("At least one RT block non-normal → Friedman Test\n\n")
  friedman_rt  <- friedman.test(Median_RT ~ Block_Name | participant_ID,
                                data = person_blocks)
  chi_rt       <- as.numeric(friedman_rt$statistic)
  df_rt        <- as.numeric(friedman_rt$parameter)
  p_rt         <- as.numeric(friedman_rt$p.value)
  test_name_rt <- "Friedman Rank Sum Test"

  tibble(Outcome = "Reaction Time", Test = test_name_rt,
         Chi_sq  = round(chi_rt, 3), df = df_rt,
         p_value = round(p_rt,   4),
         Sig     = ifelse(p_rt < 0.05, "Yes (p < 0.05)", "No (p >= 0.05)")) %>%
    kable(caption   = "Table 4. Test Result — Reaction Time",
          col.names = c("Outcome","Test","Chi-sq","df","p-value","Significant?"),
          booktabs  = TRUE) %>%
    kable_styling(latex_options     = c("striped","HOLD_position"),
                  bootstrap_options = c("striped","hover"),
                  full_width = FALSE, position = "center") %>%
    row_spec(0, bold = TRUE, background = HDR_COLOR, color = "white")
}

cat(sprintf("\nDECISION (RT): %s (p = %.4f)\n",
    ifelse(p_rt < 0.05, "REJECT H0 — RT changed significantly across blocks",
                        "KEEP H0 — RT did NOT change significantly"),
    p_rt))
```

## Test 2 — Accuracy

```{r test-acc}
if (acc_normal_all) {
  cat("All Accuracy blocks normal → Repeated Measures ANOVA\n\n")
  rm_acc        <- summary(aov(Mean_Acc ~ Block_Name +
                               Error(participant_ID / Block_Name), data = person_blocks))
  print(rm_acc)
  p_acc         <- rm_acc[["Error: participant_ID:Block_Name"]][[1]][["Pr(>F)"]][1]
  test_name_acc <- "Repeated Measures ANOVA"
  chi_acc       <- NA_real_;  df_acc <- NA_real_
  cat(sprintf("\nRM-ANOVA p-value for Accuracy: %.4f\n", p_acc))
} else {
  cat("At least one Accuracy block non-normal → Friedman Test\n\n")
  friedman_acc  <- friedman.test(Mean_Acc ~ Block_Name | participant_ID,
                                 data = person_blocks)
  chi_acc       <- as.numeric(friedman_acc$statistic)
  df_acc        <- as.numeric(friedman_acc$parameter)
  p_acc         <- as.numeric(friedman_acc$p.value)
  test_name_acc <- "Friedman Rank Sum Test"

  tibble(Outcome = "Accuracy", Test = test_name_acc,
         Chi_sq  = round(chi_acc, 3), df = df_acc,
         p_value = round(p_acc,   4),
         Sig     = ifelse(p_acc < 0.05, "Yes (p < 0.05)", "No (p >= 0.05)")) %>%
    kable(caption   = "Table 5. Test Result — Accuracy",
          col.names = c("Outcome","Test","Chi-sq","df","p-value","Significant?"),
          booktabs  = TRUE) %>%
    kable_styling(latex_options     = c("striped","HOLD_position"),
                  bootstrap_options = c("striped","hover"),
                  full_width = FALSE, position = "center") %>%
    row_spec(0, bold = TRUE, background = HDR_COLOR, color = "white")
}

cat(sprintf("\nDECISION (Accuracy): %s (p = %.4f)\n",
    ifelse(p_acc < 0.05, "REJECT H0 — Accuracy changed significantly across blocks",
                         "KEEP H0 — Accuracy did NOT change significantly"),
    p_acc))
```

## Post-Hoc Tests (only if main test is significant)

```{r post-hoc-rt}
cat("=== Post-Hoc: Reaction Time ===\n\n")
if (p_rt < 0.05) {
  pairs <- list(c("1: Early","2: Middle"), c("2: Middle","3: Late"),
                c("1: Early","3: Late"))
  ph_rt <- lapply(pairs, function(pair) {
    common <- intersect(
      person_blocks %>% filter(Block_Name == pair[1]) %>% pull(participant_ID),
      person_blocks %>% filter(Block_Name == pair[2]) %>% pull(participant_ID))
    d1 <- person_blocks %>% filter(Block_Name == pair[1], participant_ID %in% common) %>%
          arrange(participant_ID) %>% pull(Median_RT)
    d2 <- person_blocks %>% filter(Block_Name == pair[2], participant_ID %in% common) %>%
          arrange(participant_ID) %>% pull(Median_RT)
    w  <- wilcox.test(d1, d2, paired = TRUE, exact = FALSE)
    tibble(Comparison = paste(pair[1], "vs", pair[2]),
           W          = round(as.numeric(w$statistic), 1),
           p_raw      = round(as.numeric(w$p.value),   4))
  }) %>% bind_rows() %>%
    mutate(p_adj = round(p.adjust(p_raw, "bonferroni"), 4),
           Sig   = ifelse(p_adj < 0.05, "Yes", "No"))

  ph_rt %>%
    kable(caption   = "Table 6. Post-Hoc Pairwise Tests — Reaction Time (Bonferroni)",
          col.names = c("Comparison","W","Raw p","Adj. p","Significant?"),
          booktabs  = TRUE) %>%
    kable_styling(latex_options     = c("striped","HOLD_position"),
                  bootstrap_options = c("striped","hover"),
                  full_width = FALSE, position = "center") %>%
    row_spec(0, bold = TRUE, background = HDR_COLOR, color = "white")
} else {
  cat("RT test not significant — no post-hoc tests needed.\n")
}
```

```{r post-hoc-acc}
cat("=== Post-Hoc: Accuracy ===\n\n")
if (p_acc < 0.05) {
  pairs <- list(c("1: Early","2: Middle"), c("2: Middle","3: Late"),
                c("1: Early","3: Late"))
  ph_acc <- lapply(pairs, function(pair) {
    common <- intersect(
      person_blocks %>% filter(Block_Name == pair[1]) %>% pull(participant_ID),
      person_blocks %>% filter(Block_Name == pair[2]) %>% pull(participant_ID))
    d1 <- person_blocks %>% filter(Block_Name == pair[1], participant_ID %in% common) %>%
          arrange(participant_ID) %>% pull(Mean_Acc)
    d2 <- person_blocks %>% filter(Block_Name == pair[2], participant_ID %in% common) %>%
          arrange(participant_ID) %>% pull(Mean_Acc)
    w  <- wilcox.test(d1, d2, paired = TRUE, exact = FALSE)
    tibble(Comparison = paste(pair[1], "vs", pair[2]),
           W          = round(as.numeric(w$statistic), 1),
           p_raw      = round(as.numeric(w$p.value),   4))
  }) %>% bind_rows() %>%
    mutate(p_adj = round(p.adjust(p_raw, "bonferroni"), 4),
           Sig   = ifelse(p_adj < 0.05, "Yes", "No"))

  ph_acc %>%
    kable(caption   = "Table 7. Post-Hoc Pairwise Tests — Accuracy (Bonferroni)",
          col.names = c("Comparison","W","Raw p","Adj. p","Significant?"),
          booktabs  = TRUE) %>%
    kable_styling(latex_options     = c("striped","HOLD_position"),
                  bootstrap_options = c("striped","hover"),
                  full_width = FALSE, position = "center") %>%
    row_spec(0, bold = TRUE, background = HDR_COLOR, color = "white")
} else {
  cat("Accuracy test not significant — no post-hoc tests needed.\n")
}
```

---

# Final Conclusion

```{r conclusion, echo=FALSE, results='asis'}
rt_sig     <- p_rt  < 0.05
acc_sig    <- p_acc < 0.05
either_sig <- rt_sig | acc_sig

cat("## Was H6 Supported?\n\n")

cat("### Reaction Time:\n")
if (rt_sig) {
  cat(sprintf("- **REJECTED H0** (p = %.4f) — RT changed significantly across blocks.\n\n", p_rt))
} else {
  cat(sprintf("- **KEPT H0** (p = %.4f) — RT was stable across blocks.\n\n", p_rt))
}

cat("### Accuracy:\n")
if (acc_sig) {
  cat(sprintf("- **REJECTED H0** (p = %.4f) — Accuracy changed significantly across blocks.\n\n", p_acc))
} else {
  cat(sprintf("- **KEPT H0** (p = %.4f) — Accuracy was stable across blocks.\n\n", p_acc))
}

cat("### Overall Conclusion:\n\n")
if (either_sig) {
  cat(paste0(
    "**H6 is SUPPORTED.** There is evidence of performance change across the ",
    "experiment. This suggests cognitive fatigue played a role. Future studies ",
    "should model trial order as a within-subject factor and consider giving ",
    "participants scheduled breaks during long experiments.\n"
  ))
} else {
  cat(paste0(
    "**H6 is NOT supported.** Participants maintained consistent performance ",
    "across all three time blocks. This is a **positive finding** — it means the ",
    "experimental design was robust. The effects found in H1–H4 are unlikely to ",
    "be contaminated by fatigue, since performance did not degrade over time.\n"
  ))
}
```

---

# Data Flow and Working Summary

```
RAW LOG FILES (one per participant)
        |
        | Load ALL rows → isValidation check (column + Event string)
        | Apply formula: Correct Hits > (False Alarms / 2) + Missed
        v
VALID PARTICIPANTS (expected: 112)
        |
        | Keep only IR press events (exclude Practice)
        | Sort each participant's presses by Timestamp (chronological)
        v
FOR EACH PARTICIPANT:
        trial_sequence = 1, 2, 3 ... N  (in time order)
        ntile(sequence, 3) → 3 equal blocks:
            Block 1: Early  = first  1/3 of presses
            Block 2: Middle = second 1/3 of presses
            Block 3: Late   = final  1/3 of presses
        |
        | Compute per person per block:
        |   Median RT   = robust centre of RT distribution
        |   Mean Acc    = proportion of correct presses
        v
BLOCK SCORE TABLE (112 × 3 = 336 rows)
        |
        | Step D: Shapiro-Wilk per block + Q-Q plots (6 tests)
        |   If any block fails normality → Friedman Test
        |   If all pass               → Repeated Measures ANOVA
        v
MAIN TEST (Friedman or RM-ANOVA)
        p < 0.05 → significant change detected
        p >= 0.05 → no change detected
        |
        | If significant:
        |   Pairwise Wilcoxon tests (3 pairs)
        |   Bonferroni correction (multiply p by 3)
        v
FINAL DECISION
        H6 supported   → fatigue effect exists
        H6 unsupported → experiment design is robust
```

## H6 Quick Reference Table

```{r h6-summary-table, echo=FALSE}
# Build safe labels — handles both Friedman and RM-ANOVA results
rt_label <- if (!is.na(chi_rt)) {
  paste0(test_name_rt, ": Chi-sq(", df_rt, ") = ", round(chi_rt, 3),
         ", p = ", round(p_rt, 4))
} else {
  paste0(test_name_rt, ": p = ", round(p_rt, 4))
}

acc_label <- if (!is.na(chi_acc)) {
  paste0(test_name_acc, ": Chi-sq(", df_acc, ") = ", round(chi_acc, 3),
         ", p = ", round(p_acc, 4))
} else {
  paste0(test_name_acc, ": p = ", round(p_acc, 4))
}

tibble(
  Item   = c("What we tested", "Data used", "How we split",
             "Normality check", "Test for RT", "Test for Accuracy",
             "RT result", "Accuracy result",
             "Post-hoc correction", "Overall decision"),
  Detail = c(
    "Do RT and accuracy change across Early, Middle, Late blocks?",
    "IR press events for all valid participants (practice removed)",
    "Each person's presses sorted by Timestamp, split by ntile(n, 3)",
    "Shapiro-Wilk per block + Q-Q plots (6 tests total)",
    rt_label,
    acc_label,
    ifelse(rt_sig,
           paste0("Significant (p = ", round(p_rt,  4), ") — RT changed"),
           paste0("Not significant (p = ", round(p_rt,  4), ") — RT stable")),
    ifelse(acc_sig,
           paste0("Significant (p = ", round(p_acc, 4), ") — Accuracy changed"),
           paste0("Not significant (p = ", round(p_acc, 4), ") — Accuracy stable")),
    "Bonferroni correction on 3 pairwise tests (p x 3)",
    ifelse(either_sig,
           "H6 SUPPORTED — fatigue effect detected",
           "H6 NOT SUPPORTED — performance stable, robust design")
  )
) %>%
  kable(caption   = "Table 8. H6 Complete Summary",
        col.names = c("What We Checked", "Result"),
        booktabs  = TRUE) %>%
  kable_styling(
    latex_options     = c("striped", "HOLD_position"),
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE, position = "center"
  ) %>%
  column_spec(1, bold = TRUE, width = "5cm") %>%
  column_spec(2, width = "10cm") %>%
  row_spec(0, bold = TRUE, background = HDR_COLOR, color = "white") %>%
  row_spec(10, bold = TRUE, background = "#EBF5FB")
```

---

*Report generated on `r Sys.time()` using R `r R.version.string`*
