---
title: "Hypothesis 7: Does Sentence Type Interact with Voice on Recognition Accuracy?"
subtitle: "2 x 4 Non-Parametric Factorial Analysis: Voice (Active/Passive) x Sentence Type (HH/HL/LH/LL)"
author: "Team: Metamatics"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    latex_engine: xelatex
    number_sections: true
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    code_folding: show
    fig_caption: true
    highlight: tango
    df_print: paged
fontsize: 11pt
geometry: margin=1in
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{amsmath}
  - \usepackage{amssymb}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo       = TRUE,
  warning    = FALSE,
  message    = FALSE,
  fig.align  = "center",
  fig.width  = 7,
  fig.height = 4,
  out.width  = "100%",
  dpi        = 150
)

pkgs <- c("dplyr", "tidyr", "tibble", "ggplot2", "readr", "stringr",
          "purrr", "knitr", "kableExtra", "scales", "patchwork", "rstatix", "Hmisc")
for (p in pkgs) {
  if (!requireNamespace(p, quietly = TRUE))
    install.packages(p, repos = "https://cran.r-project.org")
  library(p, character.only = TRUE)
}

PAL_VOICE <- c("Active"  = "#2E86AB", "Passive" = "#A23B72")
PAL_TYPE  <- c("HH" = "#264653", "HL" = "#2A9D8F",
               "LH" = "#E9C46A", "LL" = "#E76F51")
HDR_COLOR <- "#1B3A57"

theme_h7 <- theme_minimal(base_size = 13) +
  theme(
    plot.title       = element_text(face = "bold", size = 14, colour = "#2C2C2C"),
    plot.subtitle    = element_text(colour = "grey40", size = 10.5, lineheight = 1.3),
    plot.caption     = element_text(colour = "grey55", size = 9, hjust = 0),
    legend.position  = "bottom",
    legend.title     = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(colour = "grey92"),
    axis.title       = element_text(face = "bold", size = 11),
    strip.text       = element_text(face = "bold", size = 11)
  )
theme_set(theme_h7)

output_dir <- "R_output"
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)
```

---

# What is Hypothesis 7?

## The Big Idea

In our experiment, every sentence was built using two nouns:

- **The Subject noun** (the "doer" in the sentence)
- **The Object noun** (the "receiver" in the sentence)

Each noun was either **High** (H) or **Low** (L) memorability. This gives us **4 sentence types**:

| Code | Subject | Object | Meaning |
|---|---|---|---|
| **HH** | High | High | Easy to remember both nouns |
| **HL** | High | Low  | Easy subject, hard object |
| **LH** | Low  | High | Hard subject, easy object |
| **LL** | Low  | Low  | Difficult to remember both |

We also have **2 voice conditions**: Active and Passive.
Full design = **2 (Voice) × 4 (Sentence Type) = 8 combinations**.

## The Question

> **Does the difficulty of remembering sentence wording depend BOTH on voice AND on noun memorability?**
> If passive voice hurts memory differently depending on word difficulty, there is an **interaction effect**.

---

# Formal Hypothesis

> **H0:** Voice effect on accuracy is the SAME for all sentence types. No interaction.
>
> **H7:** Voice effect on accuracy DIFFERS depending on sentence type. Interaction exists.

---

# Step A: Loading the Data

```{r load-data}
analysis_path <- file.path(output_dir, "H3H4_step3_analysis.csv")
if (!file.exists(analysis_path)) {
  stop("H3H4_step3_analysis.csv not found in R_output/. Please knit Hypo3and4_analysis.Rmd first.")
}

df_raw <- read_csv(analysis_path, show_col_types = FALSE)

cat(sprintf("Loaded: %s rows, %d participants\n",
            format(nrow(df_raw), big.mark = ","),
            n_distinct(df_raw$participant_ID)))
cat("\n=== ALL COLUMN NAMES ===\n")
cat(paste(" -", names(df_raw), collapse = "\n"))
cat("\n\n=== FIRST 5 ROWS ===\n")
print(head(df_raw, 5))
cat("\n=== UNIQUE Core_ID values (first 20) ===\n")
print(head(unique(df_raw$Core_ID), 20))
if ("Original_Voice" %in% names(df_raw)) {
  cat("\n=== UNIQUE Original_Voice values ===\n")
  print(unique(df_raw$Original_Voice))
}

type_cols <- names(df_raw)[sapply(names(df_raw), function(col) {
  vals <- unique(na.omit(as.character(df_raw[[col]])))
  any(vals %in% c("HH","HL","LH","LL"))
})]
if (length(type_cols) > 0) {
  cat(sprintf("\n=== Columns with HH/HL/LH/LL values: %s ===\n", paste(type_cols, collapse=", ")))
} else {
  cat("\n=== NO column found with HH/HL/LH/LL — will extract from Core_ID ===\n")
}
```

## Extracting the Sentence Type (HH / HL / LH / LL)

```{r extract-sentence-type}
type_col <- names(df_raw)[sapply(names(df_raw), function(col) {
  vals <- unique(na.omit(as.character(df_raw[[col]])))
  any(vals %in% c("HH","HL","LH","LL"))
})]

if (length(type_col) > 0) {
  sentence_type_vec <- as.character(df_raw[[type_col[1]]])
} else {
  raw_type <- str_extract(df_raw$Core_ID, "(?i)(HH|HL|LH|LL|HVL|LVH|LVL)")
  sentence_type_vec <- case_when(
    toupper(raw_type) == "HVL" ~ "HL",
    toupper(raw_type) == "LVH" ~ "LH",
    toupper(raw_type) == "LVL" ~ "LL",
    toupper(raw_type) %in% c("HH","HL","LH","LL") ~ toupper(raw_type),
    TRUE ~ NA_character_
  )
}

cat(sprintf("Sentence types found: %s\nNA count: %d / %d\n",
            paste(head(na.omit(unique(sentence_type_vec)), 20), collapse=", "),
            sum(is.na(sentence_type_vec)), length(sentence_type_vec)))

df <- df_raw %>%
  mutate(
    Sentence_Type = sentence_type_vec,
    Voice_Label   = case_when(
      tolower(trimws(Original_Voice)) == "a" ~ "Active",
      tolower(trimws(Original_Voice)) == "p" ~ "Passive",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(Sentence_Type), !is.na(WR_Accuracy), !is.na(Voice_Label)) %>%
  mutate(Sentence_Type = factor(Sentence_Type), Voice_Label = factor(Voice_Label)) %>%
  droplevels()

cat(sprintf("\nRows after filtering: %s\nVoice levels: %s\nSentence Type levels: %s\n",
            format(nrow(df), big.mark=","),
            paste(levels(df$Voice_Label), collapse=", "),
            paste(levels(df$Sentence_Type), collapse=", ")))
if (nrow(df) > 0) { cat("\nCell counts:\n"); print(table(df$Sentence_Type, df$Voice_Label)) }
```

---

# Step B: Descriptive Statistics

## Table 1 - Mean Accuracy per Cell

```{r desc-table}
if (nrow(df) == 0) {
  cat("No data available.\n")
  desc <- tibble(Voice_Label=character(), Sentence_Type=character(),
                 N=integer(), Mean_Acc=numeric(), Median_Acc=numeric(), SD=numeric(), SE=numeric())
} else {
  desc <- df %>%
    group_by(Voice_Label, Sentence_Type) %>%
    summarise(N=n(), Mean_Acc=round(mean(WR_Accuracy, na.rm=TRUE),4),
              Median_Acc=round(median(WR_Accuracy, na.rm=TRUE),4),
              SD=round(sd(WR_Accuracy, na.rm=TRUE),4),
              SE=round(sd(WR_Accuracy, na.rm=TRUE)/sqrt(n()),4), .groups="drop")

  desc %>%
    kable(caption="Table 1. Descriptive Statistics per Voice x Sentence Type",
          col.names=c("Voice","Type","N","Mean Acc","Median Acc","SD","SE"), booktabs=TRUE) %>%
    kable_styling(latex_options=c("striped","HOLD_position"),
                  bootstrap_options=c("striped","hover"), full_width=FALSE, position="center") %>%
    row_spec(0, bold=TRUE, background=HDR_COLOR, color="white")
}
```

## Table 2 - Summary by Sentence Type

```{r desc-type}
if (nrow(df) > 0) {
  df %>%
    group_by(Sentence_Type) %>%
    summarise(N=n(), Mean_Acc=round(mean(WR_Accuracy,na.rm=TRUE),4),
              SD=round(sd(WR_Accuracy,na.rm=TRUE),4), .groups="drop") %>%
    kable(caption="Table 2. Accuracy by Sentence Type (collapsed across Voice)",
          col.names=c("Sentence Type","N","Mean Accuracy","SD"), booktabs=TRUE) %>%
    kable_styling(latex_options=c("striped","HOLD_position"),
                  bootstrap_options=c("striped","hover"), full_width=FALSE, position="center") %>%
    row_spec(0, bold=TRUE, background=HDR_COLOR, color="white")
}
```

---

# Step C: Visualising the Data

## Figure 1 - Bar Chart

```{r fig1-bar, fig.cap="Figure 1. Mean WR accuracy per Voice x Sentence Type cell.", fig.width=7, fig.height=5}
if (nrow(desc) == 0) { cat("Figure 1 skipped.\n") } else {
  # Join individual data for jittering over the summary bars
  p1 <- ggplot() +
    geom_col(data=desc, aes(x=Sentence_Type, y=Mean_Acc, fill=Voice_Label, group=Voice_Label),
             position=position_dodge(0.75), width=0.65, alpha=0.85, colour="black", linewidth=0.3) +
    geom_jitter(data=df, aes(x=Sentence_Type, y=WR_Accuracy, fill=Voice_Label, group=Voice_Label),
                position=position_jitterdodge(jitter.width=0.15, dodge.width=0.75),
                alpha=0.25, size=1.5, shape=21, colour="grey30", show.legend=FALSE) +
    geom_errorbar(data=desc, aes(x=Sentence_Type, ymin=Mean_Acc-SE, ymax=Mean_Acc+SE, group=Voice_Label),
                  position=position_dodge(0.75), width=0.25, colour="black", linewidth=0.8) +
    scale_fill_manual(values=PAL_VOICE, name="Voice") +
    scale_y_continuous(limits=c(0,1), labels=percent_format()) +
    labs(title="Figure 1: WR Accuracy by Voice and Sentence Type",
         subtitle="Bars = group mean | Error bars = \u00b11 SE | Dots = individual participants",
         x="Sentence Type", y="Mean WR Accuracy")
  print(p1)
  ggsave(file.path(output_dir,"H7_bar_8cells.png"), p1, width=7, height=5, dpi=150)
}
```

## Figure 2 - Interaction Plot

```{r fig2-interaction, fig.cap="Figure 2. Interaction plot.", fig.width=7, fig.height=5}
if (nrow(desc) == 0) { cat("Figure 2 skipped.\n") } else {
  p2 <- ggplot(desc, aes(x=Sentence_Type, y=Mean_Acc, colour=Voice_Label, group=Voice_Label)) +
    geom_line(linewidth=1.5, alpha=0.8) + 
    geom_errorbar(aes(ymin=Mean_Acc-SE, ymax=Mean_Acc+SE), width=0.15, linewidth=1, alpha=0.7) +
    geom_point(size=5, shape=21, aes(fill=Voice_Label), colour="white", stroke=1.5) +
    scale_colour_manual(values=PAL_VOICE, name="Voice") +
    scale_fill_manual(values=PAL_VOICE, name="Voice") +
    scale_y_continuous(limits=c(0,1), labels=percent_format()) +
    labs(title="Figure 2: Interaction Plot -- Voice \u00d7 Sentence Type",
         subtitle="Parallel lines = NO interaction. Crossing = INTERACTION exists.",
         x="Sentence Type", y="Mean WR Accuracy")
  print(p2)
  ggsave(file.path(output_dir,"H7_interaction_plot.png"), p2, width=7, height=5, dpi=150)
}
```

## Figure 3 - Violin Plots

```{r fig3-violin, fig.cap="Figure 3. Accuracy distribution.", fig.width=7, fig.height=5}
if (nrow(df) == 0) { cat("Figure 3 skipped.\n") } else {
  p3 <- ggplot(df, aes(x=Sentence_Type, y=WR_Accuracy, fill=Voice_Label)) +
    geom_violin(alpha=0.35, position=position_dodge(0.8), width=0.75, colour="black", linewidth=0.5, trim=FALSE) +
    geom_boxplot(width=0.18, alpha=0.85, position=position_dodge(0.8), outlier.shape=NA, colour="grey20", fill="white") +
    geom_jitter(aes(colour=Voice_Label), position=position_jitterdodge(jitter.width=0.1, dodge.width=0.8),
                alpha=0.4, size=1.8, show.legend=FALSE) +
    stat_summary(fun=mean, geom="point", position=position_dodge(0.8), shape=23, size=4, fill="white", colour="black", stroke=1.2) +
    scale_fill_manual(values=PAL_VOICE, name="Voice") +
    scale_colour_manual(values=PAL_VOICE, name="Voice") +
    scale_y_continuous(limits=c(0,1.05), labels=percent_format(), breaks=seq(0,1,0.25)) +
    labs(title="Figure 3: Accuracy Distribution by Sentence Type and Voice",
         subtitle="Violin = full distribution | Box = median & IQR | Diamond = mean | Dots = participants",
         x="Sentence Type", y="WR Accuracy")
  print(p3)
  ggsave(file.path(output_dir,"H7_violin.png"), p3, width=7, height=5, dpi=150)
}
```

## Figure 4 - Heatmap

```{r fig4-heatmap, fig.cap="Figure 4. Heatmap of mean accuracy.", fig.width=7, fig.height=4}
if (nrow(desc) == 0) { cat("Figure 4 skipped.\n") } else {
  p4 <- ggplot(desc, aes(x=Sentence_Type, y=Voice_Label, fill=Mean_Acc)) +
    geom_tile(colour="white", linewidth=1.5, width=0.9, height=0.9) +
    geom_text(aes(label=scales::percent(Mean_Acc, accuracy=0.1)),
              size=5, fontface="bold", colour="white") +
    scale_fill_gradient(low="#2E4057", high="#06D6A0", name="Mean\nAccuracy", labels=percent_format()) +
    labs(title="Figure 4: Heatmap of Mean WR Accuracy", x="Sentence Type", y="Voice")
  print(p4)
  ggsave(file.path(output_dir,"H7_heatmap.png"), p4, width=7, height=4, dpi=150)
}
```

## Figure 5 - Faceted Boxplots

```{r fig5-facet, fig.cap="Figure 5. One panel per sentence type.", fig.width=7, fig.height=5}
if (nrow(df) == 0 || nlevels(df$Sentence_Type) < 1) { cat("Figure 5 skipped.\n") } else {
  df_plot5 <- df %>% mutate(Sentence_Type=droplevels(Sentence_Type))
  type_labels <- setNames(
    paste0(levels(df_plot5$Sentence_Type), "\n",
           c(HH="(both easy)",HL="(easy-hard)",LH="(hard-easy)",LL="(both hard)")[levels(df_plot5$Sentence_Type)]),
    levels(df_plot5$Sentence_Type))
  p5 <- ggplot(df_plot5, aes(x=Voice_Label, y=WR_Accuracy, fill=Voice_Label)) +
    geom_violin(alpha=0.25, width=0.9, trim=FALSE, colour="grey40") +
    geom_boxplot(width=0.25, alpha=0.85, outlier.shape=NA, colour="grey25") +
    geom_jitter(aes(colour=Voice_Label), width=0.1, alpha=0.35, size=1.5, show.legend=FALSE) +
    stat_summary(fun=mean, geom="point", shape=23, size=4, fill="white", colour="black", stroke=1.2) +
    scale_fill_manual(values=PAL_VOICE, guide="none") +
    scale_colour_manual(values=PAL_VOICE, guide="none") +
    scale_y_continuous(limits=c(0,1.05), labels=percent_format()) +
    facet_wrap(~Sentence_Type, nrow=1, labeller=labeller(Sentence_Type=type_labels)) +
    labs(title="Figure 5: Active vs Passive — faceted by Sentence Type",
         subtitle="Diamond = group mean | Box = median & IQR | Dots = individuals",
         x="Voice", y="WR Accuracy")
  print(p5)
  ggsave(file.path(output_dir,"H7_facet_boxplots.png"), p5, width=7, height=5, dpi=150)
}
```

---

# Step D: Normality Check

```{r normality-check}
df_person <- df %>%
  group_by(participant_ID, Voice_Label, Sentence_Type) %>%
  summarise(mean_acc=mean(WR_Accuracy, na.rm=TRUE), .groups="drop")

cells <- df_person %>% distinct(Voice_Label, Sentence_Type)

sw_results <- df_person %>%
  group_by(Voice_Label, Sentence_Type) %>%
  summarise(
    n_obs = n(),
    W = if(n() >= 3) round(shapiro.test(mean_acc)$statistic, 4) else NA_real_,
    p_sw = if(n() >= 3) round(shapiro.test(mean_acc)$p.value, 4) else NA_real_,
    .groups = "drop"
  ) %>%
  mutate(Normal = case_when(
    is.na(p_sw) ~ "Too few obs",
    p_sw > 0.05 ~ "Yes (p > 0.05)",
    TRUE ~ "No (p <= 0.05)"
  )) %>%
  mutate(
    cell_col = case_when(Normal=="Yes (p > 0.05)" ~ "#1B5E20", Normal=="No (p <= 0.05)" ~ "#B71C1C", TRUE ~ "#616161"),
    cell_bg  = case_when(Normal=="Yes (p > 0.05)" ~ "#E8F5E9", Normal=="No (p <= 0.05)" ~ "#FDECEA", TRUE ~ "#EEEEEE")
  )

sw_results %>%
  select(Voice_Label, Sentence_Type, n_obs, W, p_sw, Normal) %>%
  kable(caption="Table 3. Shapiro-Wilk Normality Test — 8 cells",
        col.names=c("Voice","Type","N obs","W","p-value","Normal?"), booktabs=TRUE) %>%
  kable_styling(latex_options=c("striped","HOLD_position"),
                bootstrap_options=c("striped","hover"), full_width=FALSE, position="center") %>%
  column_spec(6, bold=TRUE, color=sw_results$cell_col, background=sw_results$cell_bg) %>%
  row_spec(0, bold=TRUE, background=HDR_COLOR, color="white")

any_non_normal <- any(sw_results$Normal=="No (p <= 0.05)", na.rm=TRUE)
cat(sprintf("\nAny cell non-normal? %s\n",
            ifelse(any_non_normal,"YES -- will use Scheirer-Ray-Hare","NO -- may use 2-way ANOVA")))
```

## Q-Q Plots for Each Cell

```{r qq-grid, fig.cap="Figure 6. Q-Q plots for all Voice x Sentence Type cells.", fig.width=7, fig.height=7}
n_cells <- nrow(cells)
qq_list <- vector("list", n_cells)
for (i in seq_len(n_cells)) {
  v  <- as.character(cells$Voice_Label[i])
  st <- as.character(cells$Sentence_Type[i])
  d  <- df_person %>% filter(Voice_Label==v, Sentence_Type==st)
  lbl_row <- sw_results %>% filter(as.character(Voice_Label)==v, as.character(Sentence_Type)==st)
  lbl <- if (nrow(lbl_row)>0) lbl_row$Normal[1] else "No result"
  qq_list[[i]] <- ggplot(d, aes(sample=mean_acc)) +
    stat_qq(colour=PAL_VOICE[v], alpha=0.7, size=2) +
    stat_qq_line(colour="#C73E1D", linewidth=1) +
    labs(title=paste(v,"-",st), subtitle=lbl, x="Expected", y="Actual Acc")
}
qq_grid <- wrap_plots(qq_list, ncol=min(4,n_cells))
print(qq_grid)
ggsave(file.path(output_dir,"H7_qq_grid.png"), qq_grid,
       width=7, height=ceiling(n_cells/4)*3.5, dpi=150)
```

## Test Selection Decision

```{r test-decision, echo=FALSE, results='asis'}
cat("### Which test will be used?\n\n")
if (any_non_normal) {
  cat("**Scheirer-Ray-Hare Test** — at least one cell failed normality.\n")
} else {
  cat("**2-way ANOVA** — all cells passed normality.\n")
}
```

---

# Step E: Main Statistical Test

```{r scheirer-ray-hare}
df_srh <- df_person %>%
  mutate(Rank_Acc=rank(mean_acc),
         Voice_Label=droplevels(factor(Voice_Label)),
         Sentence_Type=droplevels(factor(Sentence_Type)))

n_voice <- nlevels(df_srh$Voice_Label)
n_type  <- nlevels(df_srh$Sentence_Type)
cat(sprintf("Voice levels: %d (%s)\nSentence Type levels: %d (%s)\n",
            n_voice, paste(levels(df_srh$Voice_Label),collapse=", "),
            n_type,  paste(levels(df_srh$Sentence_Type),collapse=", ")))

if (n_voice < 2 || n_type < 2) {
  cat("\n!! CANNOT RUN: one factor has < 2 levels.\n")
  H_voice<-NA_real_; p_voice<-NA_real_; df_voice<-NA_real_
  H_type <-NA_real_; p_type <-NA_real_; df_type <-NA_real_
  H_inter<-NA_real_; p_inter<-NA_real_; df_inter<-NA_real_
} else {
  N      <- nrow(df_srh)
  ss_tot <- sum((df_srh$Rank_Acc - mean(df_srh$Rank_Acc))^2)
  srh_fit <- aov(Rank_Acc ~ Voice_Label * Sentence_Type, data=df_srh)
  srh_sum <- summary(srh_fit)[[1]]
  rownames(srh_sum) <- trimws(rownames(srh_sum))

  ss_voice <- srh_sum["Voice_Label",               "Sum Sq"]
  ss_type  <- srh_sum["Sentence_Type",             "Sum Sq"]
  ss_inter <- srh_sum["Voice_Label:Sentence_Type", "Sum Sq"]
  df_voice <- srh_sum["Voice_Label",               "Df"]
  df_type  <- srh_sum["Sentence_Type",             "Df"]
  df_inter <- srh_sum["Voice_Label:Sentence_Type", "Df"]

  H_voice <- (ss_voice/ss_tot)*(N-1); p_voice <- pchisq(H_voice,df=df_voice,lower.tail=FALSE)
  H_type  <- (ss_type /ss_tot)*(N-1); p_type  <- pchisq(H_type, df=df_type, lower.tail=FALSE)
  H_inter <- (ss_inter/ss_tot)*(N-1); p_inter <- pchisq(H_inter,df=df_inter,lower.tail=FALSE)

  srh_table <- tibble(
    Effect=c("Voice (Active vs Passive)","Sentence Type (HH/HL/LH/LL)",
             "Voice x Sentence Type (Interaction)"),
    df=c(df_voice,df_type,df_inter),
    SS_Ranks=round(c(ss_voice,ss_type,ss_inter),2),
    H_statistic=round(c(H_voice,H_type,H_inter),3),
    p_value=round(c(p_voice,p_type,p_inter),4),
    Significant=case_when(
      is.na(c(p_voice,p_type,p_inter)) ~ "Cannot compute",
      c(p_voice,p_type,p_inter) < 0.05 ~ "Yes (p < 0.05)",
      TRUE ~ "No (p >= 0.05)"
    )
  ) %>%
  mutate(
    cell_col = case_when(Significant=="Yes (p < 0.05)" ~ "#1B5E20", Significant=="No (p >= 0.05)" ~ "#B71C1C", TRUE ~ "#616161"),
    cell_bg  = case_when(Significant=="Yes (p < 0.05)" ~ "#E8F5E9", Significant=="No (p >= 0.05)" ~ "#FDECEA", TRUE ~ "#EEEEEE")
  )

  srh_table %>%
    select(Effect, df, SS_Ranks, H_statistic, p_value, Significant) %>%
    kable(caption="Table 4. Scheirer-Ray-Hare Test (Non-Parametric 2-way Factorial)",
          col.names=c("Effect","df","SS (Ranks)","H statistic","p-value","Significant?"),
          booktabs=TRUE) %>%
    kable_styling(latex_options=c("striped","HOLD_position"),
                  bootstrap_options=c("striped","hover"), full_width=FALSE, position="center") %>%
    column_spec(6, bold=TRUE, color=srh_table$cell_col, background=srh_table$cell_bg) %>%
    row_spec(0, bold=TRUE, background=HDR_COLOR, color="white")
}
```

---

# Step F: Post-Hoc Tests

```{r post-hoc-type}
if (is.na(p_type)) {
  cat("Post-hoc skipped: main test could not run.\n")
} else if (isTRUE(p_type < 0.05)) {
  type_pairs <- combn(levels(df_srh$Sentence_Type), 2, simplify=FALSE)
  ph_type <- lapply(type_pairs, function(pair) {
    d1 <- df_srh %>% filter(Sentence_Type==pair[1]) %>% pull(mean_acc)
    d2 <- df_srh %>% filter(Sentence_Type==pair[2]) %>% pull(mean_acc)
    w  <- wilcox.test(d1, d2, exact=FALSE)
    tibble(Comparison=paste(pair[1],"vs",pair[2]),
           W=round(as.numeric(w$statistic),1), p_raw=round(as.numeric(w$p.value),4))
  }) %>% bind_rows() %>%
    mutate(p_adj=round(p.adjust(p_raw,method="bonferroni"),4),
           Significant=ifelse(p_adj<0.05,"Yes","No"))

  ph_type %>%
    kable(caption="Table 5. Post-Hoc Pairwise Tests (Bonferroni corrected)",
          col.names=c("Comparison","W","Raw p","Adjusted p","Significant?"), booktabs=TRUE) %>%
    kable_styling(latex_options=c("striped","HOLD_position"),
                  bootstrap_options=c("striped","hover"), full_width=FALSE, position="center") %>%
    row_spec(0, bold=TRUE, background=HDR_COLOR, color="white")
} else {
  cat(sprintf("Sentence Type NOT significant (p = %.4f). No post-hoc needed.\n", p_type))
}
```

---

# Step G: Complete From-Scratch Analysis — Voice × Lexical Interaction on Corrected Memorability

> Rigorous 5-step pipeline on `Corrected_Memorability`. Every step has
> defensive guards so the document **always** knits to PDF without crashing.

---

## Data Preparation (Cell Isolation)

> **What we do:** Aggregate trials into **8 cells** per participant:
> Active-HH, Passive-HH, Active-HL, Passive-HL, Active-LH, Passive-LH, Active-LL, Passive-LL.
> Cell score = `mean(corrected_memorability, na.rm = TRUE)`.

```{r g-data-prep}
cm_path <- file.path(output_dir, "H3H4_step3_analysis.csv")
if (!file.exists(cm_path)) stop("H3H4_step3_analysis.csv not found.")
df_cm_raw <- read_csv(cm_path, show_col_types=FALSE)

lc_raw <- str_extract(df_cm_raw$Core_ID, "(?i)(HH|HL|LH|LL|HVL|LVH|LVL)")
lc_vec <- case_when(
  toupper(lc_raw) == "HVL" ~ "HL",
  toupper(lc_raw) == "LVH" ~ "LH",
  toupper(lc_raw) == "LVL" ~ "LL",
  toupper(lc_raw) %in% c("HH","HL","LH","LL") ~ toupper(lc_raw),
  TRUE ~ NA_character_
)

df_cm <- df_cm_raw %>%
  mutate(
    Lexical_Condition = factor(toupper(lc_vec), levels=c("HH","HL","LH","LL")),
    Voice = factor(case_when(
      toupper(trimws(Original_Voice)) %in% c("A","ACTIVE","Active")   ~ "Active",
      toupper(trimws(Original_Voice)) %in% c("P","PASSIVE","Passive") ~ "Passive",
      TRUE ~ NA_character_
    ), levels=c("Active","Passive"))
  ) %>%
  filter(!is.na(Lexical_Condition), !is.na(corrected_memorability), !is.na(Voice)) %>%
  group_by(participant_ID, Lexical_Condition, Voice) %>%
  summarise(Score=mean(corrected_memorability, na.rm=TRUE), .groups="drop") %>%
  mutate(Lexical_Condition=droplevels(Lexical_Condition), Voice=droplevels(Voice))

cat("=== Data Preparation Complete ===\n")
cat(sprintf("  Total rows          : %d\n  Unique participants : %d\n  Lexical levels      : %s\n  Voice levels        : %s\n",
    nrow(df_cm), n_distinct(df_cm$participant_ID),
    paste(levels(df_cm$Lexical_Condition),collapse=", "),
    paste(levels(df_cm$Voice),collapse=", ")))
cat("\n  Cell counts:\n")
print(table(df_cm$Lexical_Condition, df_cm$Voice))
```

---

## Reference-Style Exploratory Plots

> The plots below match the style of the reference figures you shared:
> **Salmon = Active, Teal = Passive**.
> They give a clean visual summary before the formal statistical test.

```{r g-ref-colours, include=FALSE}
# Colour palette matching the reference images
PAL_REF <- c("Active" = "#F08070", "Passive" = "#40C4B5")
```

### Plot R1 — Corrected Memorability Score by Sentence Type and Voice

```{r g-ref-bar, fig.width=8, fig.height=5, fig.cap="Corrected memorability score by sentence type and voice. Error bars = ±1 SE."}
if (nrow(df_cm) == 0) {
  cat("Plot R1 skipped: no data.\n")
} else {
  ref_desc <- df_cm %>%
    group_by(Lexical_Condition, Voice) %>%
    summarise(
      Mean_CM = mean(Score, na.rm = TRUE),
      SE_CM   = sd(Score,   na.rm = TRUE) / sqrt(n()),
      .groups = "drop"
    )

  p_ref_bar <- ggplot(ref_desc,
    aes(x = Lexical_Condition, y = Mean_CM, fill = Voice, group = Voice)) +
    geom_col(position = position_dodge(0.7), width = 0.6, colour = NA) +
    geom_errorbar(
      aes(ymin = Mean_CM - SE_CM, ymax = Mean_CM + SE_CM),
      position = position_dodge(0.7), width = 0.2,
      colour = "black", linewidth = 0.7
    ) +
    scale_fill_manual(values = PAL_REF, name = "Voice") +
    scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.1))) +
    labs(
      title = "Corrected memorability score by sentence type and voice",
      x     = "Sentence type",
      y     = "Corrected memorability score"
    ) +
    theme_classic(base_size = 13) +
    theme(
      plot.title        = element_text(hjust = 0.5, size = 13),
      legend.position   = c(0.92, 0.85),
      legend.background = element_rect(fill = "white", colour = "grey80"),
      axis.line         = element_line(colour = "grey40")
    )

  print(p_ref_bar)
  ggsave(file.path(output_dir, "H7_Ref_Bar_CM.png"),
         p_ref_bar, width = 8, height = 5, dpi = 150)
}
```

### Plot R2 — Distribution of Corrected Memorability Scores (Violin)

```{r g-ref-violin, fig.width=9, fig.height=5.5, fig.cap="Distribution of corrected memorability scores across all 8 conditions. Dot = group mean."}
if (nrow(df_cm) == 0) {
  cat("Plot R2 skipped: no data.\n")
} else {
  df_violin <- df_cm %>%
    mutate(
      Condition = factor(
        paste0(Lexical_Condition, "-", Voice),
        levels = c("HH-Active","HL-Active","LH-Active","LL-Active",
                   "HH-Passive","HL-Passive","LH-Passive","LL-Passive")
      )
    )

  p_ref_violin <- ggplot(df_violin,
    aes(x = Condition, y = Score, fill = Voice)) +
    geom_violin(alpha = 0.75, colour = "black", linewidth = 0.35, trim = FALSE) +
    stat_summary(fun = mean, geom = "point", colour = "black",
                 size = 3, shape = 21, fill = "black") +
    scale_fill_manual(values = PAL_REF, name = "Voice") +
    labs(
      title = "Distribution of corrected memorability scores",
      x     = "Condition",
      y     = "Corrected memorability score"
    ) +
    theme_classic(base_size = 12) +
    theme(
      plot.title        = element_text(hjust = 0.5, size = 13),
      axis.text.x       = element_text(angle = 25, hjust = 1, size = 9),
      legend.position   = c(0.92, 0.85),
      legend.background = element_rect(fill = "white", colour = "grey80")
    )

  print(p_ref_violin)
  ggsave(file.path(output_dir, "H7_Ref_Violin_CM.png"),
         p_ref_violin, width = 9, height = 5.5, dpi = 150)
}
```

### Plot R3 — Recognition Hit Rate by Sentence Type and Voice

```{r g-ref-hitrate, fig.width=8, fig.height=5, fig.cap="Recognition (IR) hit rate by sentence type and voice. Error bars = ±1 SE."}
if (!exists("df_cm_raw") || !"WR_Accuracy" %in% names(df_cm_raw)) {
  cat("Plot R3 skipped: WR_Accuracy column not found.\n")
  cat(sprintf("Available columns: %s\n", paste(names(df_cm_raw), collapse = ", ")))
} else {
  lc_raw_r3 <- str_extract(df_cm_raw$Core_ID, "(?i)(HH|HL|LH|LL|HVL|LVH|LVL)")
  lc_vec_r3 <- case_when(
    toupper(lc_raw_r3) == "HVL" ~ "HL",
    toupper(lc_raw_r3) == "LVH" ~ "LH",
    toupper(lc_raw_r3) == "LVL" ~ "LL",
    toupper(lc_raw_r3) %in% c("HH","HL","LH","LL") ~ toupper(lc_raw_r3),
    TRUE ~ NA_character_
  )
  df_ir <- df_cm_raw %>%
    mutate(
      Lexical_Condition = factor(toupper(lc_vec_r3), levels = c("HH","HL","LH","LL")),
      Voice = factor(case_when(
        toupper(trimws(Original_Voice)) %in% c("A","ACTIVE","Active")   ~ "Active",
        toupper(trimws(Original_Voice)) %in% c("P","PASSIVE","Passive") ~ "Passive",
        TRUE ~ NA_character_), levels = c("Active","Passive"))
    ) %>%
    filter(!is.na(Lexical_Condition), !is.na(WR_Accuracy), !is.na(Voice)) %>%
    group_by(participant_ID, Lexical_Condition, Voice) %>%
    summarise(Hit_Rate = mean(WR_Accuracy, na.rm = TRUE), .groups = "drop") %>%
    mutate(Lexical_Condition = droplevels(Lexical_Condition),
           Voice             = droplevels(Voice))

  if (nrow(df_ir) == 0) {
    cat("Plot R3 skipped: empty after filtering.\n")
  } else {
    ref_ir <- df_ir %>%
      group_by(Lexical_Condition, Voice) %>%
      summarise(Mean_HR = mean(Hit_Rate, na.rm = TRUE),
                SE_HR   = sd(Hit_Rate, na.rm = TRUE) / sqrt(n()),
                .groups = "drop")

    p_ref_hr <- ggplot(ref_ir,
      aes(x = Lexical_Condition, y = Mean_HR, fill = Voice, group = Voice)) +
      geom_col(position = position_dodge(0.7), width = 0.6, colour = NA) +
      geom_errorbar(
        aes(ymin = Mean_HR - SE_HR, ymax = Mean_HR + SE_HR),
        position = position_dodge(0.7), width = 0.2,
        colour = "black", linewidth = 0.7
      ) +
      scale_fill_manual(values = PAL_REF, name = "Voice") +
      scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.25)) +
      labs(
        title = "Recognition hit rate for repeated target sentences",
        x     = "Sentence type",
        y     = "IR hit rate"
      ) +
      theme_classic(base_size = 13) +
      theme(
        plot.title        = element_text(hjust = 0.5, size = 13),
        legend.position   = c(0.92, 0.15),
        legend.background = element_rect(fill = "white", colour = "grey80")
      )

    print(p_ref_hr)
    ggsave(file.path(output_dir, "H7_Ref_HitRate.png"),
           p_ref_hr, width = 8, height = 5, dpi = 150)
  }
}
```

### Plot R4 — Reaction Time for Correctly Detected Targets (Boxplot)

```{r g-ref-rt, fig.width=9, fig.height=5.5, fig.cap="Reaction time for correctly detected repeated targets across all 8 conditions."}
# Auto-detect a reaction time column from common names
rt_col <- intersect(
  names(df_cm_raw),
  c("RT","reaction_time","Reaction_Time","rt","RT_ms","WR_RT",
    "response_time","Response_Time","mean_rt","IR_RT","ir_rt")
)

if (length(rt_col) == 0) {
  cat("Plot R4 skipped: no reaction time column found in the dataset.\n")
  cat(sprintf("Available columns: %s\n", paste(names(df_cm_raw), collapse = ", ")))
} else {
  rt_col <- rt_col[1]
  cat(sprintf("Using column '%s' for reaction time.\n", rt_col))

  lc_raw_r4 <- str_extract(df_cm_raw$Core_ID, "(?i)(HH|HL|LH|LL|HVL|LVH|LVL)")
  lc_vec_r4 <- case_when(
    toupper(lc_raw_r4) == "HVL" ~ "HL",
    toupper(lc_raw_r4) == "LVH" ~ "LH",
    toupper(lc_raw_r4) == "LVL" ~ "LL",
    toupper(lc_raw_r4) %in% c("HH","HL","LH","LL") ~ toupper(lc_raw_r4),
    TRUE ~ NA_character_
  )
  df_rt <- df_cm_raw %>%
    rename(RT_val = all_of(rt_col)) %>%
    mutate(
      Lexical_Condition = factor(toupper(lc_vec_r4), levels = c("HH","HL","LH","LL")),
      Voice = factor(case_when(
        toupper(trimws(Original_Voice)) %in% c("A","ACTIVE","Active")   ~ "Active",
        toupper(trimws(Original_Voice)) %in% c("P","PASSIVE","Passive") ~ "Passive",
        TRUE ~ NA_character_), levels = c("Active","Passive")),
      Condition = factor(
        paste0(Lexical_Condition, "-", Voice),
        levels = c("HH-Active","HL-Active","LH-Active","LL-Active",
                   "HH-Passive","HL-Passive","LH-Passive","LL-Passive"))
    ) %>%
    filter(!is.na(Condition), !is.na(RT_val), RT_val > 0)

  if (nrow(df_rt) == 0) {
    cat("Plot R4 skipped: no valid RT rows after filtering.\n")
  } else {
    p_ref_rt <- ggplot(df_rt,
      aes(x = Condition, y = RT_val, fill = Voice)) +
      geom_boxplot(outlier.colour = "grey60", outlier.size = 1.2,
                   colour = "black", linewidth = 0.4, alpha = 0.8) +
      scale_fill_manual(values = PAL_REF, name = "Voice") +
      scale_y_log10(labels = scales::comma) +
      labs(
        title = "Reaction time for correctly detected repeated targets",
        x     = "Condition",
        y     = "Mean IR reaction time (ms)"
      ) +
      theme_classic(base_size = 12) +
      theme(
        plot.title        = element_text(hjust = 0.5, size = 13),
        axis.text.x       = element_text(angle = 25, hjust = 1, size = 9),
        legend.position   = c(0.92, 0.85),
        legend.background = element_rect(fill = "white", colour = "grey80")
      )

    print(p_ref_rt)
    ggsave(file.path(output_dir, "H7_Ref_RT_Boxplot.png"),
           p_ref_rt, width = 9, height = 5.5, dpi = 150)
  }
}
```

---

## Step 1 — Per-Group Normality Testing (8 Independent Shapiro-Wilk Tests)

> **Why check each cell separately?** A 2-way ANOVA assumes the data *inside each group*
> looks like a bell curve. Checking the whole dataset hides per-cell problems.
> We run one Shapiro-Wilk per cell — 8 independent tests.
>
> **Rule:** If **any** cell has p < 0.05 → data is **Non-Parametric** → must use Scheirer-Ray-Hare.

```{r g-normality}
if (nrow(df_cm) < 3) {
  cat("!! Too few rows for normality testing.\n")
  g_is_parametric <- FALSE
  g_sw <- tibble(Voice=character(), Lexical_Condition=character(),
                 N=integer(), W=numeric(), p_value=numeric(), Verdict=character())
} else {
  g_sw <- df_cm %>%
    group_by(Voice, Lexical_Condition) %>%
    summarise(
      N       = n(),
      W       = ifelse(n()>=3, round(shapiro.test(Score)$statistic,4), NA_real_),
      p_value = ifelse(n()>=3, round(shapiro.test(Score)$p.value,4),   NA_real_),
      .groups = "drop"
    ) %>%
    mutate(Verdict = case_when(
      is.na(p_value) ~ "Too few observations",
      p_value < 0.05 ~ "Non-Parametric (p < 0.05) — NOT a bell curve",
      TRUE           ~ "Normal (p >= 0.05) — looks like a bell curve"
    ))

  g_sw %>%
    kable(caption="Shapiro-Wilk — one independent test per cell (8 cells)",
          col.names=c("Voice","Lexical","N","W","p-value","Verdict"), booktabs=TRUE) %>%
    kable_styling(latex_options=c("striped","HOLD_position"),
                  bootstrap_options=c("striped","hover"), full_width=FALSE, position="center") %>%
    row_spec(0, bold=TRUE, background=HDR_COLOR, color="white")

  g_is_parametric <- all(g_sw$p_value>=0.05, na.rm=TRUE)
}

cat("\n=== NORMALITY DECISION ===\n")
if (g_is_parametric) cat("All 8 cells passed. Data is PARAMETRIC.\n") else
  cat("At least one cell FAILED. Data is NON-PARAMETRIC.\n")
```

### Q-Q Plot Grid (2 × 4 — one panel per cell)

> Points on the red line = normal. Points curving away = NOT normal.

```{r g-qq-grid, fig.width=10, fig.height=5, fig.cap="2x4 Q-Q Plot Grid: one panel per Voice x Lexical cell."}
if (nrow(df_cm) == 0) { cat("Q-Q grid skipped: no data.\n") } else {
  g_cells   <- df_cm %>% distinct(Voice, Lexical_Condition)
  qq_panels <- vector("list", nrow(g_cells))

  for (i in seq_len(nrow(g_cells))) {
    v  <- as.character(g_cells$Voice[i])
    lc <- as.character(g_cells$Lexical_Condition[i])
    d  <- df_cm %>% filter(Voice==v, Lexical_Condition==lc)
    sw_row <- g_sw %>% filter(as.character(Voice)==v, as.character(Lexical_Condition)==lc)
    subtitle_lbl <- if (nrow(sw_row)>0) sw_row$Verdict[1] else ""

    qq_panels[[i]] <- ggplot(d, aes(sample=Score)) +
      stat_qq(colour=PAL_VOICE[v], alpha=0.7, size=2) +
      stat_qq_line(colour="#C73E1D", linewidth=1) +
      labs(title=paste(v,"—",lc), subtitle=subtitle_lbl, x="Expected", y="Observed") +
      theme_minimal(base_size=9) +
      theme(plot.title=element_text(size=10,face="bold"),
            plot.subtitle=element_text(size=7,colour="grey50"))
  }

  g_qq_plot <- wrap_plots(qq_panels, ncol=4)
  print(g_qq_plot)  # print() ensures this appears in the PDF
  ggsave(file.path(output_dir,"H7_CM_QQ_Grid.png"), g_qq_plot, width=10, height=5, dpi=150)
}
```

### Density Plot

```{r g-density, fig.width=9, fig.height=4, fig.cap="Density curves by Lexical Condition and Voice."}
if (nrow(df_cm) == 0) { cat("Density plot skipped.\n") } else {
  p_dens <- ggplot(df_cm, aes(x=Score, fill=Voice)) +
    geom_density(alpha=0.45) +
    facet_wrap(~Lexical_Condition, nrow=1) +
    scale_fill_manual(values=PAL_VOICE) +
    labs(title="Density Plot — Corrected Memorability by Condition",
         x="Corrected Memorability", y="Density") +
    theme_minimal(base_size=11)
  print(p_dens)
  ggsave(file.path(output_dir,"H7_CM_Density.png"), p_dens, width=9, height=4, dpi=150)
}
```

---

## Step 2 — Statistical Decision Logic

```{r g-test, results='asis'}
cm_n_lc    <- nlevels(df_cm$Lexical_Condition)
cm_n_voice <- nlevels(df_cm$Voice)
cat(sprintf("Factor check: Lexical = %d levels | Voice = %d levels\n\n", cm_n_lc, cm_n_voice))

if (cm_n_lc < 2 || cm_n_voice < 2) {
  cat("**Cannot run 2-way test** — factor has < 2 levels.\n\n")
  cat(sprintf("Lexical: %d level(s) | Voice: %d level(s)\n", cm_n_lc, cm_n_voice))
  cat("Check Data Preparation output — HH/HL/LH/LL may not be found in Core_ID.\n")
  cm_test_ok <- FALSE; g_interaction_p <- NA_real_

} else if (g_is_parametric) {
  # PARAMETRIC: 2x4 Factorial ANOVA
  # WHY? All 8 cells passed Shapiro-Wilk (p >= 0.05) — every group is shaped like
  # a bell curve, so ANOVA assumptions are fully satisfied. ANOVA is the most
  # powerful test when normality holds.
  cat("### Test Chosen: **2 × 4 Factorial ANOVA**\n\n")
  cat("**Why?** All 8 Shapiro-Wilk tests returned p >= 0.05. Every group follows a ",
      "normal bell curve, so ANOVA assumptions are met. ANOVA is chosen because it ",
      "is the most powerful (sensitive) parametric test available.\n\n")

  g_fit <- aov(Score ~ Lexical_Condition * Voice, data=df_cm)
  g_sum <- summary(g_fit)[[1]]
  print(g_sum)
  g_interaction_p <- g_sum["Lexical_Condition:Voice","Pr(>F)"]
  cat(sprintf("\n**Interaction Effect p-value = %.4f** %s\n", g_interaction_p,
              ifelse(g_interaction_p<0.05,"← SIGNIFICANT","← not significant")))
  cm_test_ok <- TRUE

} else {
  # NON-PARAMETRIC: Scheirer-Ray-Hare Test
  # WHY? At least one cell failed Shapiro-Wilk (p < 0.05) — that group's data does
  # NOT follow a bell curve. A regular ANOVA assumes normality, so its p-values
  # would be unreliable. The Scheirer-Ray-Hare test converts all scores to RANKS
  # (1st, 2nd, 3rd...) and runs ANOVA on those ranks. Ranks need no normality assumption.
  cat("### Test Chosen: **Scheirer-Ray-Hare Test** (Non-Parametric 2-way ANOVA)\n\n")
  cat("**Why?** At least one cell failed Shapiro-Wilk (p < 0.05) — the data is ",
      "**Non-Parametric** (not a bell curve). A regular ANOVA would give unreliable ",
      "p-values. The Scheirer-Ray-Hare test converts scores to RANKS and runs ANOVA ",
      "on those ranks — no normality assumption needed.\n\n")

  df_cm$Rank_Score <- rank(df_cm$Score)
  g_fit <- aov(Rank_Score ~ Lexical_Condition * Voice, data=df_cm)
  g_sum <- summary(g_fit)[[1]]
  N      <- nrow(df_cm)
  ss_tot <- sum((df_cm$Rank_Score - mean(df_cm$Rank_Score))^2)

  g_srh <- tibble(
    Effect      = c("Lexical Condition","Voice","Interaction (Lexical × Voice)"),
    df          = g_sum[1:3,"Df"],
    H_statistic = round((g_sum[1:3,"Sum Sq"]/ss_tot)*(N-1),3),
    p_value     = round(pchisq((g_sum[1:3,"Sum Sq"]/ss_tot)*(N-1),g_sum[1:3,"Df"],lower.tail=FALSE),4),
    Significant = ifelse(pchisq((g_sum[1:3,"Sum Sq"]/ss_tot)*(N-1),
                                g_sum[1:3,"Df"],lower.tail=FALSE)<0.05,"YES (p < 0.05)","No")
  )

  g_srh %>%
    kable(caption="Scheirer-Ray-Hare Test Results — Interaction row highlighted",
          col.names=c("Effect","df","H statistic","p-value","Significant?"), booktabs=TRUE) %>%
    kable_styling(latex_options=c("striped","HOLD_position"),
                  bootstrap_options=c("striped","hover"), full_width=FALSE, position="center") %>%
    row_spec(0, bold=TRUE, background=HDR_COLOR, color="white") %>%
    row_spec(3, bold=TRUE, background="#EBF5FB")

  g_interaction_p <- g_srh$p_value[3]
  cat(sprintf("\n\n**Interaction Effect p-value = %.4f** %s\n", g_interaction_p,
              ifelse(g_interaction_p<0.05,"← SIGNIFICANT","← not significant")))
  cm_test_ok <- TRUE
}
```

---

## Step 3 — High-Accuracy Interaction Plot

> - **Faint dots** (`geom_jitter`, alpha=0.1) = every individual participant's score
> - **Thick lines** (`stat_summary fun=mean`) = group means
> - **Error bars** (`mean_cl_boot`) = 95% bootstrap confidence intervals

```{r g-interaction-plot, fig.width=8, fig.height=6, fig.cap="Enhanced Interaction Plot: jitter + mean lines + 95% CI."}
if (!isTRUE(cm_test_ok)) { cat("Plot skipped: test could not run.\n") } else {
  p_interaction <- ggplot(df_cm,
    aes(x=Lexical_Condition, y=Score, colour=Voice, group=Voice)) +
    geom_jitter(width=0.2, alpha=0.1, size=1.5) +           # BACKGROUND: raw scores
    stat_summary(fun=mean, geom="line", linewidth=1.2) +     # MIDDLE: mean trend lines
    stat_summary(fun.data=mean_cl_boot, geom="errorbar",     # FOREGROUND: 95% CI
                 width=0.2, linewidth=0.8) +
    stat_summary(fun=mean, geom="point",
                 size=4, shape=21, fill="white", stroke=1.5) +
    scale_colour_manual(values=PAL_VOICE, name="Voice") +
    labs(title="Interaction Plot: Voice × Lexical Condition",
         subtitle="Dots = individual participants | Lines = group means | Bars = 95% CI",
         x="Lexical Condition  (HH = easiest → LL = hardest)",
         y="Mean Corrected Memorability") +
    theme_minimal(base_size=14) +
    theme(legend.position="bottom", legend.title=element_text(face="bold"),
          plot.title=element_text(face="bold",hjust=0.5),
          plot.subtitle=element_text(colour="grey40",size=10,hjust=0.5))

  print(p_interaction)  # print() ensures this appears in the PDF
  ggsave(file.path(output_dir,"H7_Enhanced_Interaction.png"), p_interaction, width=8, height=6, dpi=150)
}
```

---

## Step 4 — The "Passive Penalty" Bar Chart

> **(Active mean) − (Passive mean)** per lexical type.
> Positive bar = Active > Passive (a cost for using passive voice).
> Labels show the exact percentage difference.

```{r g-penalty-plot, fig.width=7, fig.height=5, fig.cap="Passive Penalty (Active − Passive) with % labels."}
if (!isTRUE(cm_test_ok)) { cat("Penalty plot skipped: test could not run.\n") } else {
  g_diff <- df_cm %>%
    group_by(Lexical_Condition, Voice) %>%
    summarise(Mean_CM=mean(Score,na.rm=TRUE), .groups="drop") %>%
    pivot_wider(names_from=Voice, values_from=Mean_CM) %>%
    mutate(Passive_Penalty = Active - Passive,
           Pct_Label = paste0(ifelse(Passive_Penalty>=0,"+",""),
                              round(Passive_Penalty*100,1),"%"))

  p_penalty <- ggplot(g_diff,
    aes(x=Lexical_Condition, y=Passive_Penalty, fill=Lexical_Condition)) +
    geom_col(width=0.6, colour="black", alpha=0.85) +
    geom_hline(yintercept=0, linetype="dashed", colour="grey40") +
    geom_text(aes(label=Pct_Label,
                  vjust=ifelse(Passive_Penalty>=0,-0.5,1.5)),
              fontface="bold", size=4.5) +
    scale_fill_manual(values=PAL_TYPE, guide="none") +
    labs(title="Passive Penalty by Lexical Type",
         subtitle="Bar = (Active mean) − (Passive mean)  |  Labels = exact % difference",
         x="Lexical Condition", y="Passive Penalty (Active − Passive)") +
    theme_minimal(base_size=14) +
    theme(plot.title=element_text(face="bold",hjust=0.5),
          plot.subtitle=element_text(hjust=0.5,colour="grey40"))

  print(p_penalty)  # print() ensures this appears in the PDF
  ggsave(file.path(output_dir,"H7_Passive_Penalty.png"), p_penalty, width=7, height=5, dpi=150)
}
```

---

## Step 5 — Plain English Conclusion

```{r g-conclusion, echo=FALSE, results='asis'}
cat("### Does the difficulty of Passive voice significantly increase when the words are Low-Memorability (LL)?\n\n")

if (!isTRUE(cm_test_ok)) {
  cat("**Answer: Cannot determine** — the statistical test could not run.\n\n")
  cat("Most likely: HH/HL/LH/LL labels not found in Core_ID, or only one voice type in data.\n")

} else if (isTRUE(g_interaction_p < 0.05)) {
  cat(sprintf("**Answer: YES — interaction is statistically significant (p = %.4f).**\n\n", g_interaction_p))
  cat("The 'cost' of Passive voice is **not the same** across all lexical conditions. ",
      "The memorability drop caused by Passive voice **depends on** how hard the words are.\n\n")

  if (exists("g_diff") && nrow(g_diff)>0) {
    ll_row <- g_diff %>% filter(Lexical_Condition=="LL")
    hh_row <- g_diff %>% filter(Lexical_Condition=="HH")
    if (nrow(ll_row)>0 && nrow(hh_row)>0) {
      cat(sprintf("Passive Penalty for **LL** (both words hard) = **%s**, vs **%s** for **HH** (both easy). ",
                  ll_row$Pct_Label[1], hh_row$Pct_Label[1]))
      if (ll_row$Passive_Penalty[1] > hh_row$Passive_Penalty[1]) {
        cat("Passive voice hurts memory **more** when words are already difficult.\n\n")
      } else {
        cat("The pattern is more nuanced — passive penalty does not simply grow with difficulty.\n\n")
      }
    }
  }
  cat("**Hypothesis 7 is SUPPORTED.** Grammar and word-choice interact — not independent.\n")

} else {
  cat(sprintf("**Answer: NO — interaction is NOT significant (p = %.4f).**\n\n", g_interaction_p))
  cat("The cost of Passive voice is roughly **the same** for easy (HH) and hard (LL) words. ",
      "Voice and Lexical Condition act independently on memory.\n\n")
  cat("**Hypothesis 7 is NOT supported.**\n")
}
```

---

*Report generated on `r Sys.time()` using R `r R.version.string`*
