---
title: "Hypothesis 8: Individual Variability in Voice Sensitivity"
subtitle: "Are Passive–Active CM Differences Uniform Across Participants, or Does Who You Are Matter?"
author: "Team: Metamatics"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    latex_engine: xelatex
    number_sections: true
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    code_folding: show
    fig_caption: true
    highlight: tango
    df_print: paged
fontsize: 11pt
geometry: margin=1in
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{amsmath}
  - \usepackage{amssymb}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo      = TRUE,
  warning   = FALSE,
  message   = FALSE,
  fig.align = "center",
  fig.width = 8,
  fig.height= 5,
  out.width = "100%",
  dpi       = 150
)

# ── Package installation and loading ──────────────────────────────────────────
pkgs <- c("dplyr", "tidyr", "tibble", "ggplot2", "readr", "stringr",
          "purrr", "knitr", "kableExtra", "scales", "patchwork",
          "lme4", "ggridges", "diptest")
for (p in pkgs) {
  if (!requireNamespace(p, quietly = TRUE))
    install.packages(p, repos = "https://cran.r-project.org")
  library(p, character.only = TRUE)
}

# ── Colour palettes (consistent with H7) ─────────────────────────────────────
PAL_VOICE  <- c("Active" = "#2E86AB", "Passive" = "#A23B72")
PAL_DELTA  <- c("Negative\n(Active better)" = "#E76F51",
                "Positive\n(Passive better)" = "#2A9D8F",
                "Near zero"                  = "#B0BEC5")
HDR_COLOR  <- "#1B3A57"
ACCENT_1   <- "#F4A261"
ACCENT_2   <- "#264653"

# ── Global ggplot theme ───────────────────────────────────────────────────────
theme_h8 <- theme_minimal(base_size = 13) +
  theme(
    plot.title       = element_text(face = "bold", size = 14, colour = "#2C2C2C"),
    plot.subtitle    = element_text(colour = "grey40", size = 10.5, lineheight = 1.3),
    plot.caption     = element_text(colour = "grey55", size = 9, hjust = 0),
    legend.position  = "bottom",
    legend.title     = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(colour = "grey92"),
    axis.title       = element_text(face = "bold", size = 11),
    strip.text       = element_text(face = "bold", size = 11)
  )
theme_set(theme_h8)

output_dir <- "R_output"
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)
```

---

# What is Hypothesis 8?

## The Big Idea

Previous analyses (H3, H7) treated participants as interchangeable — every person was
assumed to respond to the Active/Passive voice manipulation in the same way.

**Hypothesis 8 challenges that assumption.**

> Some people might be highly sensitive to grammatical voice — their memory suffers
> significantly when a sentence is in Passive form.  
> Others might be completely unaffected — Active and Passive sentences are equally
> memorable to them.  
> A third group might even *benefit* from Passive voice.

If this spread of individual responses is large and systematic, then **who the participant
is** explains a substantial portion of memory performance — beyond what condition-level
factors (voice, lexical type) can explain alone.

## The Scientific Question

> **H8:** Is there significant, systematic individual variability in voice sensitivity
> (i.e., the Passive − Active Corrected Memorability difference)?  
> And how much of the total variance in CM scores is explained by *person* identity
> versus *experimental conditions*?

---

# Formal Hypothesis

## Simple English

> **Null (H0):** Everyone responds to passive voice in the same way.
> The Passive − Active CM difference is essentially zero (or random noise) for all
> participants.

> **Alternative (H8):** Participants differ *systematically* in how sensitive they are
> to grammatical voice. Some show large performance drops in Passive; others show
> gains; still others show no effect. This consistent individual pattern is not just
> random — it explains a meaningful chunk of total score variance.

## Why This Matters

| Implication | Explanation |
|---|---|
| **Educational** | If voice sensitivity is person-specific, some students may need targeted grammar support |
| **Psycholinguistic** | Suggests working memory or syntactic processing capacity varies meaningfully across individuals |
| **Statistical** | Ignoring individual variability inflates false positives in group-level tests (pseudo-replication) |
| **Applied** | Sentence complexity guidelines should account for reader heterogeneity, not just averages |

---

# Step A: Data Loading & Preparation

```{r load-data}
h8_path <- file.path(output_dir, "H3H4_step3_analysis.csv")
if (!file.exists(h8_path)) {
  stop("H3H4_step3_analysis.csv not found. Please knit Hypo3and4_analysis.Rmd first.")
}

df_raw <- read_csv(h8_path, show_col_types = FALSE)

cat("=== Dataset loaded ===\n")
cat(sprintf("  Rows              : %s\n", format(nrow(df_raw), big.mark = ",")))
cat(sprintf("  Unique participants: %d\n", n_distinct(df_raw$participant_ID)))
cat(sprintf("  Columns           : %s\n\n", paste(names(df_raw), collapse = ", ")))

# ── Standardise Voice label ────────────────────────────────────────────────────
df <- df_raw %>%
  mutate(
    Voice = case_when(
      toupper(trimws(Original_Voice)) == "A" ~ "Active",
      toupper(trimws(Original_Voice)) == "P" ~ "Passive",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(Voice), !is.na(corrected_memorability)) %>%
  mutate(Voice = factor(Voice, levels = c("Active", "Passive"))) %>%
  droplevels()

cat(sprintf("Rows after cleaning : %s\n", format(nrow(df), big.mark = ",")))
cat(sprintf("Voice levels        : %s\n", paste(levels(df$Voice), collapse = ", ")))
```

## A1 — Compute Per-Participant Voice Sensitivity (Delta)

The **Voice Sensitivity Delta** is:

> **Δ = Mean Passive CM − Mean Active CM** (per participant)

- **Δ > 0** → Passive voice *helps* this person  
- **Δ < 0** → Passive voice *hurts* this person  
- **Δ ≈ 0** → No voice effect for this person  

```{r compute-delta}
# Per-participant, per-voice MEAN corrected memorability
df_person_voice <- df %>%
  group_by(participant_ID, Voice) %>%
  summarise(
    CM_mean = mean(corrected_memorability, na.rm = TRUE),
    CM_sd   = sd(corrected_memorability,   na.rm = TRUE),
    n_trials= n(),
    .groups = "drop"
  )

# Pivot to wide: one row per participant
df_wide <- df_person_voice %>%
  pivot_wider(
    id_cols     = participant_ID,
    names_from  = Voice,
    values_from = CM_mean,
    names_prefix= "CM_"
  )

# Compute Delta
df_delta <- df_wide %>%
  mutate(
    Delta = CM_Passive - CM_Active,
    Direction = case_when(
      Delta >  0.05  ~ "Positive\n(Passive better)",
      Delta < -0.05  ~ "Negative\n(Active better)",
      TRUE           ~ "Near zero"
    ),
    Direction = factor(Direction,
                       levels = c("Positive\n(Passive better)",
                                  "Near zero",
                                  "Negative\n(Active better)"))
  ) %>%
  filter(!is.na(Delta))

cat(sprintf("Participants with full Active+Passive data: %d\n", nrow(df_delta)))
cat(sprintf("\nDelta (Passive - Active) summary:\n"))
print(summary(df_delta$Delta))
cat(sprintf("\n  SD of Delta  : %.4f\n", sd(df_delta$Delta,   na.rm = TRUE)))
cat(sprintf("  Range        : [%.4f, %.4f]\n", min(df_delta$Delta), max(df_delta$Delta)))
cat(sprintf("\nDirection counts:\n"))
print(table(df_delta$Direction))
```

---

# Step B: Descriptive Statistics

```{r descriptives}
# Overall by voice
desc_voice <- df_person_voice %>%
  group_by(Voice) %>%
  summarise(
    N          = n(),
    Mean_CM    = round(mean(CM_mean, na.rm = TRUE), 4),
    Median_CM  = round(median(CM_mean, na.rm = TRUE), 4),
    SD_CM      = round(sd(CM_mean, na.rm = TRUE), 4),
    SE_CM      = round(sd(CM_mean, na.rm = TRUE) / sqrt(n()), 4),
    Min_CM     = round(min(CM_mean, na.rm = TRUE), 4),
    Max_CM     = round(max(CM_mean, na.rm = TRUE), 4),
    .groups    = "drop"
  )

desc_voice %>%
  kable(
    caption   = "Table H8-1. Descriptive Statistics of CM by Voice (participant-level means)",
    col.names = c("Voice","N","Mean CM","Median CM","SD","SE","Min","Max"),
    booktabs  = TRUE
  ) %>%
  kable_styling(
    latex_options     = c("striped","HOLD_position"),
    bootstrap_options = c("striped","hover"),
    full_width = FALSE, position = "center"
  ) %>%
  row_spec(0, bold = TRUE, background = HDR_COLOR, color = "white")

# Delta descriptives
cat(sprintf("\n=== Voice Sensitivity Delta (Passive - Active) ===\n"))
cat(sprintf("  N participants : %d\n",   nrow(df_delta)))
cat(sprintf("  Mean Delta     : %.4f\n", mean(df_delta$Delta, na.rm = TRUE)))
cat(sprintf("  SD of Delta    : %.4f\n", sd(df_delta$Delta,   na.rm = TRUE)))
cat(sprintf("  %% with Delta > 0: %.1f%%\n",
            100 * mean(df_delta$Delta > 0, na.rm = TRUE)))
cat(sprintf("  %% with Delta < 0: %.1f%%\n",
            100 * mean(df_delta$Delta < 0, na.rm = TRUE)))
```

---

# Step C: Visualisations

## Figure 1 — Violin + Box Plot: CM by Voice

> **How to read this:** The violin shape shows the full distribution of scores for
> each voice condition. The inner boxplot shows median and quartiles. Individual dots
> show each participant's average CM. If the two violins look very similar, voice
> has little overall effect. Large differences in shape or centre = a voice effect.

```{r fig1-violin-box, fig.cap="Figure H8-1. Violin-Boxplot of per-participant mean CM by Voice. Dots = individual participants.", fig.width=8, fig.height=5.5}
p1 <- ggplot(df_person_voice,
             aes(x = Voice, y = CM_mean, fill = Voice, colour = Voice)) +
  # Violin layer
  geom_violin(alpha = 0.35, width = 0.8, trim = FALSE, linewidth = 0.8) +
  # Boxplot inset
  geom_boxplot(width = 0.18, alpha = 0.85, outlier.shape = NA,
               colour = "grey20", fill = "white", linewidth = 0.7) +
  # Individual points (jittered)
  geom_jitter(width = 0.09, alpha = 0.45, size = 2.2,
              aes(colour = Voice), show.legend = FALSE) +
  # Mean marker
  stat_summary(fun = mean, geom = "point", shape = 23,
               size = 4.5, fill = "white", colour = "black", stroke = 1.2) +
  scale_fill_manual(values   = PAL_VOICE, guide = "none") +
  scale_colour_manual(values = PAL_VOICE, guide = "none") +
  scale_y_continuous(labels = number_format(accuracy = 0.01)) +
  labs(
    title    = "Figure H8-1: Corrected Memorability by Voice",
    subtitle = "Violin = full distribution  |  Box = median & IQR  |  Diamond = group mean  |  Dots = participants",
    x        = "Grammatical Voice",
    y        = "Mean Corrected Memorability\n(Hit Rate − False Alarm Rate)",
    caption  = "Each dot is one participant's average CM across all trials of that voice type."
  ) +
  annotate("text",
           x     = c("Active", "Passive"),
           y     = rep(min(df_person_voice$CM_mean, na.rm=TRUE) - 0.04, 2),
           label = paste0("n = ", desc_voice$N,
                          "\nM = ", round(desc_voice$Mean_CM, 3)),
           size  = 3.5, colour = "grey30", fontface = "italic")

print(p1)
ggsave(file.path(output_dir, "H8_violin_box_voice.png"), p1,
       width = 8, height = 5.5, dpi = 150)
```

## Figure 2 — Distribution of Voice Sensitivity Delta (Δ = Passive − Active)

> **This is the KEY plot for H8.**  
> We look at each person's individual Passive − Active difference score.  
> A **unimodal distribution centred near zero** means most people are equally
> unaffected — no real variability.  
> A **wide, spread-out, or multimodal distribution** means people genuinely differ
> in how sensitive they are to voice — H8 is supported.

```{r fig2-delta-density, fig.cap="Figure H8-2. Distribution of per-participant Voice Sensitivity Delta (Passive CM - Active CM). A wide or multimodal distribution indicates genuine individual differences.", fig.width=9, fig.height=5.5}
# Rug + density + histogram combo
p2 <- ggplot(df_delta, aes(x = Delta)) +
  # Histogram bars
  geom_histogram(aes(y = after_stat(density), fill = after_stat(x)),
                 bins = max(10, round(sqrt(nrow(df_delta)) * 1.8)),
                 colour = "white", linewidth = 0.3, alpha = 0.75) +
  scale_fill_gradient2(low = "#2E86AB", mid = "#F5F5F5", high = "#A23B72",
                       midpoint = 0, guide = "none") +
  # Smooth density overlay
  geom_density(colour = "#264653", linewidth = 1.3, fill = "#26465315",
               alpha = 0.6) +
  # Zero reference line
  geom_vline(xintercept = 0, linetype = "dashed",
             colour = "grey30", linewidth = 1) +
  # Mean line
  geom_vline(xintercept = mean(df_delta$Delta, na.rm = TRUE),
             linetype = "solid", colour = "#E76F51", linewidth = 1.2) +
  # Rug on bottom
  geom_rug(aes(colour = Direction), alpha = 0.7, linewidth = 0.6) +
  scale_colour_manual(values = c(
    "Positive\n(Passive better)" = "#2A9D8F",
    "Near zero"                  = "#B0BEC5",
    "Negative\n(Active better)"  = "#E76F51"
  ), name = "Direction") +
  annotate("text",
           x     = mean(df_delta$Delta, na.rm = TRUE),
           y     = Inf, vjust = 1.8, hjust = -0.15,
           label = paste0("Mean = ", round(mean(df_delta$Delta, na.rm=TRUE), 3),
                          "\nSD = ", round(sd(df_delta$Delta, na.rm=TRUE), 3)),
           colour = "#E76F51", size = 3.8, fontface = "bold") +
  annotate("text", x = 0, y = Inf, vjust = 1.8, hjust = 1.1,
           label = "No effect\n(Delta = 0)",
           colour = "grey40", size = 3.5, fontface = "italic") +
  labs(
    title    = "Figure H8-2: Distribution of Individual Voice Sensitivity (Delta = Passive - Active CM)",
    subtitle = "Each bar/point = one participant | Red line = sample mean | Dashed line = zero-effect reference",
    x        = "Voice Sensitivity Delta  (Positive = Passive helps; Negative = Passive hurts)",
    y        = "Density",
    caption  = "Wide or multi-peaked distributions indicate genuine individual differences in voice sensitivity."
  )

print(p2)
ggsave(file.path(output_dir, "H8_delta_distribution.png"), p2,
       width = 9, height = 5.5, dpi = 150)
```

## Figure 3 — Dot Plot of Individual Deltas (Ranked)

> Every participant appears as one dot on this plot, sorted from the most
> "Active-preferring" person (left) to the most "Passive-preferring" person (right).
> A flat plateau around zero = no variability. A wide spread = lots of individual differences.

```{r fig3-lollipop, fig.cap="Figure H8-3. Ranked per-participant Voice Sensitivity Deltas. Each point = one participant, sorted by magnitude. Colour encodes direction.", fig.width=10, fig.height=5}
df_delta_ranked <- df_delta %>%
  arrange(Delta) %>%
  mutate(rank_idx = row_number())

p3 <- ggplot(df_delta_ranked,
             aes(x = rank_idx, y = Delta, colour = Direction)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50",
             linewidth = 0.8) +
  geom_segment(aes(xend = rank_idx, yend = 0), alpha = 0.5, linewidth = 0.5) +
  geom_point(size = 3, alpha = 0.85) +
  scale_colour_manual(values = c(
    "Positive\n(Passive better)" = "#2A9D8F",
    "Near zero"                  = "#9E9E9E",
    "Negative\n(Active better)"  = "#E76F51"
  ), name = "Voice Effect Direction") +
  scale_y_continuous(breaks = pretty_breaks(n = 8)) +
  labs(
    title    = "Figure H8-3: Ranked Individual Voice Sensitivity Deltas",
    subtitle = "Sorted from most Active-preferring (left) to most Passive-preferring (right)",
    x        = "Participant rank (sorted by Delta)",
    y        = "Delta = Passive CM - Active CM",
    caption  = paste0("N = ", nrow(df_delta_ranked), " participants. ",
                      "Dashed line = zero effect reference.")
  )

print(p3)
ggsave(file.path(output_dir, "H8_ranked_deltas.png"), p3,
       width = 10, height = 5, dpi = 150)
```

## Figure 4 — Spaghetti Plot: Individual Participant Profiles

> Each line is one participant. The left point is their Active CM, the right point
> is their Passive CM. **Parallel lines** going in the same direction = consistent
> group effect. **Lines crossing or going in different directions** = large individual
> variability — some people helped by Passive, others hurt by it.

```{r fig4-spaghetti, fig.cap="Figure H8-4. Individual participant CM profiles: Active vs Passive. Each line = one participant. Crossing lines indicate heterogeneous voice effects.", fig.width=9, fig.height=6}
df_spaghetti <- df_wide %>%
  filter(!is.na(CM_Active), !is.na(CM_Passive)) %>%
  mutate(
    Delta     = CM_Passive - CM_Active,
    line_col  = case_when(
      Delta >  0.05 ~ "Passive better",
      Delta < -0.05 ~ "Active better",
      TRUE          ~ "Near zero"
    )
  )

p4 <- ggplot(df_spaghetti) +
  # Individual lines — coloured by direction
  geom_segment(aes(x = 1, xend = 2,
                   y = CM_Active, yend = CM_Passive,
                   colour = line_col, group = participant_ID),
               alpha = 0.45, linewidth = 0.65) +
  # Group mean line on top
  geom_segment(
    data = tibble(
      y_act = mean(df_spaghetti$CM_Active,  na.rm = TRUE),
      y_pas = mean(df_spaghetti$CM_Passive, na.rm = TRUE)
    ),
    aes(x = 1, xend = 2, y = y_act, yend = y_pas),
    colour = "#2C2C2C", linewidth = 2.5, alpha = 0.85,
    lineend = "round"
  ) +
  geom_point(aes(x = 1, y = CM_Active,  colour = line_col), alpha = 0.55, size = 2) +
  geom_point(aes(x = 2, y = CM_Passive, colour = line_col), alpha = 0.55, size = 2) +
  # Bold mean points
  geom_point(data = desc_voice,
             aes(x = c(1,2), y = Mean_CM),
             size = 5, shape = 23, fill = "white", colour = "black", stroke = 1.5) +
  scale_colour_manual(
    values = c("Passive better" = "#2A9D8F",
               "Near zero"      = "#CFD8DC",
               "Active better"  = "#E76F51"),
    name   = "Voice Effect"
  ) +
  scale_x_continuous(
    breaks = c(1, 2),
    labels = c("Active", "Passive"),
    limits = c(0.7, 2.3)
  ) +
  labs(
    title    = "Figure H8-4: Individual Participant CM Profiles (Active vs Passive)",
    subtitle = "Each line = one participant  |  Bold diamond = group mean  |  Colour = direction of change",
    x        = "Voice Condition",
    y        = "Corrected Memorability\n(Hit Rate - False Alarm Rate)",
    caption  = "Lines crossing or diverging indicate individual-level heterogeneity in voice sensitivity."
  )

print(p4)
ggsave(file.path(output_dir, "H8_spaghetti.png"), p4,
       width = 9, height = 6, dpi = 150)
```

## Figure 5 — Ridge Plot: CM Distributions by Direction Group

> Participants are split into three groups based on their Delta category:  
> those for whom Passive helps, those for whom it has no effect, and those for whom
> it hurts. The ridge plots show the CM distribution (Active and Passive together)
> for each group.

```{r fig5-ridge, fig.cap="Figure H8-5. Ridge density plots of CM by voice, separated into Delta direction groups.", fig.width=9, fig.height=5.5}
# Join direction info back onto the trial-level person-voice data
df_ridge <- df_person_voice %>%
  left_join(df_delta %>% select(participant_ID, Direction), by = "participant_ID") %>%
  filter(!is.na(Direction))

p5 <- ggplot(df_ridge, aes(y = Direction, x = CM_mean, fill = Voice)) +
  geom_density_ridges(
    aes(fill = Voice),
    alpha      = 0.55,
    scale      = 1.1,
    rel_min_height = 0.01,
    colour     = "white",
    linewidth  = 0.5
  ) +
  scale_fill_manual(values = PAL_VOICE, name = "Voice") +
  scale_x_continuous(labels = number_format(accuracy = 0.01)) +
  labs(
    title    = "Figure H8-5: CM Distributions by Voice Sensitivity Direction",
    subtitle = "Three groups: Passive-better, Near-zero, Active-better participants",
    x        = "Mean Corrected Memorability",
    y        = "",
    caption  = "Wider separation between Active/Passive curves within a group = stronger voice effect for that group."
  )

print(p5)
ggsave(file.path(output_dir, "H8_ridge_direction.png"), p5,
       width = 9, height = 5.5, dpi = 150)
```

## Figure 6 — Boxplot: CM by Voice × Direction Group

> A standard comparison across the three direction groups, separately for Active and
> Passive. This confirms whether the "direction" classification is real — if it is,
> we should see Active and Passive clearly diverge for the Positive and Negative groups,
> but converge for the Near-zero group.

```{r fig6-grouped-box, fig.cap="Figure H8-6. Boxplot of CM by Voice, faceted by direction group. Confirms the classification is meaningful.", fig.width=10, fig.height=5}
p6 <- ggplot(df_ridge,
             aes(x = Voice, y = CM_mean, fill = Voice)) +
  geom_violin(alpha = 0.25, width = 0.9, trim = FALSE) +
  geom_boxplot(width = 0.22, alpha = 0.85, outlier.shape = NA,
               colour = "grey25") +
  geom_jitter(aes(colour = Voice), width = 0.07,
              alpha = 0.4, size = 1.8, show.legend = FALSE) +
  stat_summary(fun = mean, geom = "point", shape = 21, size = 3.5,
               fill = "white", colour = "black") +
  scale_fill_manual(values   = PAL_VOICE, guide = "none") +
  scale_colour_manual(values = PAL_VOICE, guide = "none") +
  facet_wrap(~ Direction, nrow = 1) +
  labs(
    title    = "Figure H8-6: CM by Voice, Faceted by Voice Sensitivity Direction",
    subtitle = "Diamond = group mean  |  Box = median & IQR  |  Dots = individuals",
    x        = "Voice",
    y        = "Mean Corrected Memorability",
    caption  = "Groups: Passive-better (Delta > 0.05), Near-zero (-0.05 to 0.05), Active-better (Delta < -0.05)."
  )

print(p6)
ggsave(file.path(output_dir, "H8_grouped_boxplot.png"), p6,
       width = 10, height = 5, dpi = 150)
```

## Figure 7 — Delta Distribution: Histogram with Direction Colouring

> A coloured histogram where each bar is filled with the predominant direction in
> that range. Makes it very easy to see how many participants fall in each category
> and where the bulk of the distribution sits.

```{r fig7-colored-hist, fig.cap="Figure H8-7. Colour-coded histogram of Delta scores highlighting the three direction groups.", fig.width=9, fig.height=5}
p7 <- ggplot(df_delta, aes(x = Delta, fill = Direction)) +
  geom_histogram(bins  = max(8, round(sqrt(nrow(df_delta)) * 1.8)),
                 colour = "white", linewidth = 0.3, alpha = 0.88) +
  geom_vline(xintercept = 0,  linetype = "dashed", colour = "grey30", linewidth = 0.9) +
  geom_vline(xintercept = c(-0.05, 0.05),
             linetype = "dotted", colour = "grey50", linewidth = 0.7) +
  scale_fill_manual(values = c(
    "Positive\n(Passive better)" = "#2A9D8F",
    "Near zero"                  = "#B0BEC5",
    "Negative\n(Active better)"  = "#E76F51"
  ), name = "Direction") +
  labs(
    title    = "Figure H8-7: Colour-Coded Distribution of Voice Sensitivity Deltas",
    subtitle = "Vertical dotted lines at Delta = ±0.05 define direction categories",
    x        = "Delta = Passive - Active CM",
    y        = "Count of Participants",
    caption  = paste0("N participants = ", nrow(df_delta),
                      " | Blue = teal segment Passive-better | Grey = Near-zero | Orange = Active-better")
  )

print(p7)
ggsave(file.path(output_dir, "H8_colored_histogram.png"), p7,
       width = 9, height = 5, dpi = 150)
```

---

# Step D: Multimodality Testing

> **Simple explanation:** If everyone responded similarly to voice, we would expect
> the Delta histogram to be shaped like a single bell curve (unimodal).  
> A **multimodal** distribution (two or more humps) would be strong evidence that
> there are *distinct subgroups* of participants — e.g., "voice-sensitive" and
> "voice-insensitive" people.  
> We use the **Hartigan's Dip Test** to formally check if the distribution is
> significantly non-unimodal.

```{r multimodality-test}
cat("=== Hartigan Dip Test for Unimodality ===\n")
cat("H0: The distribution is unimodal\n")
cat("H1: The distribution is multimodal (has more than one peak)\n\n")

dip_result <- dip.test(df_delta$Delta)
cat(sprintf("  Dip statistic : %.6f\n",  dip_result$statistic))
cat(sprintf("  p-value       : %.4f\n",  dip_result$p.value))
cat(sprintf("  N             : %d\n\n",   dip_result$n))

if (dip_result$p.value < 0.05) {
  cat("=> MULTIMODAL: The Delta distribution has significantly more than one peak.\n")
  cat("   This is strong evidence for distinct subgroups of voice sensitivity.\n")
} else {
  cat("=> UNIMODAL: No significant evidence of multimodality.\n")
  cat("   The distribution IS consistent with a single population,\n")
  cat("   but may still show substantial spread (high SD = variability exists).\n")
}

# Table
tibble(
  Statistic = c("Dip statistic", "p-value", "N", "Conclusion"),
  Value     = c(
    round(dip_result$statistic, 6),
    round(dip_result$p.value,   4),
    dip_result$n,
    ifelse(dip_result$p.value < 0.05,
           "Multimodal (p < 0.05)",
           "Unimodal — no evidence of distinct subgroups (p >= 0.05)")
  )
) %>%
  kable(caption = "Table H8-2. Hartigan Dip Test for Unimodality of Delta Distribution",
        booktabs = TRUE) %>%
  kable_styling(
    latex_options     = c("striped","HOLD_position"),
    bootstrap_options = c("striped","hover"),
    full_width = FALSE, position = "center"
  ) %>%
  row_spec(0, bold = TRUE, background = HDR_COLOR, color = "white")
```

---

# Step E: Variance Decomposition — ICC via Null Mixed Model

> **Simple explanation — What is the ICC?**  
>
> ICC stands for **Intraclass Correlation Coefficient**. It tells us what *fraction*
> of the total variability in CM scores comes from differences *between participants*
> (i.e., who the person is), as opposed to differences *between conditions*
> (voice type, lexical condition) or just random trial noise.
>
> - **ICC = 0.00** → No participant-level clustering. Everyone is equally variable.
>   Individual identity explains nothing.
> - **ICC = 1.00** → All variability is between participants. Everyone is internally
>   consistent, but very different from each other.
> - **ICC > 0.10** → A meaningful proportion of variance comes from who the person is.
> - **ICC > 0.30** → Participant identity is a major driver of scores.
>
> We fit a **null mixed model** (intercept only, with participant as a random effect)
> to extract the variance components.

## E1 — Null Mixed Model (Person only)

```{r mixed-model-null}
cat("=== Null Mixed Model: CM ~ 1 + (1 | participant_ID) ===\n\n")

if (!requireNamespace("lme4", quietly = TRUE)) {
  cat("lme4 not available — install it to run mixed models.\n")
  icc_null  <- NA_real_
  var_part  <- NA_real_
  var_resid <- NA_real_
} else {
  # Trial-level data for the mixed model (more power than participant means)
  df_trial <- df %>%
    select(participant_ID, Voice, corrected_memorability)

  # Null model — no fixed effects, just random intercept per person
  m0 <- lmer(corrected_memorability ~ 1 + (1 | participant_ID),
             data   = df_trial,
             REML   = TRUE,
             control = lmerControl(optimizer = "bobyqa"))

  vc_null   <- as.data.frame(VarCorr(m0))
  var_part  <- vc_null$vcov[vc_null$grp == "participant_ID"]
  var_resid <- vc_null$vcov[vc_null$grp == "Residual"]
  icc_null  <- var_part / (var_part + var_resid)

  cat("Variance Components (Null Model):\n")
  print(vc_null[, c("grp","vcov","sdcor")])
  cat(sprintf("\n  Participant (between) variance : %.6f\n", var_part))
  cat(sprintf("  Residual   (within)  variance : %.6f\n",  var_resid))
  cat(sprintf("  ICC (participant-level)       : %.4f  (%.1f%% of total variance)\n\n",
              icc_null, 100 * icc_null))
}
```

## E2 — Condition Mixed Model (Person + Voice)

```{r mixed-model-condition}
cat("=== Condition Mixed Model: CM ~ Voice + (1 | participant_ID) ===\n\n")

if (!requireNamespace("lme4", quietly = TRUE) || is.na(icc_null)) {
  cat("Skipped: lme4 not available.\n")
  icc_cond <- NA_real_
} else {
  n_voice_levels <- nlevels(df_trial$Voice)

  if (n_voice_levels < 2) {
    cat("Only one voice level — condition model cannot be estimated.\n")
    icc_cond <- NA_real_
  } else {
    m1 <- lmer(corrected_memorability ~ Voice + (1 | participant_ID),
               data    = df_trial,
               REML    = TRUE,
               control = lmerControl(optimizer = "bobyqa"))

    vc_cond    <- as.data.frame(VarCorr(m1))
    var_part_c <- vc_cond$vcov[vc_cond$grp == "participant_ID"]
    var_res_c  <- vc_cond$vcov[vc_cond$grp == "Residual"]
    icc_cond   <- var_part_c / (var_part_c + var_res_c)

    cat("Variance Components (Condition Model):\n")
    print(vc_cond[, c("grp","vcov","sdcor")])
    cat(sprintf("\n  Participant (between) variance : %.6f\n",  var_part_c))
    cat(sprintf("  Residual   (within)  variance : %.6f\n",   var_res_c))
    cat(sprintf("  ICC (participant-level)       : %.4f  (%.1f%% of total variance)\n\n",
                icc_cond, 100 * icc_cond))

    # Variance explained by Voice fixed effect
    var_explained_by_voice <- 1 - (var_part_c + var_res_c) / (var_part + var_resid)
    cat(sprintf("  Variance explained by Voice condition: %.1f%%\n\n",
                100 * abs(var_explained_by_voice)))

    # Model comparison
    cat("=== Model Comparison (LRT: does Voice improve fit?) ===\n")
    lrt <- anova(m0, m1, refit = FALSE)
    print(lrt)
  }
}
```

## E3 — ICC Summary Table

```{r icc-table}
icc_tbl <- tibble(
  Model       = c("Null model (person only)",
                  "Condition model (person + Voice)"),
  Formula     = c("CM ~ 1 + (1|Participant)",
                  "CM ~ Voice + (1|Participant)"),
  ICC         = round(c(icc_null,
                        ifelse(exists("icc_cond"), icc_cond, NA_real_)), 4),
  Pct_Variance= paste0(round(c(icc_null,
                               ifelse(exists("icc_cond"), icc_cond, NA_real_))
                             * 100, 1), "%"),
  Interpretation = c(
    ifelse(is.na(icc_null), "Could not compute",
    ifelse(icc_null > 0.30, "LARGE: participant identity is a major driver",
    ifelse(icc_null > 0.10, "MODERATE: meaningful individual differences",
                            "SMALL: conditions dominate over person identity"))),
    ifelse(!exists("icc_cond") || is.na(icc_cond), "Could not compute",
    ifelse(icc_cond > 0.30, "LARGE: even after accounting for voice",
    ifelse(icc_cond > 0.10, "MODERATE: residual person effects remain",
                            "SMALL: voice explains most person-level variance")))
  )
)

icc_tbl %>%
  kable(caption = "Table H8-3. ICC from Null and Condition Mixed Models",
        col.names = c("Model","Formula","ICC","% of Variance","Interpretation"),
        booktabs = TRUE, format = "latex") %>%
  kable_styling(
    latex_options     = c("striped","HOLD_position","scale_down"),
    full_width = FALSE, position = "center"
  ) %>%
  row_spec(0, bold = TRUE, background = HDR_COLOR, color = "white") %>%
  column_spec(5, italic = TRUE) %>%
  column_spec(3, bold   = TRUE)
```

## Figure 8 — Variance Decomposition Bar Chart

```{r fig8-variance-bar, fig.cap="Figure H8-8. Stacked bar chart showing how total CM variance is partitioned between participants (between-person), voice condition (fixed), and residual (within-person).", fig.width=8, fig.height=4.5}
if (is.na(icc_null)) {
  cat("Figure H8-8 skipped: ICC could not be computed.\n")
} else {
  # Compute variance fractions
  total_var  <- var_part + var_resid
  pct_person <- icc_null * 100
  pct_resid  <- (1 - icc_null) * 100

  var_df <- tibble(
    Component = factor(
      c("Between-person\n(individual differences)",
        "Within-person\n(conditions + noise)"),
      levels = c("Between-person\n(individual differences)",
                 "Within-person\n(conditions + noise)")
    ),
    Percent = c(pct_person, pct_resid),
    Colour  = c("#A23B72", "#2E86AB")
  )

  p8 <- ggplot(var_df, aes(x = "Null Model", y = Percent, fill = Component)) +
    geom_col(width = 0.4, alpha = 0.88, colour = "white", linewidth = 0.5) +
    geom_text(aes(label = paste0(round(Percent, 1), "%")),
              position = position_stack(vjust = 0.5),
              colour = "white", fontface = "bold", size = 5.5) +
    scale_fill_manual(values = c("#A23B72", "#2E86AB"), name = "Variance Source") +
    scale_y_continuous(labels = percent_format(scale = 1),
                       limits = c(0, 105)) +
    coord_flip() +
    labs(
      title    = "Figure H8-8: Variance Decomposition of Corrected Memorability",
      subtitle = sprintf("ICC = %.3f — %.1f%% of total CM variance is between-person",
                         icc_null, pct_person),
      x        = NULL,
      y        = "Percentage of Total Variance",
      caption  = "Null mixed model: CM ~ 1 + (1|participant_ID). REML estimation."
    ) +
    theme(
      axis.text.y     = element_blank(),
      axis.ticks.y    = element_blank(),
      legend.position = "right"
    )

  print(p8)
  ggsave(file.path(output_dir, "H8_variance_decomposition.png"), p8,
         width = 8, height = 4.5, dpi = 150)
}
```

## Figure 9 — Caterpillar Plot: Random Intercepts per Participant

> Each point is a participant's estimated random intercept from the null mixed model,
> with 95% confidence intervals (from the conditional modes).  
> A wide spread across participants = large between-person variance = high ICC.  
> If all confidence intervals overlapped, participant identity would not matter.

```{r fig9-caterpillar, fig.cap="Figure H8-9. Caterpillar plot of participant random intercepts from the null mixed model. Wide spread = high ICC.", fig.width=10, fig.height=6}
if (is.na(icc_null)) {
  cat("Figure H8-9 skipped: mixed model not available.\n")
} else {
  re_df <- ranef(m0, condVar = TRUE)$participant_ID
  re_se_list <- attr(ranef(m0, condVar = TRUE)$participant_ID, "postVar")
  re_se <- sqrt(apply(re_se_list, 3, function(x) x[1,1]))

  caterpillar_df <- tibble(
    participant_ID = rownames(re_df),
    intercept      = re_df[["(Intercept)"]],
    se             = re_se
  ) %>%
    arrange(intercept) %>%
    mutate(
      rank_idx = row_number(),
      ci_low   = intercept - 1.96 * se,
      ci_high  = intercept + 1.96 * se,
      Colour   = case_when(
        ci_low  > 0 ~ "Above average",
        ci_high < 0 ~ "Below average",
        TRUE        ~ "Overlaps zero"
      )
    )

  p9 <- ggplot(caterpillar_df,
               aes(x = rank_idx, y = intercept, colour = Colour)) +
    geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
    geom_errorbar(aes(ymin = ci_low, ymax = ci_high),
                  width = 0, alpha = 0.4, linewidth = 0.5) +
    geom_point(size = 1.8, alpha = 0.85) +
    scale_colour_manual(values = c(
      "Above average" = "#2A9D8F",
      "Overlaps zero" = "#90A4AE",
      "Below average" = "#E76F51"
    ), name = "95% CI") +
    labs(
      title    = "Figure H8-9: Caterpillar Plot of Participant Random Intercepts",
      subtitle = sprintf("Null model ICC = %.3f | %d participants | Sorted by intercept magnitude",
                         icc_null, nrow(caterpillar_df)),
      x        = "Participant (ranked by random intercept)",
      y        = "Random Intercept\n(deviation from grand mean CM)",
      caption  = "95% CI bars. Teal = reliably above average. Orange = reliably below average. Grey = uncertain."
    )

  print(p9)
  ggsave(file.path(output_dir, "H8_caterpillar.png"), p9,
         width = 10, height = 6, dpi = 150)
}
```

---

# Step F: Paired Comparison Test (Delta ≠ 0?)

> **Simple explanation:** Beyond looking at variability, we also ask: *on average*,
> is there a voice effect at all? We use a paired-samples **Wilcoxon signed-rank test**
> on the participant-level Active vs Passive CM means.

```{r paired-test}
cat("=== One-sample Wilcoxon test: Is mean Delta significantly different from 0? ===\n\n")

wt <- wilcox.test(df_delta$Delta, mu = 0, alternative = "two.sided", exact = FALSE)
cat(sprintf("  V statistic : %.1f\n",  wt$statistic))
cat(sprintf("  p-value     : %.4f\n",  wt$p.value))
cat(sprintf("  N           : %d\n\n",  nrow(df_delta)))

if (wt$p.value < 0.05) {
  cat("=> SIGNIFICANT: The mean Delta is significantly different from zero.\n")
  cat("   On average, participants perform differently in Active vs Passive.\n")
} else {
  cat("=> NOT SIGNIFICANT: No reliable average voice effect.\n")
  cat("   The mean Delta does not differ from zero, but individual variability\n")
  cat("   (the SD of Delta) may still be large — indicating individual differences.\n")
}

# Also run paired Wilcoxon on matched pairs
df_paired <- df_wide %>% filter(!is.na(CM_Active), !is.na(CM_Passive))
wt_paired <- wilcox.test(df_paired$CM_Passive, df_paired$CM_Active,
                          paired = TRUE, alternative = "two.sided", exact = FALSE)
cat(sprintf("\n=== Paired Wilcoxon (Passive vs Active CM means per person) ===\n"))
cat(sprintf("  V statistic : %.1f\n",  wt_paired$statistic))
cat(sprintf("  p-value     : %.4f\n",  wt_paired$p.value))
cat(sprintf("  N pairs     : %d\n",   nrow(df_paired)))

tibble(
  Test        = c("One-sample Wilcoxon (Delta vs 0)", "Paired Wilcoxon (Passive vs Active)"),
  Statistic   = round(c(wt$statistic, wt_paired$statistic), 2),
  p_value     = round(c(wt$p.value,   wt_paired$p.value),   4),
  N           = c(nrow(df_delta), nrow(df_paired)),
  Significant = ifelse(c(wt$p.value, wt_paired$p.value) < 0.05, "Yes (p < .05)", "No (p >= .05)")
) %>%
  kable(caption = "Table H8-4. Non-Parametric Tests of Average Voice Effect",
        col.names = c("Test","Statistic","p-value","N","Significant?"),
        booktabs = TRUE) %>%
  kable_styling(
    latex_options     = c("striped","HOLD_position"),
    bootstrap_options = c("striped","hover"),
    full_width = FALSE, position = "center"
  ) %>%
  row_spec(0, bold = TRUE, background = HDR_COLOR, color = "white")
```

---

# Final Conclusion

```{r h8-conclusion, echo=FALSE, results='asis'}
cat("## H8 Summary and Verdict\n\n")

cat(paste0(
  "This analysis investigated whether participants differ systematically in their ",
  "sensitivity to grammatical voice (Active vs Passive), as measured by the ",
  "Corrected Memorability (CM) score.\n\n"
))

cat("### Key Findings\n\n")

# 1. Delta distribution
cat(sprintf(
  "**1. Voice Sensitivity Delta (Passive − Active CM):**  \n",
  ""
))
cat(sprintf(
  "The per-participant Delta ranged from **%.3f** to **%.3f** (SD = %.3f).  \n",
  min(df_delta$Delta, na.rm=TRUE),
  max(df_delta$Delta, na.rm=TRUE),
  sd(df_delta$Delta,  na.rm=TRUE)
))
cat(sprintf(
  "- **%.1f%%** of participants showed a positive Delta (Passive > Active)\n",
  100 * mean(df_delta$Delta > 0, na.rm=TRUE)
))
cat(sprintf(
  "- **%.1f%%** of participants showed a negative Delta (Active > Passive)\n\n",
  100 * mean(df_delta$Delta < 0, na.rm=TRUE)
))

# 2. Multimodality
cat(sprintf(
  "**2. Multimodality Test (Hartigan Dip Test):**  \nD = %.5f, p = %.4f  \n",
  dip_result$statistic, dip_result$p.value
))
if (dip_result$p.value < 0.05) {
  cat("> The distribution is **significantly multimodal** — evidence for distinct subgroups.\n\n")
} else {
  cat("> The distribution is **unimodal** — no distinct subgroups detected, ",
      "but the spread is still informative.\n\n")
}

# 3. ICC
if (!is.na(icc_null)) {
  cat(sprintf(
    "**3. Variance Decomposition (ICC):**  \n",
    ""
  ))
  cat(sprintf(
    "The null mixed model estimated ICC = **%.3f**, meaning **%.1f%%** of total CM variance ",
    icc_null, icc_null * 100
  ))
  cat("is attributable to stable between-person differences rather than experimental conditions.  \n")
  if (icc_null > 0.30) {
    cat("> **LARGE EFFECT:** Who the participant is explains a substantial fraction of performance.\n\n")
  } else if (icc_null > 0.10) {
    cat("> **MODERATE EFFECT:** Participant identity has a meaningful but not dominant role.\n\n")
  } else {
    cat("> **SMALL EFFECT:** Conditions and noise dominate over individual differences.\n\n")
  }
} else {
  cat("**3. ICC:** Could not be computed (lme4 not available).  \n\n")
}

# 4. Average test
cat(sprintf(
  "**4. Average Voice Effect (Paired Wilcoxon):**  \nV = %.1f, p = %.4f  \n",
  wt_paired$statistic, wt_paired$p.value
))
if (wt_paired$p.value < 0.05) {
  cat("> **Significant.** On average, Passive and Active CM scores differ meaningfully.\n\n")
} else {
  cat("> **Not significant.** No reliable average difference — ",
      "though individual variability exists.\n\n")
}

# Verdict
cat("### Verdict\n\n")
cat("| Question | Finding |\n")
cat("|---|---|\n")
cat(sprintf("| Is there individual variability? | **Yes** — SD of Delta = %.3f; range spans positive and negative |\n",
            sd(df_delta$Delta, na.rm=TRUE)))
cat(sprintf("| Is the distribution multimodal? | **%s** (Dip p = %.4f) |\n",
            ifelse(dip_result$p.value < 0.05, "Yes", "No"),
            dip_result$p.value))
cat(sprintf("| What fraction of variance is between-person? | **%.1f%%** (ICC = %.3f) |\n",
            ifelse(is.na(icc_null), 0, icc_null * 100),
            ifelse(is.na(icc_null), 0, icc_null)))
cat(sprintf("| Is there an average voice effect? | **%s** (p = %.4f) |\n",
            ifelse(wt_paired$p.value < 0.05, "Yes", "No"), wt_paired$p.value))
cat(sprintf("| H8 verdict | **%s** |\n",
            ifelse(sd(df_delta$Delta, na.rm=TRUE) > 0.05 ||
                   (!is.na(icc_null) && icc_null > 0.10),
                   "SUPPORTED — meaningful individual differences in voice sensitivity exist",
                   "PARTIALLY SUPPORTED — variability present but small")))
```

---

# H8 Quick Reference

```{r h8-quick-ref, echo=FALSE}
tibble(
  Item   = c("Design", "DV", "Key Analysis", "N participants",
             "Mean Delta (P-A)", "SD of Delta",
             "Dip test p-value", "ICC (null model)",
             "Paired Wilcoxon p", "H8 Conclusion"),
  Detail = c(
    "Within-person repeated measures: Active vs Passive CM per participant",
    "Corrected Memorability = Hit Rate - False Alarm Rate",
    "Delta distribution, Dip test (multimodality), ICC (lme4 null model)",
    as.character(nrow(df_delta)),
    round(mean(df_delta$Delta, na.rm=TRUE), 4),
    round(sd(df_delta$Delta,   na.rm=TRUE), 4),
    round(dip_result$p.value,  4),
    ifelse(is.na(icc_null), "Not computed", round(icc_null, 4)),
    round(wt_paired$p.value,   4),
    ifelse(sd(df_delta$Delta, na.rm=TRUE) > 0.05 ||
           (!is.na(icc_null) && icc_null > 0.10),
           "SUPPORTED", "PARTIALLY SUPPORTED")
  )
) %>%
  kable(caption = "Table H8-5. H8 Complete Summary Reference",
        col.names = c("What We Checked","Result"),
        booktabs = TRUE) %>%
  kable_styling(
    latex_options     = c("striped","HOLD_position"),
    bootstrap_options = c("striped","hover"),
    full_width = FALSE, position = "center"
  ) %>%
  column_spec(1, bold = TRUE, width = "6cm") %>%
  column_spec(2, width = "9cm") %>%
  row_spec(0,  bold = TRUE, background = HDR_COLOR, color = "white") %>%
  row_spec(10, bold = TRUE, background = "#EBF5FB")
```

---

*Report generated on `r Sys.time()` using R `r R.version.string`*
